---
title: "making figures and combined plots of respirometry data and fvfm"
author: "Ninah Munk"
date: "2024-07-09"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(rstatix)
library(Rmisc)
library(lmerTest)
library(tidyverse)
library(janitor)
library(emmeans)
library(plyr)
library(ggthemes)
library(here)
library(pbkrtest)
library(sjPlot)
library(knitr)
library(kableExtra)
library(ggpattern)
library(hrbrthemes)
library(ggpubr)
```

# Load Data
```{r}

read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Respiration/Output/full.data.cleaned.csv")-> data
read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Data/worm_damage.csv")-> worm
```

Clean Data
```{r}
#clean data
corals_to_remove <- c(11, 5, 37, 104, 71, 17, 29, 100)
data <- data[!(data$coral_num %in% corals_to_remove), ]
```


More clean up + combine data frames
```{r}
data%>%rename(coral_id = coral_num)%>%select(-X) -> data

worm%>%select(coral_id, affected, percent_affected)-> worm.damage

data%>%left_join(worm.damage, by = "coral_id")-> data2

```

Make all fixed effects factors  
```{r}

data2%>%mutate(timepoint = as.factor(timepoint),
               temp = as.factor(temp),
               wound = as.factor(wound),
               genotype = as.factor(genotype)) -> data2

```

Filter data for less than 10% worm damage
```{r}
rates.filtered = subset(data2, percent_affected <= 10)
```

Subset respiration data, run lmer model, look at 3-way anova results, make pairwise comparisons
```{r}
#RESPIRATION
rates.filtered%>%filter(!timepoint == "day0")%>%filter(rate == "Respiration") -> resp.filtered

#model for respiration
mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint + (1|genotype/coral_id), data = resp.filtered)

anova(mod, type = 3) # significant temp*timepoint interaction

emmeans(mod, pairwise ~ temp*timepoint) #make pairwise comparisons

# checking for differences in groups pre-treatments
mod<- lmer(umol.cm2.hr ~ wound*temp + (1|genotype), data = rates.filtered%>%filter(rate == "Respiration")%>%filter(timepoint == "day0"))
anova(mod, type = 3)
```
Subset net photosynthesis data, run lmer model, look at 3-way anova results, make pairwise comparisons
```{r}
#NET PHOTOSYNTHESIS
rates.filtered%>%filter(!timepoint == "day0")%>%filter(rate == "Net Photosynthesis") -> netphoto.filtered

#model for net photosynthesis
mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint + (1|genotype/coral_id), data = netphoto.filtered)

anova(mod, type = 3) # significant temp*timepoint interaction

emmeans(mod, pairwise ~ wound*temp*timepoint) #make pairwise comparisons

# checking for differences in groups pre-treatments
mod<- lmer(umol.cm2.hr ~ wound*temp + (1|genotype), data = rates.filtered%>%filter(rate == "Net Photosynthesis")%>%filter(timepoint == "day0"))
anova(mod, type = 3)
```

Subset gross photosynthesis data, run lmer model, look at 3-way anova results, make pairwise comparisons
```{r}
#GROSS PHOTOSYNTHESIS
rates.filtered%>%filter(!timepoint == "day0")%>%filter(rate == "Gross Photosynthesis") -> grossphoto.filtered

#model for gross photosynthesis
mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint + (1|genotype/coral_id), data = grossphoto.filtered)

anova(mod, type = 3) # significant effect of time and temperature 

emmeans(mod, pairwise ~ temp*timepoint) #make pairwise comparisons

# checking for differences in groups pre-treatments
mod<- lmer(umol.cm2.hr ~ wound*temp + (1|genotype), data = rates.filtered%>%filter(rate == "Gross Photosynthesis")%>%filter(timepoint == "day0"))
anova(mod, type = 3)

```

calculate P:R, run lmer model, look at 3-way anova results, make pairwise comparisons
```{r}
rates.filtered%>%filter(!rate == "Net Photosynthesis")%>%select(rate, coral_id, temp, wound, genotype, timepoint, umol.cm2.hr)%>%pivot_wider(names_from = rate, values_from = umol.cm2.hr)%>%clean_names() -> rates.wide

rates.wide%>%mutate(ratio = ((11*gross_photosynthesis)/(24*respiration))) ->ratio

ratio%>%filter(!timepoint == "day0")-> ratio.filtered

#model for P:R
mod<- lmer(ratio ~ wound*temp*timepoint + (1|genotype/coral_id), data = ratio.filtered)

anova(mod, type = 3) #  significant temp*timepoint interaction

emmeans(mod, pairwise ~ temp*timepoint) #make pairwise comparisons

# checking for differences in groups pre-treatments
mod<- lmer(ratio ~ wound*temp + (1|genotype), data = ratio%>%filter(timepoint == "day0"))
anova(mod, type = 3)

```

```{r}
resp.filtered%>%mutate(timepoint = case_when(timepoint == "day0" ~ "0",
                                     timepoint == "day1" ~ "1",
                                     timepoint == "day10" ~ "10",
                                     timepoint == "day19" ~ "19"))%>%mutate(timepoint = as.factor(timepoint)) -> resp.filtered
```


Respiration figure
```{r}
# ggplot(data = resp.filtered, aes(x = timepoint, y = umol.cm2.hr, shape = temp, col = temp))+
#   stat_summary(fun=mean, geom="point", aes(group=temp), size = 7.5, position = position_dodge(width = 0.3))+
#   stat_summary(fun.data = mean_se, geom = "errorbar", aes(group=temp), width = 0.2, position = position_dodge(width = 0.3))+
#   stat_summary( fun = "mean", geom = "line", size = .4, position = position_dodge(width = 0.3), linetype = "dashed", aes(group = temp))+
#   ylab("O\u00b2 \u00b5mol/cm\u00b2/hr")+
#   xlab("Day")+
#   facet_wrap(~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")), ncol = 1, scales = "free")+ 
#   labs(fill = "Temperature")+
#  scale_shape_manual(values = c(1, 19), labels = c("Ambient", "Elevated"))+
#   scale_color_manual(values = c('blue', 'red'), labels = c("Ambient", "Elevated"))+
#   guides(col = guide_legend("Temperature"), shape = guide_legend("Temperature"))+
#   theme_base()+
#    theme_few(base_size = 30)+
#   theme(panel.spacing=unit(2,"lines"))+
#   theme(legend.position = "right" ) -> resp

ggplot(data = resp.filtered, aes(x = timepoint, y = umol.cm2.hr, shape = temp, col = temp))+
  stat_summary(fun=mean, geom="point", aes(group=temp), size = 7.5, position = position_dodge(width = 0.5))+
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(group=temp), width = 0.2, position = position_dodge(width = 0.5))+
  stat_summary( fun = "mean", geom = "line", size = .4, position = position_dodge(width = 0.5), linetype = "dashed", aes(group = temp))+
  ylab("O\u00b2 \u00b5mol/cm\u00b2/hr")+
  xlab("Day")+
  #facet_wrap(~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")), ncol = 1, scales = "free")+ 
  labs(fill = "Temperature")+
 scale_shape_manual(values = c(1, 19), labels = c("Ambient", "Warming"))+
  scale_color_manual(values = c('blue', 'red'), labels = c("Ambient", "Warming"))+
  guides(col = guide_legend("Temperature"), shape = guide_legend("Temperature"))+
  theme_base()+
   theme_few(base_size = 30)+
  theme(panel.spacing=unit(2,"lines"))+
  theme(legend.position = "none" ) -> resp

print(resp)

annotate_figure(
  resp,
  left = text_grob("Respiration",
                   color = "black", rot = 90, size = 30, face = "bold"),
   # bottom = text_grob("Time (Day)", color = "black",
   #                   face = "bold", size = 20)
  )-> resp.fig

ggsave("resp.fig2.png", plot = resp.fig, path = here(),
       width = 20,
       height = 15,
       units = "in")
 
```

```{r}
netphoto.filtered%>%mutate(timepoint = case_when(timepoint == "day0" ~ "0",
                                     timepoint == "day1" ~ "1",
                                     timepoint == "day10" ~ "10",
                                     timepoint == "day19" ~ "19"))%>%mutate(timepoint = as.factor(timepoint)) -> netphoto.filtered
```

net photosynthesis figure
```{r}
ggplot(data = netphoto.filtered, aes(x = timepoint, y = umol.cm2.hr, shape = temp, col = temp))+
  stat_summary(fun=mean, geom="point", aes(group=temp), size = 7.5, position = position_dodge(width = 0.5))+
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(group=temp), width = 0.2, position = position_dodge(width = 0.5))+
  stat_summary( fun = "mean", geom = "line", size = .4, position = position_dodge(width = 0.5), linetype = "dashed", aes(group = temp))+
  ylab("O\u00b2 \u00b5mol/cm\u00b2/hr")+
  xlab("Day")+
  facet_wrap(~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")), ncol = 1)+ 
  labs(fill = "Temperature")+
 scale_shape_manual(values = c(1, 19), labels = c("Ambient", "Warming"))+
  scale_color_manual(values = c('blue', 'red'), labels = c("Ambient", "Warming"))+
  guides(col = guide_legend("Temperature"), shape = guide_legend("Temperature"))+
  theme_base()+
   theme_few(base_size = 30)+
  theme(panel.spacing=unit(2,"lines"))+
  theme(legend.position = "right") -> net.photo.woundfacet

ggplot(data = netphoto.filtered, aes(x = timepoint, y = umol.cm2.hr, shape = wound, col = temp))+
  stat_summary(fun=mean, geom="point", aes(group=wound), size = 7.5, position = position_dodge(width = 0.5))+
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(group=wound), width = 0.2, position = position_dodge(width = 0.5))+
  stat_summary( fun = "mean", geom = "line", size = .4, position = position_dodge(width = 0.5), linetype = "dashed", aes(group = wound))+
  ylab("O\u00b2 \u00b5mol/cm\u00b2/hr")+
  xlab("Day")+
  #ylim(.2, 0.47)+
  facet_wrap(~temp, labeller = labeller(temp = c("H" = "Warming", "A" = "Ambient")), ncol = 1)+ 
  labs(fill = "Injury")+
 scale_shape_manual(values = c(16, 17, 15),labels = c("Non-injured", "Fragmentation", "Abrasion"))+
  scale_color_manual(values = c('blue', 'red'), labels = c("Ambient", "Warming"))+
  guides(col = guide_legend("Temperature"), shape = guide_legend("Injury"))+
  theme_base()+
   theme_few(base_size = 30)+
  theme(panel.spacing=unit(2,"lines"))+
  theme(legend.position = "right") ->net.photo.tempfacet


# annotate_figure(
#   net.photo.woundfacet,
#   left = text_grob("Net Photosynthesis",
#                    color = "black", rot = 90, size = 30, face = "bold"),
#    # bottom = text_grob("Time (Day)", color = "black",
#    #                   face = "bold", size = 20)
#   )-> net.photo.woundfacet.fig
# 
# annotate_figure(
#   net.photo.tempfacet,
#   left = text_grob("Net Photosynthesis",
#                    color = "black", rot = 90, size = 30, face = "bold"),
#    # bottom = text_grob("Time (Day)", color = "black",
#    #                   face = "bold", size = 20)
#   )-> net.photo.tempfacet.fig

# ggarrange(net.photo.woundfacet, net.photo.tempfacet,                                                # First row with scatter plot
#           ncol = 2, 
#           labels = c("A", "B")                                        # Labels of the scatter plot
#           ) -> fig2

annotate_figure(
  net.photo.woundfacet,
  left = text_grob("Net Photosynthesis",
                   color = "black", rot = 90, size = 30, face = "bold"),
   # bottom = text_grob("Time (Day)", color = "black",
   #                   face = "bold", size = 20)
  )-> net.photo.fig

ggsave("net.photo.fig2.png", plot = net.photo.fig, path = here(),
       width = 15,
       height = 20,
       units = "in")

```


```{r}
grossphoto.filtered%>%mutate(timepoint = case_when(timepoint == "day0" ~ "0",
                                     timepoint == "day1" ~ "1",
                                     timepoint == "day10" ~ "10",
                                     timepoint == "day19" ~ "19"))%>%mutate(timepoint = as.factor(timepoint)) -> grossphoto.filtered
```


Gross photosynthesis figure
```{r}

ggplot(data = grossphoto.filtered, aes(x = timepoint, y = umol.cm2.hr, shape = temp, col = temp))+
  stat_summary(fun=mean, geom="point", aes(group=temp), size = 7.5, position = position_dodge(width = 0.5))+
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(group=temp), width = 0.2, position = position_dodge(width = 0.5))+
  stat_summary( fun = "mean", geom = "line", size = .4, position = position_dodge(width = 0.5), linetype = "dashed", aes(group = temp))+
  ylab("O\u00b2 \u00b5mol/cm\u00b2/hr")+
  xlab("Day")+
 # facet_wrap(~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")), ncol = 1, scales = "free")+ 
  labs(fill = "Temperature")+
 scale_shape_manual(values = c(1, 19), labels = c("Ambient", "Warming"))+
  scale_color_manual(values = c('blue', 'red'), labels = c("Ambient", "Warming"))+
  guides(col = guide_legend("Temperature"), shape = guide_legend("Temperature"))+
  theme_base()+
   theme_few(base_size = 30)+
  theme(panel.spacing=unit(2,"lines"))+
  theme(legend.position = "none" ) -> grossp

print(grossp)

annotate_figure(
  grossp,
  left = text_grob("Gross Photosynthesis",
                   color = "black", rot = 90, size = 30, face = "bold"),
   # bottom = text_grob("Time (Day)", color = "black",
   #                   face = "bold", size = 20)
  )-> grossp.fig

ggsave("grossp.fig2.png", plot = grossp.fig, path = here(),
       width = 20,
       height = 15,
       units = "in")
 
```



```{r}
ratio.filtered%>%mutate(timepoint = case_when(timepoint == "day0" ~ "0",
                                     timepoint == "day1" ~ "1",
                                     timepoint == "day10" ~ "10",
                                     timepoint == "day19" ~ "19"))%>%mutate(timepoint = as.factor(timepoint)) -> ratio.filtered
```

P:R fig
```{r}
ggplot(data = ratio.filtered, aes(x = timepoint, y = ratio, col = temp, group = temp, shape = temp))+
  #geom_jitter(size = .5, alpha = .5)+
 stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, position = position_dodge(width = 0.5)) +
  stat_summary( fun = "mean", geom = "point", size = 7.5, position = position_dodge(width = 0.5), aes(group = temp))+
  stat_summary( fun = "mean", geom = "line", size = .4, position = position_dodge(width = 0.5), linetype = "dashed", aes(group = temp))+
 #geom_smooth(data=ratio.filtered, aes(x=timepoint, y=ratio, group = temp), alpha=0.3, method = "lm")+
  ylab(NULL)+
  xlab("Day")+
 #facet_grid(~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")), scales = "free_y", space="free")+ 
  theme_classic()+
  labs(color = "Temperature")+
  scale_color_manual(values = c('blue', 'red'), labels = c("Ambient", "Warming"))+
  scale_shape_manual(values = c(1, 19), labels = c("Ambient", "Warming"))+
   theme_few(base_size = 30) +
  theme(panel.spacing=unit(2,"lines"))+
  guides(color = guide_legend("Temperature"), shape = guide_legend("Temperature"))+
  theme(strip.text = element_blank())+ theme(legend.position = "right") -> PR

annotate_figure(
  PR,
  left = text_grob("Daily P:R",
                   color = "black", rot = 90, size = 30, face = "bold"),
   # bottom = text_grob("Time (Day)", color = "black",
   #                   face = "bold", size = 20)
  )-> PR

ggsave("PR.fig.png", plot = PR, path = here(),
       width = 20,
       height = 15,
       units = "in")
```

combining net photosynthesis, respiration, and P:R plots
```{r}
annotate_figure(
  resp,
  left = text_grob("Respiration",
                   color = "black", rot = 90, size = 20, face = "bold")
  ) -> resp.fig

annotate_figure(
  net.photo,
  left = text_grob("Net Photosynthesis",
                   color = "black", rot = 90, size = 20, face = "bold")
   # bottom = text_grob("Time (Day)", color = "black",
   #                   face = "bold", size = 20)
  )-> net.photo.fig

annotate_figure(
  PR,
  left = text_grob("Daily P:R",
                   color = "black", rot = 90, size = 20, face = "bold"),
   # bottom = text_grob("Time (Day)", color = "black",
   #                   face = "bold", size = 20)
  )-> PR.fig

ggarrange(resp.fig, net.photo.fig, PR.fig,                                                # First row with scatter plot
          nrow = 3, 
          labels = c("A", "B", "C")                                        # Labels of the scatter plot
          ) -> fig2


ggsave("R.NP.PR.fig.jpg", plot = fig2, path = here(),
       width = 16,
       height = 12,
       units = "in")

```
combining gross photosynthesis, respiration, and P:R plots
```{r}
ggarrange(resp.fig, grossp.fig, PR,                                                # First row with scatter plot
          nrow = 3, 
          labels = c("A", "B", "C"), font.label = list(size = 25)                                       # Labels of the scatter plot
          ) -> fig2


ggsave("R.GP.PR.fig.png", plot = fig2, path = here(),
       width = 20,
       height = 30,
       units = "in")

```

