---
title: "Respiration, photosynthesis, daily P:R"
author: "Ninah Munk"
date: "2024-06-24"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(rstatix)
library(Rmisc)
library(lmerTest)
library(tidyverse)
library(janitor)
library(emmeans)
library(plyr)
library(ggthemes)
library(here)
library(pbkrtest)
library(sjPlot)
library(knitr)
library(kableExtra)
library(ggpattern)
library(hrbrthemes)
library(ggpubr)
```

# Load Data
```{r}

read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Respiration/Output/full.data.cleaned.csv")-> data
read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Data/worm_damage.csv")-> worm
```

Clean Data
```{r}
#clean data
corals_to_remove <- c(11, 5, 37, 104, 71, 17, 29, 100)
data <- data[!(data$coral_num %in% corals_to_remove), ]
```


More clean up + combine data frames
```{r}
data%>%rename(coral_id = coral_num)%>%select(-X) -> data

worm%>%select(coral_id, affected, percent_affected)-> worm.damage

data%>%left_join(worm.damage, by = "coral_id")-> data2

```

Change timepoint to be a continuous variable  
```{r}

data2%>%mutate(timepoint = case_when(timepoint == "day0" ~ "0",
                                     timepoint == "day1" ~ "1",
                                     timepoint == "day10" ~ "10",
                                     timepoint == "day19" ~ "19"))%>%mutate(timepoint = as.numeric(timepoint),
                                                                            temp = as.factor(temp),
                                                                            wound = as.factor(wound),
                                                                            genotype = as.factor(genotype)) -> data2

```

Filter data for less than 10% worm damage
```{r}
rates.filtered = subset(data2, percent_affected <= 10)
```

# RESPIRATION
Full LMM for respiration rates. This is not including the pre-experiment timepoint (day 0/-7). 
wound, temp, timepoint 3-way interaction is highly non-significant 
```{r}
rates.filtered%>%filter(!timepoint == 0)%>%filter(rate == "Respiration") -> resp.filtered

mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint + (1|genotype/coral_id), data = resp.filtered, contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3) 

```
Reduced LMM for respiration rates. 
** Temp and time interaction is significant ** 
```{r}
mod<- lmer(umol.cm2.hr ~ wound*temp + temp*timepoint + (1|genotype/coral_id), data = resp.filtered, contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3) 
```
Checking respiration rates prior to onset of experiment (day 0/-7). 
** no significant differences **
```{r}
mod<- lmer(umol.cm2.hr ~ wound*temp  + (1|genotype), data = rates.filtered%>%filter(rate == "Respiration")%>%filter(timepoint == 0), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod)
```
Model for respiration at day 1
```{r}
# Subset data for a single time point
time1 <- resp.filtered%>%filter(timepoint == 1)
# Mixed-effects model for a single time point
model_time1 <- lmer(umol.cm2.hr ~ temp*wound + (1|genotype), data=time1, contrasts = list(wound = "contr.sum", temp = "contr.poly"))
summary(model_time1)
# ANOVA on the mixed model
anova(model_time1)
# Post-hoc tests
emmeans(model_time1, pairwise ~ temp*wound)
# Visualize data
ggplot(time1, aes(x=temp, y=umol.cm2.hr, color=temp)) + 
 # geom_point() + 
  stat_summary(fun=mean, geom="point", aes(group=temp), size=5)+
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, position = position_dodge(width = 0.4))

```

Model for respiration at day 10
```{r}
time2 <- resp.filtered%>%filter(timepoint == 10)
model_time2 <- lmer(umol.cm2.hr ~ wound*temp + (1|genotype), data=time2, contrasts = list(wound = "contr.sum", temp = "contr.poly"))
summary(model_time2)
anova(model_time2)
ggplot(time2, aes(x=temp, y=umol.cm2.hr, color=temp)) + 
  stat_summary(fun=mean, geom="point", aes(group=temp), size=5)+
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, position = position_dodge(width = 0.4)) 
```

Model for respiration at day 19
```{r}
time3 <- resp.filtered%>%filter(timepoint == 19)
model_time3 <- lmer(umol.cm2.hr ~ temp*wound + (1|genotype), data=time2, contrasts = list(wound = "contr.sum", temp = "contr.poly"))
summary(model_time3)
anova(model_time3)
ggplot(time3, aes(x=temp, y=umol.cm2.hr, color=temp)) + 
  stat_summary(fun=mean, geom="point", aes(group=temp), size=5)+
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, position = position_dodge(width = 0.4))

```


Respiration figure that pairs with result from full lmer model with time. 
```{r}

ggplot(data = resp.filtered, aes(x = timepoint, y = umol.cm2.hr, col = temp, linetype = temp))+
  #geom_jitter(size = .5, alpha = .5)+
  #stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, position = position_dodge(width = 0.4)) +
 # stat_summary( fun = "mean", geom = "point", size = 2, position = position_dodge(width = 0.4))+ 
  #stat_summary( fun = "mean", geom = "line", size = .5, position = position_dodge(width = 0.4), linetype = "dashed", aes(group = temp))+
  geom_smooth(data=resp.filtered, aes(x=timepoint, y=umol.cm2.hr, col = temp, linetype = temp), alpha=0.3, method = "lm")+
  ylab("Respiration (O\u00b2 \u00b5mol/cm\u00b2/hr)")+
  xlab("Time")+
 # facet_grid(~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")), scales = "free_y", space="free")+ 
  theme_classic()+
  labs(color = "Temperature")+
  scale_color_manual(values = c('blue', 'red'), labels = c("Ambient", "Elevated"))+
   theme_few(base_size = 12)+
  annotate("text", x=12, y=.44, label= "p < 0.01", size = 6)+
  scale_linetype_discrete(name="Temperature", 
                          breaks=c("A", "H"), 
                          labels = c("Ambient", "Elevated"))+
  guides(col = guide_legend("Temperature"), linetype = guide_legend("Temperature"))+ 
  theme(legend.position = "none") -> resp


ggsave("resp.png", plot = resp, path = here(),
       width = 10,
       height = 8,
       units = "in")
```

resp figure that pairs with net photo figure
```{r}
ggplot(data = resp.filtered, aes(x = timepoint, y = umol.cm2.hr, shape = temp, col = temp))+
  stat_summary(fun=mean, geom="point", aes(group=temp), size = 3, position = position_dodge(width = 0.3))+
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(group=temp), width = 0.2, position = position_dodge(width = 0.3))+
  stat_summary( fun = "mean", geom = "line", size = .4, position = position_dodge(width = 0.3), linetype = "dashed", aes(group = temp))+
  ylab("O\u00b2 (\u00b5mol/cm\u00b2/hr)")+
  xlab(NULL)+
  facet_grid(~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")), scales = "free")+ 
  labs(fill = "Temperature")+
 scale_shape_manual(values = c(1, 19), labels = c("Ambient", "Elevated"))+
  scale_color_manual(values = c('blue', 'red'), labels = c("Ambient", "Elevated"))+
  guides(col = guide_legend("Temperature"), shape = guide_legend("Temperature"))+
  theme_base()+
   theme_few(base_size = 20)+
  theme(panel.spacing=unit(1,"lines"))+
  theme(legend.position = "none" ) -> resp
        #legend.text = element_text(size = 12), 
   # legend.title = element_text(size = 14)
```


# NET PHOTOSYNTHESIS
Full LMM model for net photosynthetic rates.
** significant 3-way interaction between wound, temp, and time. **
```{r}
rates.filtered%>%filter(!timepoint == 0)%>%filter(rate == "Net Photosynthesis") -> netphoto.filtered
#full model 
mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint + (1|genotype/coral_id), data = netphoto.filtered, contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3) 

```

Checking net photosynthesis rates prior to onset of experiment (day 0/-7). 
** no significant differences **
```{r}
mod<- lmer(umol.cm2.hr ~ wound*temp  + (1|genotype), data = rates.filtered%>%filter(rate == "Net Photosynthesis")%>%filter(timepoint == 0), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3)
```
model for net photosynthesis at day 1
```{r}
# Subset data for a single time point
time1 <- netphoto.filtered%>%filter(timepoint == 1)
# Mixed-effects model for a single time point
model_time1 <- lmer(umol.cm2.hr ~ wound*temp + (1|genotype), data=time1, contrasts = list(wound = "contr.sum", temp = "contr.poly"))
summary(model_time1)
# ANOVA on the mixed model
anova(model_time1)
# Visualize data
ggplot(time1, aes(x=temp, y=umol.cm2.hr, color=temp)) + 
 # geom_point() + 
  stat_summary(fun=mean, geom="point", aes(group=temp), size=5)+
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, position = position_dodge(width = 0.4))

```
Model for net photosynthesis at day 10
```{r}
time2 <- netphoto.filtered%>%filter(timepoint == 10)
model_time2 <- lmer(umol.cm2.hr ~ wound*temp + (1|genotype), data=time2, contrasts = list(wound = "contr.sum", temp = "contr.poly"))
summary(model_time2)
anova(model_time2)
# Post-hoc tests
emmeans(model_time2, pairwise ~ temp*wound)
ggplot(time2, aes(x=temp, y=umol.cm2.hr, color=temp)) + 
  stat_summary(fun=mean, geom="point", aes(group=temp), size=5)+
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, position = position_dodge(width = 0.4))+ facet_wrap(~wound)
```
model for net photosynthesis at day 19
```{r}
time3 <- netphoto.filtered%>%filter(timepoint == 19)
model_time3 <- lmer(umol.cm2.hr ~ wound*temp + (1|genotype), data=time3, contrasts = list(wound = "contr.sum", temp = "contr.poly"))
summary(model_time3)
anova(model_time3)
# Post-hoc tests
emmeans(model_time3, pairwise ~ temp*wound)
```
Net photosynthesis figure that pairs with result from model
```{r}

ggplot(data = netphoto.filtered, aes(x = timepoint, y = umol.cm2.hr, shape = temp, linetype = temp))+
  geom_smooth(data=netphoto.filtered, aes(x=timepoint, y=umol.cm2.hr, col = temp, linetype = temp), alpha=0.3, method = "lm")+
  ylab("Net Photosynthesis (O\u00b2 \u00b5mol/cm\u00b2/hr)")+
  xlab("Time (Day)")+
 facet_grid(~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")), scales = "free_y", space="free")+ 
  theme_classic()+
  labs(color = "Temperature")+
  scale_color_manual(values = c('blue', 'red'), labels = c("Ambient", "Elevated"))+
   theme_few(base_size = 12)+
 # annotate("text", x=12, y=.44, label= "p < 0.01", size = 6)+
  scale_linetype_discrete(name="Temperature", 
                          breaks=c("A", "H"), 
                          labels = c("Ambient", "Elevated"))+
  guides(col = guide_legend("Temperature"), linetype = guide_legend("Temperature"))+
  theme(legend.position = "right")-> net.photo

annotate_figure(
  net.photo,
  top = text_grob("p = 0.035",
                   color = "black", size = 15)
  ) -> net.photo.fig

ggsave("net.photo.jpg", plot = net.photo.fig, path = here(),
       width = 10,
       height = 8,
       units = "in")
```


Net photosynthesis figure that pairs with respiration figure
```{r}

ggplot(data = netphoto.filtered, aes(x = timepoint, y = umol.cm2.hr, shape = temp, col = temp))+
  stat_summary(fun=mean, geom="point", aes(group=temp), size = 3, position = position_dodge(width = 0.3))+
  stat_summary(fun.data = mean_se, geom = "errorbar", aes(group=temp), width = 0.2, position = position_dodge(width = 0.3))+
  stat_summary( fun = "mean", geom = "line", size = .4, position = position_dodge(width = 0.3), linetype = "dashed", aes(group = temp))+
  ylab("Net photosynthesis (O\u00b2 \u00b5mol/cm\u00b2/hr)")+
  xlab("Time (day)")+
  facet_grid(~wound, labeller = labeller(wound = c("0" = " ", "1" = " ", "2" = " ")), scales = "free")+ 
  labs(fill = "Temperature")+
 scale_shape_manual(values = c(1, 19), labels = c("Ambient", "Elevated"))+
  scale_color_manual(values = c('blue', 'red'), labels = c("Ambient", "Elevated"))+
  guides(col = guide_legend("Temperature"), shape = guide_legend("Temperature"))+
  theme_base()+
   theme_few(base_size = 15)+
  theme(panel.spacing=unit(1,"lines"))+
  theme(legend.position = "bottom") -> net.photo
 

```
summary statistics for NP
```{r}
# Compared to control corals at elevated temperature, final mean  NP is  1.21 and 1.17 times greater in fragmented and abraded corals, respectively.

#average NP control corals day19: 0.3171588
#average NP frag corals day 19: 0.3734911
#average NP abraded corals day 19: 0.3832824


#take the larger num and divide by smaller number
0.3832824/0.3171588 # 1.208487 
0.3734911/0.3171588 # 1.177615

# By the end of the study, mean NP increased x at elevated temperature in control corals, while NP increased by x and x% in fragmented and abrased corals, respectively.

#average NP control corals day 1: 0.2986437
#average NP frag corals day 1: 0.2982166
#average NP abraded corals day 19: 0.2748130
# (final - initial) / initial
#control: 
  ((0.3171588 - 0.2986437) / 0.2986437) * 100
# fragment
  ((0.3734911 - 0.2982166) / 0.2982166) * 100
# abrasion
  ((0.3832824 - 0.2748130) / 0.2748130) * 100
# Compared to control corals at elevated temperature, NP increased x and x times more in fragmented and abraded corals, respectively after 19 days. 

# Pooled across injury status, day 10 NP was x times greater than day 1 R at elevated temperature. Pooled across injury status, the difference in mean NP between temperatures at day 10 was x times greater than the difference between mean R at day 1. 

# ambient day 1 mean NP pooled across injury: 0.2905473
# elevated day 1 mean NP pooled across injury: 0.2905696
# ambient day 10 mean NP pooled across injury:0.2694341
# elevated day 10 mean NP pooled across injury: 0.3871514

0.3871514 - 0.2694341 #NP difference day 10: 0.1177173
0.4458603 - 0.4009609 # R difference day 1: 0.0448994

0.1177173 / 0.0448994 # 2.621801 times higher NP than R

# elevated day 1 mean R pooled across injury: 0.4458603
# ambient day 1 mean R pooled across injury: 0.4009609
summarySE(netphoto.filtered, measurevar = "umol.cm2.hr", groupvars = c("temp", "timepoint"),
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE) 

summarySE(resp.filtered, measurevar = "umol.cm2.hr", groupvars = c("temp", "timepoint"),
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE) 
```



# GROSS PHOTOSYNTHESIS

Full LMM model for gross photosynthetic rates.
** timepoint is significant **
```{r}
rates.filtered%>%filter(!timepoint == 0)%>%filter(rate == "Gross Photosynthesis") -> grossphoto.filtered

mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint + (1|genotype/coral_id), data = grossphoto.filtered%>%filter(!timepoint ==0), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3) 
```

Checking gross photosynthesis rates prior to onset of experiment (day 0/-7)
** no significant differences **
```{r}
mod<- lmer(umol.cm2.hr ~ wound*temp  + (1|genotype), data = rates.filtered%>%filter(rate == "Gross Photosynthesis")%>%filter(timepoint == 0), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3)
```
 Gross Photosynthesis figure that pairs with results from model. 
```{r}
ggplot(data = grossphoto.filtered, aes(x = timepoint, y = umol.cm2.hr))+
  geom_smooth(data=grossphoto.filtered, aes(x=timepoint, y=umol.cm2.hr), alpha=0.3, method = "lm")+
  #geom_jitter(size = .5, alpha = .5)+
  # stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, position = position_dodge(width = 0.2)) +
  # stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.2))+ 
  # stat_summary( fun = "mean", geom = "line", size = 1, position = position_dodge(width = 0.2), linetype = "dashed")+
  ylab("Gross Photosynthesis (O\u00b2 \u00b5mol/cm\u00b2/hr)")+
  xlab("Time (Day)")+
 # facet_grid(~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")), scales = "free_y", space="free")+ 
  theme_classic()+
  #labs(color = "Temperature")+
  #scale_color_manual(values = c('blue', 'red'), labels = c("Ambient (~28)", "Elevated (~30)"))+
   theme_few(base_size = 12)+
  annotate("text", x=12, y=.77, label= "p = 0.04", size = 6) -> gross.photo


ggsave("gross.photo.jpg", plot = gross.photo, path = here(),
       width = 10,
       height = 8,
       units = "in")


```

summary statistics for GP
```{r}
# GP day 1: 0.7142854
# GP day 19: 0.7537166
((0.7537166 - 0.7142854)/0.7142854)*100

summarySE(grossphoto.filtered, measurevar = "umol.cm2.hr", groupvars = c( "temp", "timepoint"),
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE)
```


# DAILY P:R RATIO

wrangle data frame and calculate ratio
```{r}
rates.filtered%>%filter(!rate == "Net Photosynthesis")%>%select(rate, coral_id, temp, wound, genotype, timepoint, umol.cm2.hr)%>%pivot_wider(names_from = rate, values_from = umol.cm2.hr)%>%clean_names() -> rates.wide

rates.wide%>%mutate(ratio = ((11*gross_photosynthesis)/(24*respiration))) ->ratio
```

Full LMM model for daily P:R 
** significant interaction between temp and time ** 
```{r}
ratio%>%filter(!timepoint == 0)-> ratio.filtered

mod<- lmer(ratio ~ wound*temp*timepoint + (1|genotype/coral_id), data = ratio.filtered, contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3)

```

Checking daily P:R prior to onset of experiment (day 0/-7)
** no significant differences **
```{r}
mod<- lmer(ratio ~ wound*temp + (1|genotype), data = ratio%>%filter(timepoint == 0), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3)
```
model for P:R day 1
```{r}
# Subset data for a single time point
time1 <- ratio.filtered%>%filter(timepoint == 1)
# Mixed-effects model for a single time point
model_time1 <- lmer(ratio ~ wound*temp + (1|genotype), data=time1, contrasts = list(wound = "contr.sum", temp = "contr.poly"))
summary(model_time1)
# ANOVA on the mixed model
anova(model_time1)

``` 
 model for P:R day 10
```{r}
time2 <- ratio.filtered%>%filter(timepoint == 10)
model_time2 <- lmer(ratio ~ wound*temp + (1|genotype), data=time2, contrasts = list(wound = "contr.sum", temp = "contr.poly"))
summary(model_time2)
anova(model_time2)
emmeans(model_time2, pairwise ~ temp*wound) # no significant pairwise comparisons

ggplot(time2, aes(x=temp, y=ratio, color=temp)) + 
  stat_summary(fun=mean, geom="point", aes(group=temp), size=5)+
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, position = position_dodge(width = 0.4)) 

```
 
 P:R model for day 19
```{r}
time3 <- ratio.filtered%>%filter(timepoint == 19)
model_time3 <- lmer(ratio ~ wound*temp + (1|genotype), data=time3, contrasts = list(wound = "contr.sum", temp = "contr.poly"))
summary(model_time3)
anova(model_time3)
emmeans(model_time3, pairwise ~ temp*wound)
```
 
 Daily P:R figure that pairs with result from model
```{r}
ggplot(data = ratio.filtered, aes(x = timepoint, y = ratio, col = temp, linetype = temp))+
 #  #geom_jitter(size = .5, alpha = .5)+
 # stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, position = position_dodge(width = 0.3)) +
 #  stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.3), aes(group = temp))+
 #  stat_summary( fun = "mean", geom = "line", size = .4, position = position_dodge(width = 0.3), linetype = "dashed", aes(group = temp))+
 geom_smooth(data=ratio.filtered, aes(x=timepoint, y=ratio, col = temp, linetype = temp), alpha=0.3, method = "lm")+
  ylab("Daily P:R")+
  xlab("Time (Day)")+
 #facet_grid(~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")), scales = "free_y", space="free")+ 
  theme_classic()+
  labs(color = "Temperature")+
  scale_color_manual(values = c('blue', 'red'), labels = c("Ambient", "Elevated"))+
 # scale_shape_manual(values = c(1, 19), labels = c("Ambient (~28)", "Elevated (~30)"))+
   theme_few(base_size = 12) +
  theme(panel.spacing=unit(1,"lines"))+
  theme(strip.text = element_blank())+
  annotate("text", x=12, y=.92, label= "p < 0.01", size = 6)+
  scale_linetype_discrete(name="Temperature", 
                          breaks=c("A", "H"), 
                          labels = c("Ambient", "Elevated"))+
  guides(col = guide_legend("Temperature"), linetype = guide_legend("Temperature"))+ 
  theme(legend.position = "none") -> PR


ggsave("PR.png", plot = PR, path = here(),
       width = 12,
       height = 8,
       units = "in")
```

mean ratios figure
```{r}
ggplot(data = ratio.filtered, aes(x = timepoint, y = ratio, col = temp, group = temp, shape = temp))+
  #geom_jitter(size = .5, alpha = .5)+
 stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, position = position_dodge(width = 0.3)) +
  stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.3), aes(group = temp))+
  stat_summary( fun = "mean", geom = "line", size = .4, position = position_dodge(width = 0.3), linetype = "dashed", aes(group = temp))+
 #geom_smooth(data=ratio.filtered, aes(x=timepoint, y=ratio, group = temp), alpha=0.3, method = "lm")+
  ylab("Daily P:R")+
  xlab("Time")+
 facet_grid(~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")), scales = "free_y", space="free")+ 
  theme_classic()+
  labs(color = "Temperature")+
  scale_color_manual(values = c('blue', 'red'), labels = c("Ambient (~28)", "Elevated (~30)"))+
  scale_shape_manual(values = c(1, 19), labels = c("Ambient (~28)", "Elevated (~30)"))+
   theme_few(base_size = 20) +
  theme(panel.spacing=unit(1,"lines"))+
  guides(color = guide_legend("Temperature"), shape = guide_legend("Temperature"))+
  theme(strip.text = element_blank())+ theme(legend.position = "none")

print(PR)
ggsave("PR.png", plot = PR, path = here(),
       width = 12,
       height = 8,
       units = "in")
```


# FIGURES 

```{r}
rates.filtered%>%filter(!timepoint == 0) -> rates.filtered.2

ggplot(data = rates.filtered.2, aes(x = timepoint, y = umol.cm2.hr, col = temp, shape = temp))+
  #geom_smooth(data=rates.filtered.2, aes(x=timepoint, y=umol.cm2.hr), alpha=0.3, method = "lm")+
  #geom_jitter(size = .5, alpha = .5)+
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, position = position_dodge(width = 0.2)) +
   stat_summary( fun = "mean", geom = "point", size = 2.5, position = position_dodge(width = 0.2), aes(group = temp))+
   stat_summary( fun = "mean", geom = "line", size = .5, position = position_dodge(width = 0.2), linetype = "dashed", aes(group = temp))+
  ylab("Oxygen Flux (umol/cm2/hr)")+
  xlab("Day")+
  facet_grid(rate~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")), scales = "free_y", space="free")+ 
  theme_classic()+
  labs(color = "Temperature")+
  scale_color_manual(values = c('blue', 'red'), labels = c("Ambient (~28)", "Elevated (~30)"))+
  scale_shape_manual(values = c(1, 19), labels = c("Ambient (~28)", "Elevated (~30)"))+
   theme_few(base_size = 15) +
  guides(color = guide_legend("Temperature"), shape = guide_legend("Temperature"))+ 
  theme(legend.position = c(0.75, 0.93))


#save plot

# ggsave("all.rates.png", plot = all.rates, path = here(),
#        width = 10,
#        height = 8,
#        units = "in")
```

```{r}
respmod<- lmer(umol.cm2.hr ~ wound*temp + temp*timepoint + (1|genotype/coral_id), data = resp.filtered)
netphotomod<- lmer(umol.cm2.hr ~ wound*temp*timepoint + (1|genotype/coral_id), data = netphoto.filtered)
grossmod<- lmer(umol.cm2.hr ~ wound*temp*timepoint  + (1|genotype/coral_id), data = grossphoto.filtered)

tab_model(respmod, netphotomod, grossmod)

anova(respmod) -> aov
anova(netphotomod)-> aov2

install.packages("xtable")
require(xtable)
print(xtable(aov, aov2), type = "html")

```
combining net photosynthesis, respiration, and P:R plots
```{r}
annotate_figure(
  resp,
  left = text_grob("Respiration",
                   color = "black", rot = 90, size = 20, face= "bold")
  ) -> resp.fig

annotate_figure(
  net.photo,
  left = text_grob("Net Photosynthesis",
                   color = "black", rot = 90, size = 20, face= "bold"),
   bottom = text_grob("Time (Day)", color = "black",
                     face = "bold", size = 20)
  )-> net.photo.fig

ggarrange(
  resp.fig, net.photo.fig,  ncol = 1, labels = c("A", " B"),
  common.legend = TRUE, legend = "bottom") -> resp.photo.fig

#save plot

ggexport(resp.photo.fig, filename = "resp.photo.fig.png")

ggsave("resp.photo.fig.jpg", plot = resp.photo.fig, path = here(),
       width = 12,
       height = 8,
       units = "in")

```

arrange plots displaying model results
```{r}
ggarrange(net.photo,                                                 # First row with scatter plot
          ggarrange(resp, gross.photo, PR, ncol = 3, labels = c("B", "C", "D")), # Second row with box and dot plots
          nrow = 2, 
          labels = "A"                                        # Labels of the scatter plot
          ) -> rates.model.figure

annotate_figure(
  rates.model.figure,
  top = text_grob("p = 0.35",
                   color = "black", size = 20)
  ) -> rates.model.figure.2

ggsave("model.fig.2.jpg", plot = rates.model.figure.2, path = here(),
       width = 12,
       height = 8,
       units = "in")
```

