---
title: "Rates Analysis"
author: "Ninah Munk"
date: "2024-06-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rstatix)
library(Rmisc)
library(lmerTest)
library(tidyverse)
library(janitor)
library(emmeans)
library(plyr)
library(ggthemes)
library(here)
library(pbkrtest)
library(sjPlot)
library(knitr)
library(kableExtra)
```

# Load and Clean Data
```{r}

read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Respiration/Output/full.data.cleaned.csv")-> data
read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Data/worm_damage.csv")-> worm
```

# Clean Data
```{r}
#clean data
corals_to_remove <- c(11, 5, 37, 104, 71, 17, 29, 100)
data <- data[!(data$coral_num %in% corals_to_remove), ]
```


Clean up some data and combine data frames
```{r}
data%>%rename(coral_id = coral_num)%>%select(-X) -> data

worm%>%select(coral_id, affected, percent_affected)-> worm.damage

data%>%left_join(worm.damage, by = "coral_id")-> data2

```


Change timepoint to a continuous variable and numeric 
```{r}

data2%>%mutate(timepoint = case_when(timepoint == "day0" ~ "0",
                                     timepoint == "day1" ~ "1",
                                     timepoint == "day10" ~ "10",
                                     timepoint == "day19" ~ "19"))%>%mutate(timepoint = as.numeric(timepoint),
                                                                            temp = as.factor(temp),
                                                                            wound = as.factor(wound),
                                                                            genotype = as.factor(genotype)) -> data2

```


# plotting rates against worm damage. All rates are reduced with increasing worm damage.
```{r}
ggplot(data2, aes(x = percent_affected, y = umol.cm2.hr))+geom_point()+geom_smooth(method = "lm")+facet_grid( rate ~ timepoint)

```

# RESPIRATION 
Run Models: The full model includes fixed effects wound, temp, and timepoint. Percent_affected is included as a covariate. 
```{r}
#full model 
mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint*percent_affected + (1|genotype/coral_id), data = data2%>%filter(rate == "Respiration"), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3)

```

Filter out worms: Filter data to remove worm effect determined from growth data. this threshold is below 10 percent worm damage.
```{r}
rates.filtered = subset(data2, percent_affected <= 10)
```

Check that percent affected is no longer significant in full model
```{r}

mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint*percent_affected + (1|genotype/coral_id), data = rates.filtered%>%filter(rate == "Respiration"), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3) #percent affected is not longer significant 


mod<- lmer(umol.cm2.hr ~ wound*temp + temp*timepoint*percent_affected + (1|genotype/coral_id), data = rates.filtered%>%filter(rate == "Respiration"))
anova(mod, type = 3)# percent affected not significant in reduced model
```


Run models with filtered data without percent affected as a covariate. There is a significant interaction between temperature and timepoint in all three models. This becomes more significant after filtering out initial timepoint, so this tells me this interaction is driven primarily by the following timepoints. 
```{r}
#full model 
mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint + (1|genotype/coral_id), data = rates.filtered%>%filter(rate == "Respiration")%>%filter(!timepoint == 0), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3) # significant interaction between temp and time


```

Model for Respiration broken into each timepoint 
```{r}
# DAY 0 
mod<- lmer(umol.cm2.hr ~ wound*temp  + (1|genotype), data = rates.filtered%>%filter(rate == "Respiration")%>%filter(timepoint == 0), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3)

# DAY 1
mod<- lmer(umol.cm2.hr ~ wound*temp  + (1|genotype), data = rates.filtered%>%filter(rate == "Respiration")%>%filter(timepoint == 1), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3)
lsmeans(mod, pairwise~wound*temp, adjust="tukey")

# DAY 10
mod<- lmer(umol.cm2.hr ~ wound*temp  + (1|genotype), data = rates.filtered%>%filter(rate == "Respiration")%>%filter(timepoint == 10), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3) 

# DAY 19
mod<- lmer(umol.cm2.hr ~ wound*temp  + (1|genotype), data = rates.filtered%>%filter(rate == "Respiration")%>%filter(timepoint == 19), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3) # temp is highly significant
```

```{r}

rates.filtered <- rates.filtered %>%
  mutate(wound2=case_when(
    wound== "0" ~ "Control",
    wound== "1" ~ "Injured",
    wound == "2"~ "Injured"))

mod<- lmer(umol.cm2.hr ~ wound2*temp  + (1|genotype), data = rates.filtered%>%filter(rate == "Respiration")%>%filter(timepoint == 1), contrasts = list(wound2 = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3)

lsmeans(mod, pairwise~wound2*temp, adjust="tukey")

```


```{r}
rates.filtered%>%filter(rate == "Respiration")%>%filter(timepoint == 1) -> mean.resp
mean.resp%>%unite("treatment", c("wound2", "temp"), remove = FALSE) -> mean.resp

# Calculating mean growth for each group to plot.
mean.resp%>%group_by(wound2, temp)%>%summarize(Mean = mean(umol.cm2.hr, na.rm=TRUE),
                                                        'N'=length(umol.cm2.hr ))-> mean

mean.resp%>%aggregate(umol.cm2.hr ~ treatment, FUN = function(x) sd(x)/sqrt(length(x)))%>%separate(col = treatment, into = c("wound2", "temp")) -> std_err

# Assuming all contrasts were significant for simplicity; this may change based on actual results.
mean$letter <- c("abc", "ac", "b","ca")

dodge<-position_dodge(width=0.6) # this offsets the points so they don't overlap

  ggplot(mean, aes(x = temp, y = Mean, colour = temp, fill = temp))+
  facet_wrap(~wound2, labeller = labeller(wound = c("Control" = "No Injury", "Injured" = "Injury")))+
  geom_bar(stat="identity", position="dodge") +geom_errorbar(aes(ymin=Mean-std_err$umol.cm2.hr, ymax=Mean+std_err$umol.cm2.hr), width=0, position=position_dodge(width = 0.9))+
    geom_text(aes(label=letter), nudge_y=0.32, color="black")+
  geom_point(data = mean.resp, aes(x = temp, y = umol.cm2.hr, colour = temp), shape = 21, alpha = 0.7,
             position = position_jitterdodge(jitter.width = 0.3, jitter.height=0.2, dodge.width=0.9))+
  scale_color_manual(values = c('black', 'black'), guide="none")+
  scale_fill_manual(values = c('white', 'grey'), labels=c("Ambient", "Elevated"), guide = "none")+
  scale_x_discrete(labels = c('Ambient','Elevated'))+
xlab(expression("Temperature"))+
  ylab(expression(paste("Respiration Rate (mg/cm2/day)")))+
  theme(axis.text.x=element_text(colour="black"))+
  theme(axis.text.y=element_text(colour="black"))+theme_few(base_size = 12)
```


```{r}
# Respiration figure that pairs with full model
ggplot(data = rates.filtered%>%filter(!timepoint == 0)%>%filter(rate == "Respiration"), aes(x = timepoint, y = umol.cm2.hr, col = temp))+
  #geom_jitter(size = .5, alpha = .5)+
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, position = position_dodge(width = 0.2)) +
  stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.2))+ 
  stat_summary( fun = "mean", geom = "line", size = 1, position = position_dodge(width = 0.2), linetype = "dashed", aes(group = temp))+
  ylab("O2 Rate (umol/cm2/hr)")+
  xlab("Day")+
  #facet_grid(rate~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")), scales = "free_y", space="free")+ 
  theme_classic()+
  labs(color = "Temperature")+
  scale_color_manual(values = c('blue', 'red'), labels = c("Ambient (~28)", "Elevated (~30)"))+
   theme_few(base_size = 20)
```



```{r}
rates.filtered%>%filter(rate == "Respiration")%>%filter(timepoint == 1) -> mean.resp
mean.resp%>%unite("treatment", wound:temp, remove = FALSE) -> mean.resp

# Calculating mean growth for each group to plot.
mean.resp%>%group_by(wound, temp)%>%summarize(Mean = mean(umol.cm2.hr, na.rm=TRUE),
                                                        'N'=length(umol.cm2.hr ))-> mean

mean.resp%>%aggregate(umol.cm2.hr ~ treatment, FUN = function(x) sd(x)/sqrt(length(x)))%>%separate(col = treatment, into = c("wound", "temp")) -> std_err

# Assuming all contrasts were significant for simplicity; this may change based on actual results.
mean$letter <- c("abc", "abc", "a","bc", "abc", "bc")

dodge<-position_dodge(width=0.6) # this offsets the points so they don't overlap

  ggplot(mean, aes(x = temp, y = Mean, colour = temp, fill = temp))+
  facet_wrap(~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")))+
  geom_bar(stat="identity", position="dodge") +geom_errorbar(aes(ymin=Mean-std_err$umol.cm2.hr, ymax=Mean+std_err$umol.cm2.hr), width=0, position=position_dodge(width = 0.9))+
    geom_text(aes(label=letter), nudge_y=0.32, color="black")+
  geom_point(data = mean.resp, aes(x = temp, y = umol.cm2.hr, colour = temp), shape = 21, alpha = 0.7,
             position = position_jitterdodge(jitter.width = 0.3, jitter.height=0.2, dodge.width=0.9))+
  scale_color_manual(values = c('black', 'black', 'black'), guide="none")+
  scale_fill_manual(values = c('white', 'grey'), labels=c("Ambient", "Elevated"), guide = "none")+
  scale_x_discrete(labels = c('Ambient','Elevated'))+
xlab(expression("Temperature"))+
  ylab(expression(paste("Respiration Rate (mg/cm2/day)")))+
  theme(axis.text.x=element_text(colour="black"))+
  theme(axis.text.y=element_text(colour="black"))+theme_few(base_size = 12)

# print(figure_1)
# 
# #save plot
# 
# ggsave("figure_1.png", plot = figure_1, path = here(),
#        width = 8,
#        height = 8,
#        units = "in")
```


# NET PHOTOSYNTHESIS
Run Models: The full model includes fixed effects wound, temp, and timepoint. Percent_affected is included as a covariate. 
```{r}
#full model 
mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint*percent_affected + (1|genotype/coral_id), data = data2%>%filter(rate == "Net Photosynthesis"), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3)

#reduced
mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint + percent_affected + (1|genotype/coral_id), data = data2%>%filter(rate == "Net Photosynthesis"), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3)

```

Filter data to remove worm effect determined from growth data. this threshold is below 10 percent worm damage. check that percent affected is no longer significant in full model
```{r}
rates.filtered = subset(data2, percent_affected <= 10)

mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint*percent_affected + (1|genotype/coral_id), data = rates.filtered%>%filter(rate == "Net Photosynthesis"), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3) #percent affected is not longer significant 

#reduced
mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint + temp*timepoint*percent_affected + wound*temp*percent_affected + (1|genotype/coral_id), data = rates.filtered%>%filter(rate == "Net Photosynthesis"))
anova(mod, type = 3)

#reduced again
mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint + temp*timepoint*percent_affected + (1|genotype/coral_id), data = rates.filtered%>%filter(rate == "Net Photosynthesis"))
anova(mod, type = 3)# percent affected not significant in reduced model
```

Run models with filtered data without percent affected as a covariate.
```{r}
#full model 
mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint + (1|genotype/coral_id), data = rates.filtered%>%filter(rate == "Net Photosynthesis")%>%filter(!timepoint ==0), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3) # significant interaction between temp, time

lsmeans(mod, pairwise~wound*temp*timepoint, adjust="tukey")

```

Model for Net Photosynthesis broken into each timepoint 
```{r}
# DAY 0
mod<- lmer(umol.cm2.hr ~ wound*temp  + (1|genotype), data = rates.filtered%>%filter(rate == "Net Photosynthesis")%>%filter(timepoint == 0), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3)

# DAY 1
mod<- lmer(umol.cm2.hr ~ wound*temp  + (1|genotype), data = rates.filtered%>%filter(rate == "Net Photosynthesis")%>%filter(timepoint == 1), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3) 

# DAY 10
mod<- lmer(umol.cm2.hr ~ wound*temp  + (1|genotype), data = rates.filtered%>%filter(rate == "Net Photosynthesis")%>%filter(timepoint == 10), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3) # temp is highly significant

lsmeans(mod, pairwise~wound*temp, adjust="tukey")

# DAY 19
mod<- lmer(umol.cm2.hr ~ wound*temp  + (1|genotype), data = rates.filtered%>%filter(rate == "Net Photosynthesis")%>%filter(timepoint == 19), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3) #  temp*wound interaction highly significant

lsmeans(mod, pairwise~wound*temp, adjust="tukey")
```


```{r}
# Net photosynthesis figure that pairs with full model
ggplot(data = rates.filtered%>%filter(!timepoint == 0)%>%filter(rate == "Net Photosynthesis"), aes(x = timepoint, y = umol.cm2.hr, col = temp))+
  #geom_jitter(size = .5, alpha = .5)+
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, position = position_dodge(width = 0.2)) +
  stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.2))+ 
  stat_summary( fun = "mean", geom = "line", size = 1, position = position_dodge(width = 0.2), linetype = "dashed", aes(group = temp))+
  ylab("O2 Rate (umol/cm2/hr)")+
  xlab("Day")+
  facet_grid(rate~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")), scales = "free_y", space="free")+ 
  theme_classic()+
  labs(color = "Temperature")+
  scale_color_manual(values = c('blue', 'red'), labels = c("Ambient (~28)", "Elevated (~30)"))+
   theme_few(base_size = 20)
```


```{r}
rates.filtered%>%filter(rate == "Net Photosynthesis")%>%filter(timepoint == 10) -> mean.netphoto
mean.netphoto%>%unite("treatment", wound:temp, remove = FALSE) -> mean.netphoto

# Calculating mean growth for each group to plot.
mean.netphoto%>%group_by(wound, temp)%>%summarize(Mean = mean(umol.cm2.hr, na.rm=TRUE),
                                                        'N'=length(umol.cm2.hr ))-> mean

mean.netphoto%>%aggregate(umol.cm2.hr ~ treatment, FUN = function(x) sd(x)/sqrt(length(x)))%>%separate(col = treatment, into = c("wound", "temp")) -> std_err

# Assuming all contrasts were significant for simplicity; this may change based on actual results.
mean$letter <- c("ac", "b", "a","abc", "abc", "b")

dodge<-position_dodge(width=0.6) # this offsets the points so they don't overlap

  ggplot(mean, aes(x = temp, y = Mean, colour = temp, fill = temp))+
  facet_wrap(~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")))+
  geom_bar(stat="identity", position="dodge") +geom_errorbar(aes(ymin=Mean-std_err$umol.cm2.hr, ymax=Mean+std_err$umol.cm2.hr), width=0, position=position_dodge(width = 0.9))+
    geom_text(aes(label=letter), nudge_y=0.32, color="black")+
  geom_point(data = mean.netphoto, aes(x = temp, y = umol.cm2.hr, colour = temp), shape = 21, alpha = 0.7,
             position = position_jitterdodge(jitter.width = 0.3, jitter.height=0.2, dodge.width=0.9))+
  scale_color_manual(values = c('black', 'black', 'black'), guide="none")+
  scale_fill_manual(values = c('white', 'grey'), labels=c("Ambient", "Elevated"), guide = "none")+
  scale_x_discrete(labels = c('Ambient','Elevated'))+
xlab(expression("Temperature"))+
  ylab(expression(paste("Net Photosynthesis Rate (mg/cm2/day)")))+
  theme(axis.text.x=element_text(colour="black"))+
  theme(axis.text.y=element_text(colour="black"))+theme_few(base_size = 12)

#print(figure_1)

#save plot

# ggsave("figure_1.png", plot = figure_1, path = here(),
#        width = 8,
#        height = 8,
#        units = "in")
```

# GROSS PHOTOSYNTHESIS
Run Models: The full model includes fixed effects wound, temp, and timepoint. Percent_affected is included as a covariate. 
```{r}
#full model 
mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint*percent_affected + (1|genotype/coral_id), data = data2%>%filter(rate == "Gross Photosynthesis"), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3)

#reduced
mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint + percent_affected + (1|genotype/coral_id), data = data2%>%filter(rate == "Gross Photosynthesis"), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3)

```

Filter data to remove worm effect determined from growth data. this threshold is below 10 percent worm damage. check that percent affected is no longer significant in full model
```{r}
rates.filtered = subset(data2, percent_affected <= 10)

mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint*percent_affected + (1|genotype/coral_id), data = rates.filtered%>%filter(rate == "Gross Photosynthesis"), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3) #percent affected is not longer significant 

#reduced
mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint + timepoint*percent_affected + temp*percent_affected + (1|genotype/coral_id), data = rates.filtered%>%filter(rate == "Gross Photosynthesis"))
anova(mod, type = 3)# percent affected not significant in reduced model
```

Run models with filtered data without percent affected as a covariate.
```{r}
#full model 
mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint + (1|genotype/coral_id), data = rates.filtered%>%filter(rate == "Gross Photosynthesis")%>%filter(!timepoint ==0), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3) # temperature and timepoint is significant

lsmeans(mod, pairwise~wound*temp, adjust="tukey")

```

Model for Net Photosynthesis broken into each timepoint 
```{r}
# DAY 0
mod<- lmer(umol.cm2.hr ~ wound*temp  + (1|genotype), data = rates.filtered%>%filter(rate == "Gross Photosynthesis")%>%filter(timepoint == 0), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3)

# DAY 1
mod<- lmer(umol.cm2.hr ~ wound*temp  + (1|genotype), data = rates.filtered%>%filter(rate == "Gross Photosynthesis")%>%filter(timepoint == 1), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3) 

# DAY 10
mod<- lmer(umol.cm2.hr ~ wound*temp  + (1|genotype), data = rates.filtered%>%filter(rate == "Gross Photosynthesis")%>%filter(timepoint == 10), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3) # temp is highly significant

lsmeans(mod, pairwise~ wound + temp, adjust="tukey")

# DAY 19
mod<- lmer(umol.cm2.hr ~ wound*temp  + (1|genotype), data = rates.filtered%>%filter(rate == "Gross Photosynthesis")%>%filter(timepoint == 19), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3) #temp*wound interaction highly significant

lsmeans(mod, pairwise~wound*temp, adjust="tukey")
```

```{r}
ggplot(data = rates.filtered%>%filter(!timepoint == 0)%>%filter(rate == "Gross Photosynthesis"), aes(x = timepoint, y = umol.cm2.hr, col = temp))+
  #geom_jitter(size = .5, alpha = .5)+
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, position = position_dodge(width = 0.2)) +
  stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.2))+ 
  stat_summary( fun = "mean", geom = "line", size = 1, position = position_dodge(width = 0.2), linetype = "dashed", aes(group = temp))+
  ylab("O2 Rate (umol/cm2/hr)")+
  xlab("Day")+
  #facet_grid(rate~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")), scales = "free_y", space="free")+ 
  theme_classic()+
  labs(color = "Temperature")+
  scale_color_manual(values = c('blue', 'red'), labels = c("Ambient (~28)", "Elevated (~30)"))+
   theme_few(base_size = 20)

ggplot(data = rates.filtered%>%filter(!timepoint == 0)%>%filter(rate == "Gross Photosynthesis"), aes(x = timepoint, y = umol.cm2.hr))+
  #geom_jitter(size = .5, alpha = .5)+
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, position = position_dodge(width = 0.2)) +
  stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.2))+ 
  stat_summary( fun = "mean", geom = "line", size = 1, position = position_dodge(width = 0.2), linetype = "dashed")+
  ylab("O2 Rate (umol/cm2/hr)")+
  xlab("Day")+
  #facet_grid(rate~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")), scales = "free_y", space="free")+ 
  theme_classic()+
  #labs(color = "Temperature")+
  #scale_color_manual(values = c('blue', 'red'), labels = c("Ambient (~28)", "Elevated (~30)"))+
   theme_few(base_size = 20)
```



```{r}
rates.filtered%>%filter(rate == "Gross Photosynthesis")%>%filter(timepoint == 10) -> mean.grossphoto
mean.grossphoto%>%unite("treatment", wound:temp, remove = FALSE) -> mean.grossphoto

# Calculating mean growth for each group to plot.
mean.grossphoto%>%group_by(wound, temp)%>%summarize(Mean = mean(umol.cm2.hr, na.rm=TRUE),
                                                        'N'=length(umol.cm2.hr ))-> mean

mean.grossphoto%>%aggregate(umol.cm2.hr ~ treatment, FUN = function(x) sd(x)/sqrt(length(x)))%>%separate(col = treatment, into = c("wound", "temp")) -> std_err

# Assuming all contrasts were significant for simplicity; this may change based on actual results.
#mean$letter <- c("ac", "b", "a","abc", "abc", "b")

dodge<-position_dodge(width=0.6) # this offsets the points so they don't overlap

  ggplot(mean, aes(x = temp, y = Mean, colour = temp, fill = temp))+
  facet_wrap(~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")))+
  geom_bar(stat="identity", position="dodge") +geom_errorbar(aes(ymin=Mean-std_err$umol.cm2.hr, ymax=Mean+std_err$umol.cm2.hr), width=0, position=position_dodge(width = 0.9))+
   # geom_text(aes(label=letter), nudge_y=0.32, color="black")+
  geom_point(data = mean.grossphoto, aes(x = temp, y = umol.cm2.hr, colour = temp), shape = 21, alpha = 0.7,
             position = position_jitterdodge(jitter.width = 0.3, jitter.height=0.2, dodge.width=0.9))+
  scale_color_manual(values = c('black', 'black', 'black'), guide="none")+
  scale_fill_manual(values = c('white', 'grey'), labels=c("Ambient", "Elevated"), guide = "none")+
  scale_x_discrete(labels = c('Ambient','Elevated'))+
xlab(expression("Temperature"))+
  ylab(expression(paste("Gross Photosynthesis Rate (mg/cm2/day)")))+
  theme(axis.text.x=element_text(colour="black"))+
  theme(axis.text.y=element_text(colour="black"))+theme_few(base_size = 12)

print(figure_1)

#save plot

ggsave("figure_1.png", plot = figure_1, path = here(),
       width = 8,
       height = 8,
       units = "in")
```
summary values 
```{r}
summarySE(rates.filtered%>%filter(rate == "Gross Photosynthesis"), measurevar = "umol.cm2.hr", groupvars = c("timepoint", "temp", "wound"),
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE) -> GP.summary

summarySE(rates.filtered%>%filter(rate == "Gross Photosynthesis"), measurevar = "umol.cm2.hr", groupvars = c("timepoint", "temp"),
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE) -> GP.summary

summary(rates.filtered%>%filter(rate == "Gross Photosynthesis"))
```



# plotting all rates
```{r}
summarySE(data = rates.filtered, measurevar = "umol.cm2.hr", groupvars = c("rate","wound", "temp"),
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE)



rates.filtered%>%unite("treatment", c(wound, temp, timepoint), remove = FALSE) -> all.rates

# Calculating mean rate for each group to plot.
all.rates%>%group_by(rate, wound, temp, timepoint)%>%summarize(Mean = mean(umol.cm2.hr, na.rm=TRUE),
                                                        'N'=length(umol.cm2.hr ))-> mean

all.rates%>%aggregate(umol.cm2.hr ~ treatment, FUN = function(x) sd(x)/sqrt(length(x)))%>%separate(col = treatment, into = c("wound", "temp", "timepoint")) -> std_err

dodge<-position_dodge(width=0.6) # this offsets the points so they don't overlap

  ggplot(mean, aes(x = timepoint, y = Mean, colour = temp, fill = temp))+
  facet_grid(rate~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")))+
  geom_point(stat="identity", position="dodge")+
 scale_color_manual(values = c('blue', 'red'), labels=c("Ambient", "Elevated"))+
  #scale_fill_manual(values = c('blue', 'red'))+
  #scale_x_discrete(labels = c('Ambient','Elevated'))+
xlab(expression("Day"))+
  ylab(expression(paste("umol.cm2.hr")))+
   theme_few(base_size = 12)


```


```{r}
ggplot(data = rates.filtered%>%filter(!timepoint == 0), aes(x = timepoint, y = umol.cm2.hr, col = temp))+
  #geom_jitter(size = .5, alpha = .5)+
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, position = position_dodge(width = 0.2)) +
  stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.2))+ 
  stat_summary( fun = "mean", geom = "line", size = 1, position = position_dodge(width = 0.2), linetype = "dashed", aes(group = temp))+
  ylab("O2 Rate (umol/cm2/hr)")+
  xlab("Day")+
  facet_grid(rate~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")), scales = "free_y", space="free")+ 
  theme_classic()+
  labs(color = "Temperature")+
  scale_color_manual(values = c('blue', 'red'), labels = c("Ambient (~28)", "Elevated (~30)"))+
   theme_few(base_size = 15) -> all.rates
print(all.rates)

#save plot

ggsave("all.rates.png", plot = all.rates, path = here(),
       width = 10,
       height = 8,
       units = "in")
```


# PR ratios
```{r}
rates.filtered%>%filter(!rate == "Net Photosynthesis")%>%select(rate, coral_id, temp, wound, genotype, timepoint, umol.cm2.hr)%>%pivot_wider(names_from = rate, values_from = umol.cm2.hr)%>%clean_names() -> rates.wide

rates.wide%>%mutate(ratio = ((12*gross_photosynthesis)/(24*respiration))) ->ratio

#full model 
mod<- lmer(ratio ~ wound*temp*timepoint + (1|genotype/coral_id), data = ratio, contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3)

summary(ratio)
summarySE(ratio, measurevar = "ratio", groupvars = c("timepoint", "temp", "wound"),
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE) -> summary.ratio
ratio.filtered = subset(ratio%>%filter(timepoint == "10"), ratio > 1)
```

model at each timepoint
```{r}
# DAY 0
mod<- lmer(ratio ~ wound*temp  + (1|genotype), data = ratio%>%filter(timepoint == 0), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3)

# DAY 1
mod<- lmer(ratio ~ wound*temp  + (1|genotype), data = ratio%>%filter(timepoint == 1), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3)

# DAY 10
mod<- lmer(ratio ~ wound*temp  + (1|genotype), data = ratio%>%filter(timepoint == 10), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3)
lsmeans(mod, pairwise~wound*temp, adjust="tukey") # no significant pairwise comparisons

# DAY 19
mod<- lmer(ratio ~ wound*temp  + (1|genotype), data = ratio%>%filter(timepoint == 19), contrasts = list(wound = "contr.sum", temp = "contr.poly"))
anova(mod, type = 3)
lsmeans(mod, pairwise~wound*temp, adjust="tukey")
```


P:R plot
```{r}
ggplot(data = ratio%>%filter(!timepoint == 0), aes(x = timepoint, y = ratio, col = temp))+
  #geom_jitter(size = .5, alpha = .5)+
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, position = position_dodge(width = 0.2)) +
  stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.2))+ 
  stat_summary( fun = "mean", geom = "line", size = 1, position = position_dodge(width = 0.2), linetype = "dashed", aes(group = temp))+
  ylab("Daily P:R ratio")+
  xlab("Day")+
  facet_grid(~wound, labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")), scales = "free_y", space="free")+ 
  theme_classic()+
  labs(color = "Temperature")+
  scale_color_manual(values = c('blue', 'red'), labels = c("Ambient (~28)", "Elevated (~30)"))+
   theme_few(base_size = 20)
```



