---
title: "Metabolic rate analysis"
author: "Ninah Munk"
date: "2024-05-17"
output: html_document
---
# Load packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(emmeans)
library(lme4)
library(Matrix)
library(readxl)
library(lmerTest)
library(nlme)
library(car)
library(ggpubr)
library(rstatix)
library(Rmisc)
library(ggthemes)
library(effectsize)
library(here)

```


# Load and Clean Data
```{r}

read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Respiration/Output/full.data.cleaned.csv")-> data
read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Data/worm_damage.csv")-> worm
```

Clean up some data and combine data frames
```{r}
data <- rename(data, replace = c("coral_num" = "coral_id"))
data%>%select(-X) -> data

worm%>%select(coral_id, affected, percent_affected)-> worm.damage

data%>%left_join(worm.damage, by = "coral_id")-> data2

```

Change timepoint to a continuous variable and numeric 
```{r}

data2%>%mutate(timepoint = case_when(timepoint == "day0" ~ "0",
                                     timepoint == "day1" ~ "1",
                                     timepoint == "day10" ~ "10",
                                     timepoint == "day19" ~ "19"))%>%mutate(timepoint = as.numeric(timepoint),
                                                                            temp = as.factor(temp),
                                                                            wound = as.factor(wound),
                                                                            genotype = as.factor(genotype)) -> data2

```


```{r}
ggplot(data2%>%filter(timepoint == "0"), aes(x = percent_affected, y = umol.cm2.hr))+geom_point()+geom_smooth(method = "lm")+facet_grid(~rate)

```


# RESPIRATION 
Run Models: The full model includes fixed effects wound, temp, and timepoint. Percent_affected is included as a covariate. 
```{r}
#full model 
mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint*percent_affected + (1|genotype/coral_id), data = data2%>%filter(rate == "Respiration"))
anova(mod)

#reduced model
mod2<- lmer(umol.cm2.hr ~ wound*temp*timepoint + temp*timepoint*percent_affected + (1|genotype/coral_id), data = data2%>%filter(rate == "Respiration"))
anova(mod2)

```

Filter data to remove worm effect determined from growth data. this threshold is below 10 percent worm damage. check that percent affected is no longer significant in full model
```{r}
rates.filtered = subset(data2, percent_affected <= 10)

mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint*percent_affected + (1|genotype/coral_id), data = rates.filtered%>%filter(rate == "Respiration"))
anova(mod) #percent affected is not longer significant 

mod<- lmer(umol.cm2.hr ~ wound + temp*timepoint*percent_affected + (1|genotype/coral_id), data = rates.filtered%>%filter(rate == "Respiration"))
anova(mod)# percent affected not significant in reduced model
```


Run models with filtered data without percent affected as a covariate. There is a significant interaction between temperature and timepoint in all three models. This becomes more significant after filtering out initial timepoint, so this tells me this interaction is driven primarily by the following timepoints. 
```{r}

#full model 
mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint + (1|genotype/coral_id), data = rates.filtered%>%filter(rate == "Respiration"))
anova(mod)
summary(mod)
options(es.use_symbols = TRUE)
parameters::model_parameters(anova(mod))
eta_squared(mod)

```

Percent change in respiration from pre-treatment day 0 to post treatment day 1 
```{r}
resp.change <- rates.filtered%>%select(-drymass,-SA,-umol.hr)%>%pivot_wider(names_from = timepoint, values_from = umol.cm2.hr) %>% mutate(change1 = ((`1` - `0`)/`0`)*100,
                                                                                                                                          change2 = ((`10` - `0`)/`0`)*100,
                                                                                                                                          change3 = ((`19` - `0`)/`0`)*100)

```


## respiration effect sizes with models
```{r}
#full model
change1<- lmer(change1 ~ wound*temp + (1|genotype), data = resp.change%>%filter(rate == "Respiration"))

#reduced model 
change1<- lmer(change1 ~ wound + temp + (1|genotype), data = resp.change%>%filter(rate == "Respiration"))
summary(change1)
anova(change1)
library(broom.mixed)
library(forcats)


tidy_bm <- tidy(change1, conf.int = TRUE, effects = "fixed") %>%
  filter(term != "(Intercept)")

ggplot(tidy_bm, aes(y = term, xmin = conf.low, xmax = conf.high, x = estimate, color = p.value < 0.05)) +
  geom_pointrange() +  # Point range plot to show estimates and intervals
  scale_color_manual(values = c("red", "blue"), 
                     labels = c("Not Significant", "Significant"),
                     guide = guide_legend(title = "Significance")) +
  labs(title = "Estimates and Confidence Intervals for Model Terms",
       x = "Estimate and Confidence Interval",
       y = "Model Term") +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 10),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 10),
    legend.key=element_rect(fill="white"),
    strip.text = element_text(size = 10),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.text.x = element_text(size = 10),
    axis.ticks = element_blank(), 
    axis.title.x = element_text(size = 10),
    axis.line = element_line(color = "black") 
  )


```

```{r}
#full model change 2
change2<- lmer(change2 ~ wound*temp + (1|genotype), data = resp.change%>%filter(rate == "Respiration"))
summary(change2)
anova(change2)

#reduced model change 2
change2<- lmer(change2 ~ wound + temp + (1|genotype), data = resp.change%>%filter(rate == "Respiration"))
summary(change2)
anova(change2)

tidy_bm <- tidy(change2, conf.int = TRUE, effects = "fixed") %>%
  filter(term != "(Intercept)")

ggplot(tidy_bm, aes(y = term, xmin = conf.low, xmax = conf.high, x = estimate, color = p.value < 0.05)) +
  geom_pointrange() +  # Point range plot to show estimates and intervals
  scale_color_manual(values = c("red", "blue"), 
                     labels = c("Not Significant", "Significant"),
                     guide = guide_legend(title = "Significance")) +
  labs(title = "Estimates and Confidence Intervals for Model Terms",
       x = "Estimate and Confidence Interval",
       y = "Model Term") +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 10),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 10),
    legend.key=element_rect(fill="white"),
    strip.text = element_text(size = 10),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.text.x = element_text(size = 10),
    axis.ticks = element_blank(), 
    axis.title.x = element_text(size = 10),
    axis.line = element_line(color = "black") 
  )

```


```{r}
#full model change 3
change3<- lmer(change3 ~ wound*temp + (1|genotype), data = resp.change%>%filter(rate == "Respiration"))
summary(change3)
anova(change3)
lsmeans(change3, pairwise~wound*temp, adjust="tukey")

tidy_bm <- tidy(change3, conf.int = TRUE, effects = "fixed") %>%
  filter(term != "(Intercept)")

ggplot(tidy_bm, aes(y = term, xmin = conf.low, xmax = conf.high, x = estimate, color = p.value < 0.05)) +
  geom_pointrange() +  # Point range plot to show estimates and intervals
  scale_color_manual(values = c("red", "blue"), 
                     labels = c("Not Significant", "Significant"),
                     guide = guide_legend(title = "Significance")) +
  labs(title = "Estimates and Confidence Intervals for Model Terms",
       x = "Estimate and Confidence Interval",
       y = "Model Term") +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 10),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 10),
    legend.key=element_rect(fill="white"),
    strip.text = element_text(size = 10),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.text.x = element_text(size = 10),
    axis.ticks = element_blank(), 
    axis.title.x = element_text(size = 10),
    axis.line = element_line(color = "black") 
  )

```
## respiration - percent change
```{r}

resp.change%>% pivot_longer(cols = c("change1", "change2", "change3"),
               names_to = "timepoint",
               values_to = "percent.change") -> rate.change.long


summarySE(data = rate.change.long, measurevar = "percent.change", groupvars = c("rate", "temp", "timepoint","wound"),
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE) -> summary

summarySE(data = rate.change.long, measurevar = "percent.change", groupvars = c("rate", "temp", "timepoint"),
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE) -> summary2

dodge<-position_dodge(width=0.6) # this offsets the points so they don't overlap

ggplot(summary%>%filter(timepoint == "change1"), aes(x = temp, y = percent.change, colour = temp, fill = temp))+
  facet_wrap(rate~wound, scales = "free", labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")))+
  geom_bar(stat="identity", position="dodge") +geom_errorbar(aes(ymin=percent.change-se, ymax=percent.change+se), width=0, position=position_dodge(width = 0.9))+
  geom_point(data = rate.change.long%>%filter(timepoint == "change1"), aes(x = temp, y = percent.change, colour = temp), shape = 21, alpha = 0.7,
             position = position_jitterdodge(jitter.width = 0.3, jitter.height=0.2, dodge.width=0.9))+
  scale_color_manual(values = c('black', 'black', 'black'), guide="none")+
  scale_fill_manual(values = c('white', 'grey'), labels=c("Ambient", "Elevated"), guide = "none")+
  scale_x_discrete(labels = c('Ambient','Elevated'))+
xlab(expression("Temperature"))+
  ylab(expression(paste("Rate (mg/cm2/day)")))+
  theme(axis.text.x=element_text(colour="black"))+
  theme(axis.text.y=element_text(colour="black"))+theme_few(base_size = 12)

ggplot(summary, aes(x = timepoint, y = percent.change, fill = temp, colour =temp))+
  facet_wrap(~rate, scale = "free")+
  geom_bar(stat="identity", position= position_dodge(width = 0.9))+
  geom_errorbar(aes(ymin=percent.change-se, ymax=percent.change+se), width=0, position=position_dodge(width = 0.9))+
  geom_point(data = rate.change.long, aes(x = timepoint, y = percent.change, colour = temp, fill = temp), shape = 21, alpha = 0.5,position = position_jitterdodge(jitter.width = 0.3, jitter.height=0.2, dodge.width=0.9))+
  scale_color_manual(values = c('black', 'black'), guide="none")+
 scale_fill_manual(values = c('white', 'darkgrey'), guide = "none")+
  scale_x_discrete(labels = c('1','10', '19'))+
xlab(expression("Day"))+
  ylab(expression(paste("Percent Change in Rate")))+
  theme(axis.text.x=element_text(colour="black"))+
  theme(axis.text.y=element_text(colour="black"))+theme_few(base_size = 12)

rate.change.long$rate <- factor(rate.change.long$rate, levels = c("Respiration", "Gross Photosynthesis", "Net Photosynthesis"))
summary$rate <- factor(summary$rate, levels = c("Respiration", "Gross Photosynthesis", "Net Photosynthesis"))
rates.filtered$rate <- factor(rates.filtered$rate, levels = c("Respiration", "Gross Photosynthesis", "Net Photosynthesis"))

ggplot(data = rate.change.long, aes(x = timepoint,y = percent.change, fill = temp))+facet_grid(rate ~ wound, scales = "free", labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")))+
            stat_summary(geom = "col", fun = mean, width = 0.7,col = c("black"), stat = "identity", position= position_dodge(width = 0.9)) +
 # geom_point(data = rate.change.long, aes(x = timepoint, y = percent.change, fill = temp), shape = 21, alpha = 0.5,position = position_jitterdodge(jitter.width = 0.3, jitter.height=0.2, dodge.width=0.9))+
  geom_errorbar(data = summary, aes(x = timepoint, ymin=percent.change-se, ymax=percent.change+se), width=0, position=position_dodge(width = 0.9))+
  theme_light(base_size = 20) +
  scale_x_discrete(labels = c('1','10', '19'))+
  scale_fill_manual(values = c('white', 'darkgrey'))+
xlab(expression("Day"))+
  ylab(expression(paste("Mean Percent Change")))+
  theme(axis.text.x=element_text(colour="black"))+
  theme(axis.text.y=element_text(colour="black"))+theme_few(base_size = 12)

figure_1 <- ggplot(data = rates.filtered, aes(x = as.factor(timepoint),y = umol.cm2.hr, fill = temp))+facet_grid(rate~wound, scales = "free", labeller = labeller(wound = c("0" = "No Injury", "1" = "Fragmentation", "2" = "Abrasion")))+
            stat_summary(geom = "col", fun = mean, width = 0.7,col = c("black"), stat = "identity", position= position_dodge(width = 0.9)) +
  geom_point(data = rates.filtered, aes(x = as.factor(timepoint), y = umol.cm2.hr, fill = temp), shape = 21, alpha = 0.5,position = position_jitterdodge(jitter.width = 0.3, jitter.height=0.2, dodge.width=0.9))+
  #geom_errorbar(data = summary, aes(x = timepoint, ymin=percent.change-se, ymax=percent.change+se), width=0, position=position_dodge(width = 0.9))+
  theme_light(base_size = 40) +
  scale_x_discrete(labels = c('0','1','10', '19'))+
  scale_fill_manual(values = c('white', 'darkgrey'), guide = "none")+
xlab(expression("Day"))+
  ylab(expression(paste("O2 rate (umol/cm2/hr)")))+
  theme(axis.text.x=element_text(colour="black"))+
  theme(axis.text.y=element_text(colour="black"))+theme_few(base_size = 24)

ggsave("figure_3.png", plot = figure_1, path = here(),
       width = 8,
       height = 8,
       scale =2,
       units = "in")

```


```{r}
summary(rates.filtered%>%filter(rate == "Respiration", !timepoint == "0"))

summarySE(rates.filtered%>%filter(rate == "Respiration", !timepoint == "0"), measurevar = "umol.cm2.hr", groupvars = c("timepoint", "temp", "wound"),
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE) -> summary
```



## Respiration Figure with time as continuous variable. Acclimation to temperature? 
```{r}
ggplot(data = rates.filtered%>%filter(rate == "Respiration"), aes(x = timepoint, y = umol.cm2.hr, color = temp))+geom_point()+geom_smooth(method = "lm")

ggplot(data = rates.filtered%>%filter(rate == "Respiration", !timepoint == "0"), aes(x = timepoint, y = umol.cm2.hr, color = temp))+geom_point()+geom_smooth(method = "lm")+facet_wrap(~wound)

# double check that resp is the same at day 0 

```


# NET PHOTOSYNTHESIS 
Run Models: The full model includes fixed effects wound, temp, and timepoint. Percent_affected is included as a covariate. 
```{r}
#full model 
mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint + (1|genotype/coral_id), data = rates.filtered%>%filter(rate == "Net Photosynthesis"))
summary(mod)
Anova(mod)
anova(mod)


```

Summary Statistics
```{r}
summary(rates.filtered%>%filter(rate == "Net Photosynthesis")) 

summarySE(rates.filtered%>%filter(rate == "Net Photosynthesis"), measurevar = "umol.cm2.hr", groupvars = c("timepoint", "temp", "wound"),
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE) -> summary.net
```


Net Photosynthesis Figure. Wound, temperature, and time interact to increase Net photosynthesis ? 
```{r}
ggplot(data = rates.filtered%>%filter(rate == "Net Photosynthesis"), aes(x = timepoint, y = umol.cm2.hr, color = temp))+geom_point()+geom_smooth(method = "lm")

ggplot(data = rates.filtered%>%filter(rate == "Net Photosynthesis"), aes(x = timepoint, y = umol.cm2.hr, color = temp))+geom_point()+geom_smooth(method = "lm")+facet_wrap(~wound)

```

Percent change models
```{r}
#full model
change1<- lmer(change1 ~ wound*temp + (1|genotype), data = resp.change%>%filter(rate == "Net Photosynthesis"))

#reduced model 
change1<- lmer(change1 ~ wound + temp + (1|genotype), data = resp.change%>%filter(rate == "Net Photosynthesis"))
summary(change1)
anova(change1)



tidy_bm <- tidy(change1, conf.int = TRUE, effects = "fixed") %>%
  filter(term != "(Intercept)")

ggplot(tidy_bm, aes(y = term, xmin = conf.low, xmax = conf.high, x = estimate, color = p.value < 0.05)) +
  geom_pointrange() +  # Point range plot to show estimates and intervals
  scale_color_manual(values = c("red", "blue"), 
                     labels = c("Not Significant", "Significant"),
                     guide = guide_legend(title = "Significance")) +
  labs(title = "Estimates and Confidence Intervals for Model Terms",
       x = "Estimate and Confidence Interval",
       y = "Model Term") +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 10),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 10),
    legend.key=element_rect(fill="white"),
    strip.text = element_text(size = 10),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.text.x = element_text(size = 10),
    axis.ticks = element_blank(), 
    axis.title.x = element_text(size = 10),
    axis.line = element_line(color = "black") 
  )


```


```{r}
#full model change 2
change2<- lmer(change2 ~ wound*temp + (1|genotype), data = resp.change%>%filter(rate == "Net Photosynthesis"))
summary(change2)
anova(change2)

#reduced model change 2
change2<- lmer(change2 ~ wound + temp + (1|genotype), data = resp.change%>%filter(rate == "Net Photosynthesis"))
summary(change2)
anova(change2)

tidy_bm <- tidy(change2, conf.int = TRUE, effects = "fixed") %>%
  filter(term != "(Intercept)")

ggplot(tidy_bm, aes(y = term, xmin = conf.low, xmax = conf.high, x = estimate, color = p.value < 0.05)) +
  geom_pointrange() +  # Point range plot to show estimates and intervals
  scale_color_manual(values = c("red", "blue"), 
                     labels = c("Not Significant", "Significant"),
                     guide = guide_legend(title = "Significance")) +
  labs(title = "Estimates and Confidence Intervals for Model Terms",
       x = "Estimate and Confidence Interval",
       y = "Model Term") +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 10),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 10),
    legend.key=element_rect(fill="white"),
    strip.text = element_text(size = 10),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.text.x = element_text(size = 10),
    axis.ticks = element_blank(), 
    axis.title.x = element_text(size = 10),
    axis.line = element_line(color = "black") 
  )

```


```{r}
#full model change 3
change3<- lmer(change3 ~ wound*temp + (1|genotype), data = resp.change%>%filter(rate == "Net Photosynthesis"))

#reduced model
change3<- lmer(change3 ~ wound + temp + (1|genotype), data = resp.change%>%filter(rate == "Net Photosynthesis"))
summary(change3)
anova(change3)
lsmeans(change3, pairwise~wound*temp, adjust="tukey")

tidy_bm <- tidy(change3, conf.int = TRUE, effects = "fixed") %>%
  filter(term != "(Intercept)")

ggplot(tidy_bm, aes(y = term, xmin = conf.low, xmax = conf.high, x = estimate, color = p.value < 0.05)) +
  geom_pointrange() +  # Point range plot to show estimates and intervals
  scale_color_manual(values = c("red", "blue"), 
                     labels = c("Not Significant", "Significant"),
                     guide = guide_legend(title = "Significance")) +
  labs(title = "Estimates and Confidence Intervals for Model Terms",
       x = "Estimate and Confidence Interval",
       y = "Model Term") +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 10),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 10),
    legend.key=element_rect(fill="white"),
    strip.text = element_text(size = 10),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.text.x = element_text(size = 10),
    axis.ticks = element_blank(), 
    axis.title.x = element_text(size = 10),
    axis.line = element_line(color = "black") 
  )

```


# GROSS PHOTOSYNTHESIS 
Run Models: The full model includes fixed effects wound, temp, and timepoint. Percent_affected is included as a covariate. 
```{r}
#full model 
mod<- lmer(umol.cm2.hr ~ wound*temp*timepoint + (1|genotype/coral_id), data = rates.filtered%>%filter(rate == "Gross Photosynthesis"))
summary(mod)
anova(mod)

summary(rates.filtered%>%filter(rate == "Gross Photosynthesis"))
```

Gross Photosynthesis Figure 
```{r}
ggplot(data = rates.filtered%>%filter(rate == "Gross Photosynthesis"), aes(x = timepoint, y = umol.cm2.hr, color = temp))+geom_point()+geom_smooth(method = "lm")

ggplot(data = rates.filtered%>%filter(rate == "Gross Photosynthesis"), aes(x = timepoint, y = umol.cm2.hr, color = temp))+geom_point()+geom_smooth(method = "lm")+facet_wrap(~wound)

summarySE(rates.filtered%>%filter(rate == "Gross Photosynthesis"), measurevar = "umol.cm2.hr", groupvars = c("timepoint", "temp", "wound"),
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE) -> summary.gross

```

# All rates. 
```{r}
ggplot(data = rates.filtered%>%filter(!timepoint == 0), aes( x = timepoint, y = umol.cm2.hr, color = temp))+ geom_point(position = "jitter")+geom_smooth(method = "lm")+ facet_wrap(rate~wound) 

```

```{r}
rates.filtered%>%filter(!rate == "Net Photosynthesis")%>%select(rate, coral_id, temp, wound, genotype, timepoint, umol.cm2.hr)%>%pivot_wider(names_from = rate, values_from = umol.cm2.hr)%>%clean_names() -> rates.wide

rates.wide%>%mutate(ratio = ((12*gross_photosynthesis)/(24*respiration))) ->ratio

ggplot(data = ratio, aes( x = as.factor(timepoint), y = ratio, color = temp))+ geom_boxplot()+geom_point(position = "jitter", alpha = .5)+ facet_wrap(~wound) 

#full model 
mod<- lmer(ratio ~ wound*temp*timepoint + (1|genotype/coral_id), data = ratio)
summary(mod)
anova(mod)

summary(ratio)
summarySE(ratio, measurevar = "ratio", groupvars = c("timepoint", "temp", "wound"),
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE) -> summary.ratio
ratio.filtered = subset(ratio%>%filter(timepoint == "10"), ratio > 1)

```


# Getting summary statistics for all three rates. I'm interested in 1) the range of rates (minimum and maximum) and 2) 
```{r}
# RESPIRATION - temperature*time
summarySE(rates.filtered%>%filter(rate == "Respiration"), measurevar = "umol.cm2.hr", groupvars = c("timepoint", "temp"),
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE) -> summary.resp

summarySE(rate.change.long%>%filter(rate == "Respiration"), measurevar = "percent.change", groupvars = c("timepoint", "temp"),
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE) -> summary.resp.change

# NET PHOTOSYNTHESIS - time + temperature*time
summarySE(rates.filtered%>%filter(rate == "Net Photosynthesis"), measurevar = "umol.cm2.hr", groupvars = c("timepoint", "temp"), na.rm = FALSE, conf.interval = 0.95, .drop = TRUE) -> summary.net
# GROSS PHOTOSYNTHESIS - temperature + time
summarySE(rates.filtered%>%filter(rate == "Gross Photosynthesis"), measurevar = "umol.cm2.hr", groupvars = c("timepoint", "temp"),
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE) -> summary.gross
summarySE(rate.change.long%>%filter(rate == "Gross Photosynthesis"), measurevar = "percent.change", groupvars = c("timepoint", "temp"),
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE) -> summary.gross.change

# PERCENT CHANGE
summarySE(rate.change.long%>%filter(rate == "Net Photosynthesis"), measurevar = "percent.change", groupvars = c("timepoint", "temp", "wound"),
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE) -> summary.net.change
# P:R
summarySE(ratio, measurevar = "ratio", groupvars = c("timepoint", "temp", "wound"), na.rm = FALSE, conf.interval = 0.95, .drop = TRUE) -> summary.ratio
```

```{r}
# PERCENT CHANGE
summarySE(rate.change.long, measurevar = "percent.change", groupvars = c("rate", "timepoint", "temp", "wound"),
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE) -> summary.all.change


summary.all.change%>%mutate(timepoint = case_when(timepoint == "change1" ~ "1",
                                     timepoint == "change2" ~ "10",
                                     timepoint == "change3" ~ "19"))-> summary.all.change

ggplot(data = summary.all.change, aes(x = as.numeric(timepoint), y = percent.change, col = temp, shape = wound))+geom_point()+facet_wrap(rate~temp, scales = "free")
```






