---
title: "Acropora Regeneration Results"
author: "Ninah Munk"
date: "2024-02-14"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(janitor)
library(emmeans)
library(lme4)
library(Matrix)
library('readxl')
```

# Growth Rate Analysis
Read in weight data. 
```{r}
weight<- read.csv("Growth/Output/normalized_weight/initial.csv") %>% mutate(timepoint = "initial")%>% rename(weight = initial_g_cm2)
weight2<- read.csv("Growth/Output/normalized_weight/24hr.csv")%>% mutate(timepoint = "day1")%>% rename(weight = hr24_g_cm2)
weight3<- read.csv("Growth/Output/normalized_weight/day10.csv")%>% mutate(timepoint = "day10")%>% rename(weight = day10_g_cm2)
weight4<- read.csv("Growth/Output/normalized_weight/final.csv")%>% mutate(timepoint = "final")%>% rename(weight = final_g_cm2)
sample.info<- read.csv("Growth/Output/normalized_weight/samp_info.csv")

rbind(weight, weight2, weight3, weight4)%>%
  pivot_wider(names_from = timepoint,
              values_from = weight) -> weight.wide              

```

Binding data frames together, making it wide format, and cleaning the data (i.e., removing any corals which were compromised in some way during the experiment). Then I'm checking the relationship between initial weight and final weight.  
```{r}
rbind(weight, weight2, weight3, weight4)%>%
  pivot_wider(names_from = timepoint,
              values_from = weight) -> weight.wide  

corals_to_remove <- c(11, 5, 37, 104, 71, 17, 29)
weight.clean <- weight.wide[!(weight.wide$coral_id %in% corals_to_remove), ]

summary(weight.clean) #mean initial size is greater than mean final size --> suggests that corals did not regain size that they lost by end of study 
hist(weight.clean$initial - weight.clean$final) # initial - final weight would be negative if corals were larger at end of exp than at the start  

```
Calculating Growth Rate across the entire duration of the experiment. 
```{r}
growth.rate<- weight.clean%>%mutate(full.growth = ((final - day1)*1000)/19) #multiply by 1000 to convert from grams to mg, divide by 19 to get growth per day
# need to check coral #100, has much higher growth )rate then everything else 
growth.rate<- left_join(growth.rate, sample.info, by = "coral_id")
growth.rate$wound<- as.factor(growth.rate$wound)

ggplot(growth.rate, aes(x = wound, y = full.growth, col = temp))+
  geom_boxplot()+ggtitle("Growth Rate (over entire duration of exp)")

# checking to see if initial coral size was correlated with growth rate 
ggplot(growth.rate, aes (x=initial, y = full.growth))+ geom_point()
ggplot(growth.rate, aes (x=day1, y = full.growth))+ geom_point()

# checking how much weight corals lost from wounding treatments
growth.rate%>%filter(wound == 2) -> abrasion 
growth.rate%>%filter(wound == 1) -> fragment 

abrasion%>% mutate(weight.lost = initial - day1)%>%select(coral_id, weight.lost, wound) -> abrasion
fragment%>% mutate(weight.lost = initial - day1)%>%select(coral_id, weight.lost, wound) -> fragment
mean(abrasion$weight.lost)
mean(fragment$weight.lost)

rbind(abrasion, fragment)-> wound.weight.loss

ggplot(data = wound.weight.loss, aes( x = wound, y = weight.lost, col = wound))+ geom_boxplot()+geom_point()


```
Calculating Growth Rate between timepoints. Day 1 to Day 10 // Day 10 to Day 19
```{r}
growth.rate%>%mutate(growth.pt1 = ((day10 - day1)*1000)/9)%>%mutate(growth.pt2 = ((final - day10)*1000)/9) -> growth.rate

ggplot(growth.rate, aes (x=wound, y = growth.pt1, col = temp))+
  geom_boxplot()+
  ggtitle("Growth Rate (first 1/2 of exp)")

ggplot(growth.rate, aes (x=wound, y = growth.pt2, col = temp))+
  geom_boxplot()+
  ggtitle("Growth Rate (second 1/2 of exp)")

ggplot(growth.rate, aes (x=initial, y = growth.pt1))+ geom_point()
ggplot(growth.rate, aes (x=day1, y = growth.pt1))+ geom_point()

```
```{r}
summary(growth.rate)
#make data long again
growth.rate %>% 
  pivot_longer(cols = c(growth.pt1, growth.pt2),
               names_to = "exp_period",
               values_to = "rate")-> growth.rate.long
  # unite("treatment", wound:temp, remove = F) %>% 
  # left_join(weight, by = c("coral_num" = "coral_id")) %>% 
  # rename(initial = `0`) -> growth.rate.long

ggplot(data = growth.rate.long, aes(x = exp_period, y = rate, shape = wound))+ 
  stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.1))+ 
  stat_summary( fun = "mean", geom = "line", size = 1, position = position_dodge(width = 0.1), linetype = "dashed", aes(group = wound))+ 
  facet_wrap(~temp)+ 
  theme_classic()

```
# Surface Area Analysis
```{r}
SA.initial<- read.csv("Surface_Area/Output/initial_surface_areas.csv")%>%mutate(timepoint = "initial")
SA.postwound<- read.csv("Surface_Area/Output/post_wound_surface_areas.csv")%>%rename(SA = SA_post_wound)%>%mutate(timepoint = "postwound")
SA.final<- read.csv("Surface_Area/Output/final_surface_areas.csv")%>%select(coral_id, CSA_cm2)%>%rename(SA = CSA_cm2)%>%mutate(timepoint = "final")

rbind(SA.initial, SA.postwound, SA.final) -> SA.long

SA.long%>%
  pivot_wider(names_from = timepoint,
              values_from = SA) -> SA.wide  

summary(SA.wide)

SA.wide %>% 
  rename(initial_SA = initial,
         postwound_SA = postwound,
         final_SA = final)%>%
  left_join(weight.wide, by= "coral_id")%>%
  rename(initial_weight = initial,
         day1_weight = day1,
         day10_weight = day10, 
         final_weight = final)-> SAxWeight

#clean data
corals_to_remove <- c(11, 5, 37, 104, 71, 17, 29)
SAxWeight <- SAxWeight[!(SAxWeight$coral_id %in% corals_to_remove), ]

ggplot(data = SAxWeight, aes(x= initial_weight, y = final_weight))+geom_point()
ggplot(data = SAxWeight, aes(x= initial_weight, y = initial_SA))+geom_point()
ggplot(data = SAxWeight, aes(x= day1_weight, y = final_SA))+geom_point()


```


# Wax Dip SA Standard Curve
```{r}
#wax dipping calibration: calculate surface area of dowels 
calibration<- read.csv('Surface_Area/Data/20230712_wax_calibration.csv')%>%clean_names()%>%
  mutate(wax_weight_g = postwax_weight_g - prewax_weight_g)%>%
  mutate(cal_radius_mm = diameter_mm / 2)%>%
  mutate(cal_radius_cm = cal_radius_mm /10)%>%
  mutate(height_cm = height_mm / 10)%>%
  mutate(CSA_cm2= ((2*3.14*cal_radius_cm*height_cm) + 3.14*(cal_radius_cm)^2)) #curved surface area (CSA) = 2piRH + piR^2 (one area of circle for top of coral)

#calculate the curve coefficients for slope and intercept to apply as the standard
stnd.curve <- lm(CSA_cm2~wax_weight_g, data=calibration)
plot(CSA_cm2~wax_weight_g, data=calibration)
stnd.curve$coefficients
summary(stnd.curve)$r.squared

#bring in the datasheet with coral samples 
smpls<- read.csv("Surface_Area/Data/20230712_wax_weights.csv")%>%clean_names()%>%
#subtract postwax weight from prewax weight
mutate(wax_weight_g = postwax_weight_g - prewax_weight_g)
#Calculate surface area using the standard curve
smpls$SA <- stnd.curve$coefficients[2] * smpls$wax_weight_g + stnd.curve$coefficients[1]
```
```{r}
# wax weight X coral weight 
smpls%>%select(coral_id, wax_weight_g)%>%
  left_join(weight.wide, by = "coral_id")-> waxXweight

ggplot(data = waxXweight, aes(x= final, y = wax_weight_g))+geom_point()

stnd.curve <- lm(final~wax_weight_g, data=waxXweight)
plot(final~wax_weight_g, data=waxXweight)
stnd.curve$coefficients
summary(stnd.curve)$r.squared


bw_initial<- read_xlsx("Growth/Data/bouyantweight_initial.xlsx", sheet = 1)%>%select(coral_id, bouyantweight_g)%>%left_join(SAxWeight, by = "coral_id")
ggplot(data = bw_initial, aes(x = bouyantweight_g, y = final_SA ))+geom_point()
ggplot(data = bw_initial, aes(x = bouyantweight_g, y = initial_weight ))+geom_point()
```




Load Respiration and Photosynthesis Rates.
```{r}
data<- read.csv("Respiration/Output/full.data.cleaned.csv") %>% 
  left_join(weight, by = c("coral_num" = "coral_id"))
```


Filter data frame by Respiration rate and time points
```{r}

R.0 <- data%>% 
  filter(Rate%in% c("Respiration"))%>%filter(timepoint%in% c("0"))%>%
  mutate(wound = as.factor(wound), 
         temp = as.factor(temp),
         genotype = as.factor(genotype),
         date = as.factor(date))

ggplot(data = R.0, aes(x = wound, y = umol.cm2.hr, col= temp)) + geom_point(position = position_dodge(width = 0.2), size = .5)+
  stat_summary(
    fun = "mean", 
    geom = "point", 
    size = 5, 
    position = position_dodge(width = 0.2)
  ) +
  stat_summary(
    fun.data = "mean_cl_normal", 
    geom = "errorbar", 
    width = 0.2, 
    position = position_dodge(width = 0.2)
  ) +
  ggtitle("Day 0 (pre-wounding)") +
  ylab('Rate (umol.cm2.hr)') +
  xlab('Wound Type')


R.1 <- data%>% 
  filter(Rate%in% c("Respiration"))%>%filter(timepoint%in% c("1"))%>%
  mutate(wound = as.factor(wound), 
         temp = as.factor(temp),
         genotype = as.factor(genotype),
         date = as.factor(date))

ggplot(data = R.1, aes(x = wound, y = umol.cm2.hr, col= temp)) + geom_point(position = position_dodge(width = 0.2), size = .5)+
  stat_summary(
    fun = "mean", 
    geom = "point", 
    size = 5, 
    position = position_dodge(width = 0.2)
  ) +
  stat_summary(
    fun.data = "mean_cl_normal", 
    geom = "errorbar", 
    width = 0.2, 
    position = position_dodge(width = 0.2)
  ) +
  ggtitle("Day 1 (post-wounding)") +
  ylab('Rate (umol.cm2.hr)') +
  xlab('Wound Type')


R.10 <- data%>% 
  filter(Rate%in% c("Respiration"))%>%filter(timepoint%in% c("10"))%>%
  mutate(wound = as.factor(wound), 
         temp = as.factor(temp),
         genotype = as.factor(genotype),
         date = as.factor(date))

ggplot(data = R.10, aes(x = wound, y = umol.cm2.hr, col= temp)) + geom_point(position = position_dodge(width = 0.2), size = .5)+
  stat_summary(
    fun = "mean", 
    geom = "point", 
    size = 5, 
    position = position_dodge(width = 0.2)
  ) +
  stat_summary(
    fun.data = "mean_cl_normal", 
    geom = "errorbar", 
    width = 0.2, 
    position = position_dodge(width = 0.2)
  ) +
  ggtitle("Day 10 (post-wounding)") +
  ylab('Rate (umol.cm2.hr)') +
  xlab('Wound Type')


R.19 <- data%>% 
  filter(Rate%in% c("Respiration"))%>%filter(timepoint%in% c("19"))%>%
  mutate(wound = as.factor(wound), 
         temp = as.factor(temp),
         genotype = as.factor(genotype),
         date = as.factor(date))

ggplot(data = R.19, aes(x = wound, y = umol.cm2.hr, col= temp)) + geom_point(position = position_dodge(width = 0.2), size = .5)+
  stat_summary(
    fun = "mean", 
    geom = "point", 
    size = 5, 
    position = position_dodge(width = 0.2)
  ) +
  stat_summary(
    fun.data = "mean_cl_normal", 
    geom = "errorbar", 
    width = 0.2, 
    position = position_dodge(width = 0.2)
  ) +
  ggtitle("Day 19 (post-wounding)") +
  ylab('Rate (umol.cm2.hr)') +
  xlab('Wound Type')
```

# Julianna's probably dumb code

Get a wide data frame version of respiration rates. For each coral calculate change in rate at each time point from initial rate. 

```{r}
data %>% 
  filter(Rate == "Net Photosythesis") %>% 
  mutate(umol.cm2.hr = umol.cm2.hr * SA) %>% 
  select(-X, -date, -file.names.full, -chamber_vol, -SA) %>% 
  pivot_wider(names_from = timepoint,
              values_from = umol.cm2.hr) %>% 
  mutate(Change1.0 = `1` - `0`,
         Change10.0 = `10`-`0`,
         Change19.0 = `19`-`0` ) -> wideData

```

Now going to make this long again for plotting purposes

```{r}
wideData %>% 
  pivot_longer(cols = c(Change1.0, Change10.0, Change19.0),
               names_to = "Time_point",
               values_to = "ChangeFrom0_respir") %>% 
  unite("treatment", wound:temp, remove = F) %>% 
  left_join(weight, by = c("coral_num" = "coral_id")) %>% 
  rename(initial = `0`) -> freakylongData
```


Plot

```{r}
freakylongData %>% 
  ggplot(aes(x = treatment, y = ChangeFrom0_respir)) + 
  geom_boxplot() + 
  geom_point(position = position_dodge(width = 0.2), size = .5) +
  facet_wrap(~Time_point) +
  geom_hline(yintercept = 0, linetype = "dashed") 

ggplot(freakylongData %>% filter(Time_point == "Change19.0"), 
       aes(x = initial, y = ChangeFrom0_respir, col = as.factor(wound))) +
  geom_point()+
  geom_smooth(method = "lm")+
  facet_wrap(~temp)
```
Can we make a dumb model

```{r}
summary(lm(ChangeFrom0_respir ~ treatment*initial_g_cm2.y, data = freakylongData %>% filter(Time_point == ("Change19.0"))))

summary(lm(ChangeFrom0_respir ~ treatment*initial, data = freakylongData %>% filter(Time_point == "Change19.0")))

summary(lm(ChangeFrom0_respir~initial, 
           data = freakylongData %>% filter(Time_point == "Change19.0" & treatment == "2_H")))

m1 <- lme(Change1.0 ~ as.factor(wound) + temp + as.factor(wound):temp, random = ~1|genotype, data = wideData)
  summary(m1)
  
m2 <- lmer(Change1.0 ~ as.factor(wound) + temp + (1|genotype), data = wideData)
  summary(m2)
  
AIC(m1, m2) # only thing that matters is which one is lower and to be actually lower, needs to be a difference of 2. Can see model 2 is better 
```

```{r}
wideData %>% 
  rename(T0 = `0`,
         T1 = `1`,
         T10 = `10`,
         T19 = `19`) -> wideData

m1 <- lme(T1 ~ as.factor(wound) + temp + as.factor(wound):temp + T0, random = ~1|genotype, data = wideData)
  summary(m1)

wideData %>% 
  ggplot(aes(x = T1, y = T19, color = as.factor(wound))) +
  geom_point() 


freakylongData %>% 
  ggplot(aes(x = as.factor(wound), y = ChangeFrom0_respir, color = as.factor(temp))) +
  geom_boxplot() 
  
```
```{r}
m3 <- glm(Change10.0 ~ as.factor(wound) + temp + as.factor(wound):temp, data = wideData)
  summary(m3)
  
m4 <- glm(Change10.0 ~ as.factor(wound) + temp, data = wideData)
  summary(m4)
  
AIC(m3, m4) # only thing that matters is which one is lower and to be actually lower, needs to be a difference of 2. Can see model 4 is better 
```
```{r}
m5 <- glm(Change19.0 ~ as.factor(wound) + temp + as.factor(wound):temp, data = wideData)
  summary(m5)
  
m6 <- glm(Change19.0 ~ as.factor(wound) + temp, data = wideData)
  summary(m6)
  
AIC(m5, m6) # only thing that matters is which one is lower and to be actually lower, needs to be a difference of 2. Can see model 4 is better 
```

# Back to Ninah's dope code


```{r}
control_mean <- data %>% 
  filter(Rate == "Respiration", temp == "A", wound == 0) %>% 
  group_by(timepoint) %>% 
  reframe(respiration = mean(umol.cm2.hr))

difference_from_control <- data %>% filter(Rate == "Respiration")%>%
  mutate(treatment = paste(temp, wound, sep = "_")) %>% 
  filter(treatment != "A_0") %>% 
  left_join(control_mean, by = "timepoint") %>% 
  mutate(resp_difference = umol.cm2.hr - respiration)


difference_from_control %>% 
  ggplot(aes(x = as.factor(wound), y = resp_difference, col= temp)) + 
  geom_boxplot() + 
  geom_point(position = position_dodge(width = 0.2), size = .5) +
  facet_wrap(.~timepoint) +
  geom_hline(yintercept = 0, linetype = "dashed")
```









Analysis of initial (pre-wound) time point 
```{r}
#check distribution of data 
ggplot(data = R.0, aes(x=(umol.cm2.hr)))+geom_histogram()
ggplot(data = R.0, aes(x=log(umol.cm2.hr)))+geom_histogram()
# log transform umol.cm2.hr to fit assumptions of normality
R.0$log_umol <- log(R.0$umol.cm2.hr)

Rmodel <- lmer(umol.cm2.hr ~ wound + temp + (1|date), data = R.0) 
Rmodel <- lm(umol.cm2.hr ~ wound + temp , data = R.0)

summary(Rmodel)
# checking distribution of residuals
residuals <- resid(Rmodel)
hist(residuals)
qqnorm(residuals)
#qqline(residuals)
shapiro.test(residuals) #p = .83
plot(Rmodel)

# Visual inspection of heteroscedasticity
plot(fitted(Rmodel), resid(Rmodel), xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs. Fitted")+abline(h = 0, col = "red", lty = 2)

# Alternatively, you can use the ggplot2 package for a smoother plot
ggplot(data.frame(fitted = fitted(Rmodel), residuals = resid(Rmodel)), aes(x = fitted, y = residuals)) +
  geom_point() +
  geom_smooth(se = FALSE, method = "loess") +
  labs(x = "Fitted Values", y = "Residuals", title = "Residuals vs. Fitted")

# run Chi-Squared Type III ANOVA
anovaIII.results<-aov(Rmodel, type="III")
print(anovaIII.results)
# Obtain  estimated marginal means (EMMs) for all levels of the predictors
emmeans_results <- emmeans(Rmodel, ~ wound + temp)
# Display the EMMs
print(emmeans_results)
# Pairwise comparisons for the levels within each predictor
pairwise_results <- pairs(emmeans_results)
# Display the pairwise comparisons
print(pairwise_results)

```

Analysis of Day 1 (24 hrs post-wounding) time point
```{r}
#check distribution of data 
ggplot(data = R.1, aes(x=(umol.cm2.hr)))+geom_histogram()
ggplot(data = R.1, aes(x=log(umol.cm2.hr)))+geom_histogram()
# log transform umol.cm2.hr to fit assumptions of normality
R.1$log_umol <- log(R.1$umol.cm2.hr)

#Rmodel <- lmer(log_umol ~ wound + temp , data = R.0) 
R.1model <- lm(log_umol ~ wound + temp , data = R.1)

summary(R.1model)
# checking distribution of residuals
residuals <- resid(R.1model)
hist(residuals)
qqnorm(residuals)
#qqline(residuals)
shapiro.test(residuals) #p = .1715

# Visual inspection of heteroscedasticity
plot(fitted(R.1model), resid(R.1model), xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs. Fitted")+abline(h = 0, col = "red", lty = 2)

# Alternatively, you can use the ggplot2 package for a smoother plot
ggplot(data.frame(fitted = fitted(R.1model), residuals = resid(R.1model)), aes(x = fitted, y = residuals)) +
  geom_point() +
  geom_smooth(se = FALSE, method = "loess") +
  labs(x = "Fitted Values", y = "Residuals", title = "Residuals vs. Fitted")

# run Chi-Squared Type III ANOVA
anovaIII.results<-aov(R.1model, type="III")
print(anovaIII.results)
# Obtain  estimated marginal means (EMMs) for all levels of the predictors
emmeans_results <- emmeans(R.1model, ~ wound + temp)
# Display the EMMs
print(emmeans_results)
# Pairwise comparisons for the levels within each predictor
pairwise_results <- pairs(emmeans_results)
# Display the pairwise comparisons
print(pairwise_results)
```


```{r}
P.0 <- data%>% 
  filter(Rate%in% c("Gross Photosynthesis"))%>%filter(timepoint%in% c("0"))%>%
  mutate(wound = as.factor(wound), 
         temp = as.factor(temp),
         genotype = as.factor(genotype),
         date = as.factor(date))

ggplot(data = P.0, aes(x = wound, y = umol.cm2.hr, col= temp)) + geom_point(position = position_dodge(width = 0.2), size = .5)+
  stat_summary(
    fun = "mean", 
    geom = "point", 
    size = 5, 
    position = position_dodge(width = 0.2)
  ) +
  stat_summary(
    fun.data = "mean_cl_normal", 
    geom = "errorbar", 
    width = 0.2, 
    position = position_dodge(width = 0.2)
  ) +
  ggtitle("Day 0 (pre-wounding)") +
  ylab('Rate (umol.cm2.hr)') +
  xlab('Wound Type')


P.1 <- data%>% 
  filter(Rate%in% c("Gross Photosynthesis"))%>%filter(timepoint%in% c("1"))%>%
  mutate(wound = as.factor(wound), 
         temp = as.factor(temp),
         genotype = as.factor(genotype),
         date = as.factor(date))

ggplot(data = P.1, aes(x = wound, y = umol.cm2.hr, col= temp)) + geom_point(position = position_dodge(width = 0.2), size = .5)+
  stat_summary(
    fun = "mean", 
    geom = "point", 
    size = 5, 
    position = position_dodge(width = 0.2)
  ) +
  stat_summary(
    fun.data = "mean_cl_normal", 
    geom = "errorbar", 
    width = 0.2, 
    position = position_dodge(width = 0.2)
  ) +
  ggtitle("Day 1 (post-wounding)") +
  ylab('Rate (umol.cm2.hr)') +
  xlab('Wound Type')


P.10 <- data%>% 
  filter(Rate%in% c("Gross Photosynthesis"))%>%filter(timepoint%in% c("10"))%>%
  mutate(wound = as.factor(wound), 
         temp = as.factor(temp),
         genotype = as.factor(genotype),
         date = as.factor(date))

ggplot(data = P.10, aes(x = wound, y = umol.cm2.hr, col= temp)) + geom_point(position = position_dodge(width = 0.2), size = .5)+
  stat_summary(
    fun = "mean", 
    geom = "point", 
    size = 5, 
    position = position_dodge(width = 0.2)
  ) +
  stat_summary(
    fun.data = "mean_cl_normal", 
    geom = "errorbar", 
    width = 0.2, 
    position = position_dodge(width = 0.2)
  ) +
  ggtitle("Day 10 (post-wounding)") +
  ylab('Rate (umol.cm2.hr)') +
  xlab('Wound Type')


P.19 <- data%>% 
  filter(Rate%in% c("Gross Photosynthesis"))%>%filter(timepoint%in% c("19"))%>%
  mutate(wound = as.factor(wound), 
         temp = as.factor(temp),
         genotype = as.factor(genotype),
         date = as.factor(date))

ggplot(data = P.19, aes(x = wound, y = umol.cm2.hr, col= temp)) + geom_point(position = position_dodge(width = 0.2), size = .5)+
  stat_summary(
    fun = "mean", 
    geom = "point", 
    size = 5, 
    position = position_dodge(width = 0.2)
  ) +
  stat_summary(
    fun.data = "mean_cl_normal", 
    geom = "errorbar", 
    width = 0.2, 
    position = position_dodge(width = 0.2)
  ) +
  ggtitle("Day 19 (post-wounding)") +
  ylab('Rate (umol.cm2.hr)') +
  xlab('Wound Type')
```





```{r}
ggplot(full.data, aes(x = timepoint, y = fv_fm, shape = wound, col = temp))+
  geom_point(size = .5, position = position_dodge(width = 0.5))+
  stat_summary(fun = mean, geom = "point", size = 5 , position = position_dodge(width = 0.5))+
  stat_summary(
    fun.data = "mean_cl_normal", 
    geom = "errorbar", 
    width = 0.2, 
    position = position_dodge(width = 0.5)
  )+
  ggtitle("ACR Photosynthetic Efficiency")+
  ylab('Photosynthetic Efficiency (Fv/Fm)') +
  xlab('Days')
```




