---
title: "ACR_Regen_Growth"
author: "Ninah Munk"
date: "2024-02-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("readxl")
library("janitor")
library("ggplot2")
library('readxl')
library("dplyr")
library("emmeans")
library("lme4")
library("Matrix")
library("nlme")
```

# Calculating Skeletal Mass from Buoyant Weight 

Read in Raw Data. Combine into one long data frame.
```{r}
weight_initial<- read_xlsx("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Growth/Data/bouyantweight_initial.xlsx", sheet= "raw_data")%>%clean_names()%>%select(-calibration)%>%mutate(timepoint = "initial")
weight_24hr<- read_xlsx("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Growth/Data/bouyantweight_24hr.xlsx", sheet= "raw_data")%>%clean_names()%>%select(-bouyantweight_redo_if_applicable)%>%mutate(timepoint = "postwound")
weight_day10<- read_xlsx("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Growth/Data/bouyantweight_day10.xlsx", sheet= "raw_data")%>%clean_names()%>%select(-bouyantweight_redo_if_applicable)%>%mutate(timepoint = "day10")
weight_final<- read_xlsx("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Growth/Data/bouyantweight_final.xlsx", sheet= "raw_data")%>%clean_names()%>%mutate(timepoint = "final")

rbind(weight_initial, weight_24hr, weight_day10, weight_final)%>%select(-date, -group, -temp_c, -salinity_ppt, -salt_temp_c ) -> weight.long
```


```{r}
weight.long%>% 
  mutate(density_stopper = (air_weight_g * 0.9965)/(air_weight_g - fresh_weight_g))%>%
  mutate(density_sw = (air_weight_g - salt_weight_g)/ (air_weight_g / density_stopper))%>%
  mutate(dry_mass = bouyantweight_g / (1 - (density_sw/density_aragonite)))%>%
  mutate(coral_vol = (dry_mass / (density_aragonite - density_sw))) -> skeletal.mass

#clean data
corals_to_remove <- c(11, 5, 37, 104, 71, 17, 29)
skeletal.mass.clean <- skeletal.mass[!(skeletal.mass$coral_id %in% corals_to_remove), ]
  
# coral dry mass and volume should be positively correlated with buoyant weight
ggplot( data = skeletal.mass.clean, aes(x= bouyantweight_g, y = dry_mass))+ geom_point() + ggtitle("Bouyant Weight X Dry Mass")
ggplot( data = skeletal.mass.clean, aes(x= bouyantweight_g, y = coral_vol))+ geom_point()+ ggtitle("Bouyant Weight X Coral Volume")
```
 
# Calculating Coral Surface Area 
 
 ## Geometric coral surface areas
```{r}
# read in SA and calculate SA of a curved cylinder of each branch for every coral frag
data<- read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Surface_Area/Data/geometric_SA_initial.csv")%>%clean_names()%>%
  mutate(branch_height_cm = branch_height_mm / 10) %>%
  mutate(avg_diameter_mm= (diameter_base_mm + diameter_tip_mm) / 2)%>%
  mutate(radius_cm = (avg_diameter_mm/2)/10)%>%
  mutate(radius_tip_cm = (diameter_tip_mm/2)/10)%>%
  mutate(CSA_cm2= (2*3.14*(radius_cm*branch_height_cm) + 3.14*(radius_tip_cm)^2))

# add surface areas of branches together 
  summarized_initial_data <- data %>%
  group_by(coral_id) %>%
  summarize(SA = sum(CSA_cm2))
  
# calculate the mean apical tip diameter to use for removing SA of abrasion wound (type 2) from initial surface areas
wound_2_avg_tip_diameter<- mean(data$diameter_tip_mm)

# read in data sheet with height, tip, & base measurements for fragment wound (type 1)
wound_amount<-read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Surface_Area/Data/wound_type_1_measurements.csv")%>%clean_names()
# filling in height for wound type 2 (10 mm ~ 1cm)
wound_amount$branch_h_mm <- ifelse(wound_amount$wound == 2 & wound_amount$coral_id %in% unique(wound_amount$coral_id[wound_amount$wound == 2]), 10, wound_amount$branch_h_mm)
#filling in tip diameter for wound type 2 based on average diameter of apical branch tips from initial geometric measurements (see above)
wound_amount$tip_d_mm <- ifelse(wound_amount$wound == 2 & wound_amount$coral_id %in% unique(wound_amount$coral_id[wound_amount$wound == 2]), wound_2_avg_tip_diameter, wound_amount$tip_d_mm)
# make all other NAs in the data frame zero 
wound_amount[is.na(wound_amount)] <- 0

# calculate SA of removed coral tissue from wounding treatments
data2<- wound_amount%>%
  mutate(branch_height_cm = branch_h_mm / 10) %>%
  mutate(avg_diameter_mm= (base_d_mm + tip_d_mm) / 2)%>%
  mutate(radius_cm = (avg_diameter_mm/2)/10)%>%
  mutate(radius_tip_cm = (tip_d_mm/2)/10)%>%
  mutate(CSA_cm2= (2*3.14*(radius_cm*branch_height_cm) + 3.14*(radius_tip_cm)^2))

# new data frame with SA to subtract from initial SA
SA_to_subtract<-data2%>%select(coral_id, wound, CSA_cm2)
# join data frame with data frame of initial SA's
initialSA_woundSA<- left_join(summarized_initial_data, SA_to_subtract, by= 'coral_id')
# subtract removed SA from initial SA to get post wound SA
SA_postwound<-initialSA_woundSA%>%mutate(SA_postwound = SA - CSA_cm2)
```

try making SA removed equal for all fragmented corals
```{r}
#mean(SA_to_subtract$CSA_cm2[SA_to_subtract$wound ==1])
#SA_to_subtract$CSA_cm2[SA_to_subtract$wound == 1] <- 1.07
```


 ## Wax dipping method calibration curve
```{r}
#wax dipping calibration: calculate surface area of dowels 
calibration<- read.csv('/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Surface_Area/Data/20230712_wax_calibration.csv')%>%clean_names()%>%
  mutate(wax_weight_g = postwax_weight_g - prewax_weight_g)%>%
  mutate(cal_radius_mm = diameter_mm / 2)%>%
  mutate(cal_radius_cm = cal_radius_mm /10)%>%
  mutate(height_cm = height_mm / 10)%>%
  mutate(CSA_cm2= ((2*3.14*cal_radius_cm*height_cm) + 3.14*(cal_radius_cm)^2)) #curved surface area (CSA) = 2piRH + piR^2 (one area of circle for top of coral)

#calculate the curve coefficients for slope and intercept to apply as the standard
stnd.curve <- lm(CSA_cm2~wax_weight_g, data=calibration)
plot(CSA_cm2~wax_weight_g, data=calibration)
stnd.curve$coefficients
summary(stnd.curve)$r.squared

#bring in the datasheet with coral samples 
smpls<- read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Surface_Area/Data/20230712_wax_weights.csv")%>%clean_names()%>%
#subtract postwax weight from prewax weight
mutate(wax_weight_g = postwax_weight_g - prewax_weight_g)
#Calculate surface area using the standard curve
smpls$SA <- stnd.curve$coefficients[2] * smpls$wax_weight_g + stnd.curve$coefficients[1]

# Can we create another standard curve using final wax weights and final coral skeletal mass
skeletal.mass%>%filter(timepoint == "final")%>%select(coral_id, dry_mass)-> final.mass

# combine wax weight and coral weight 
smpls%>%select(coral_id, wax_weight_g)%>%
  left_join(final.mass, by = "coral_id")-> waxXmass

stnd.curve1 <- lm(dry_mass~wax_weight_g, data=waxXmass)
plot(dry_mass~wax_weight_g, data=waxXmass)
stnd.curve1$coefficients
summary(stnd.curve1)$r.squared #decent rsquared

#bring in sample info to filter out controls
sample.info<- read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Respiration/Data/samp_info.csv")

# calculate mass of removed coral skeleton
skeletal.mass%>%select(coral_id, timepoint, dry_mass)%>%pivot_wider(names_from = timepoint,
              values_from = dry_mass)%>%mutate(removed.mass = initial - postwound)%>%left_join(sample.info, by = "coral_id")%>%filter(!wound == 0) -> removed.mass

# lets check that final SA is positively correlated to final weight 
smpls%>%left_join(final.mass, by = "coral_id")-> final.SAXmass
ggplot(data = final.SAXmass, aes(x = dry_mass, y = SA))+ geom_point()

stnd.curve2 <- lm(dry_mass~SA, data=final.SAXmass)
plot(dry_mass~SA, data=final.SAXmass)
stnd.curve2$coefficients
summary(stnd.curve2)$r.squared #decent rsquared

```
 
 SA's derived geometrically are positively correlated to wax dip derived SA
```{r}
SA_postwound%>%select(coral_id, SA, SA_postwound)%>%rename(SA_initial = SA)%>%left_join(smpls%>%select(coral_id, SA)%>%rename(SA_final = SA))%>%left_join(sample.info)-> all.SA

ggplot(data = all.SA, aes(x= SA_initial, y = SA_final, col = temp))+geom_point()

ggplot(data = all.SA, aes(x= SA_postwound, y = SA_final, col = temp))+geom_point()+
  stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x)

stnd.curve3 <- lm(SA_initial~SA_final, data=all.SA)
plot(SA_initial~SA_final, data=all.SA)
stnd.curve3$coefficients
summary(stnd.curve3)$r.squared

```
 
Corals weight more at day 1 post wounding then they do at initial weighing (i.e., Corals grew after initial weighing so subtracting post wound weight from initial weight to get weight of removed mass wont work). This makes sense because corals had high growth rates before exp started (see below growth rates of unwounded corals pre start of experiment). Lets standardize all postwound, day 10, and final measurements by final SA derived by wax dip method, but lets subtract geometrically calculated SA of wounds from it for standardizing the postwound(day1) measurements. 

```{r}
smpls%>%left_join(SA_to_subtract, by = "coral_id")%>%mutate(postwound.SA = SA - CSA_cm2)%>%select(coral_id, postwound.SA) -> postwound.SA
```

Combine Initial, Postwound, and Final Surface Areas into one dataframe
```{r}
summarized_initial_data%>%mutate(timepoint = "initial") -> SA.1
postwound.SA%>%rename(SA = postwound.SA)%>%mutate(timepoint = "postwound") -> SA.2
smpls%>%select(coral_id, SA)%>%mutate(timepoint = "day10") -> SA.3
smpls%>%select(coral_id, SA)%>%mutate(timepoint = "final") -> SA.4


rbind(SA.1, SA.2, SA.3, SA.4) -> SA.long
#clean data
corals_to_remove <- c(11, 5, 37, 104, 71, 17, 29, 100) #100 is an outlier it has double the growth rate of the next fastest growing coral 
SA.long.clean <- SA.long[!(SA.long$coral_id %in% corals_to_remove), ]

```


# Growth Rates 

## Calculate Growth Rates not normalized by SA
```{r}
skeletal.mass.clean%>%select(coral_id, timepoint, dry_mass)%>%pivot_wider(names_from = timepoint,
              values_from = dry_mass)%>%
  rename(initial_weight = initial,
         day1_weight = postwound,
         day10_weight = day10, 
         final_weight = final) -> skeletal.mass.clean.wide

SA.long.clean%>%pivot_wider(names_from = timepoint,
              values_from = SA)%>%
  rename(initial.SA = initial,
         day1.SA = postwound,
         day10.SA = day10, 
         final.SA = final) -> SA.clean.wide

SA.clean.wide%>%left_join(skeletal.mass.clean.wide, by = "coral_id") -> mass.SA

mass.SA%>%mutate(full.growth = (((final_weight - day1_weight)/19)*1000),
                 pt1.growth = (((day10_weight - day1_weight)/9)*1000),
                 pt2.growth = (((final_weight - day10_weight)/9)*1000))-> growth
growth.not.normal<- growth%>%left_join(sample.info, by = "coral_id")

growth.not.normal$wound<- as.factor(growth.not.normal$wound)

ggplot(growth.not.normal, aes(x = wound, y = full.growth, col = temp))+
  geom_boxplot()+ggtitle("Growth Rate over entire duration of exp (not normalized by SA)")

# checking to see if initial coral size was correlated with growth rate 
ggplot(growth.not.normal, aes (x=initial_weight, y = full.growth))+ geom_point()
ggplot(growth.not.normal, aes (x=day1_weight, y = full.growth))+ geom_point()

ggplot(data = growth.not.normal, aes(x = day1.SA, y = full.growth, col = temp))+
  geom_point()+
  geom_smooth(method='lm')+
  facet_wrap(~wound)+
  ggtitle("Growth (mg) ~ Initial size")

ggplot(data = growth.not.normal, aes(x = day1.SA, y = full.growth, col = wound))+
  geom_point()+
  geom_smooth(method='lm')+
  #facet_wrap(~temp)+
  ggtitle("Growth (mg) ~ Initial size")

summary(growth.not.normal)
#make data long again
growth.not.normal %>% 
  pivot_longer(cols = c(pt1.growth, pt2.growth),
               names_to = "exp_period",
               values_to = "growthrate")-> growth.not.normal.long

ggplot(data = growth.not.normal.long, aes(x = exp_period, y = growthrate, shape = wound))+ 
  stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.1))+ 
  stat_summary( fun = "mean", geom = "line", size = 1, position = position_dodge(width = 0.1), linetype = "dashed", aes(group = wound))+ 
  facet_wrap(~temp)+ 
  theme_classic()+
  ggtitle("not standardized growth = mg/day")

ggplot(data = growth.not.normal.long, aes(x = day1.SA, y = growthrate, col = wound))+
  geom_point()+
  geom_smooth(method='lm')+
  facet_wrap(~exp_period + temp)+
  ggtitle("Growth (mg) ~ Initial size")



```


```{r}
model<- aov(full.growth ~ wound + as.factor(temp), data = growth.not.normal)
summary(model)

model2<- glm(full.growth ~ wound + as.factor(temp), data = growth.not.normal)
summary(model2)

model3<- lm(full.growth ~ wound + as.factor(temp) , data = growth.not.normal)
summary(model3)

emmeans_results <- emmeans(model2, ~ wound + temp)
# Display the EMMs
print(emmeans_results)
# Pairwise comparisons for the levels within each predictor
pairwise_results <- pairs(emmeans_results)
# Display the pairwise comparisons
print(pairwise_results)

```



## Normalize growth by timepoint specific SA
```{r}
growth%>%mutate(full.growth.rate = (full.growth / final.SA),
                pt1.growth.rate = (pt1.growth / day1.SA),
                pt2.growth.rate = (pt2.growth / final.SA))%>%left_join(sample.info, by = "coral_id") -> growth.normal
growth.normal$wound<- as.factor(growth.normal$wound)

ggplot(growth.normal, aes(x = wound, y = full.growth.rate, col = temp))+
  geom_boxplot()+ggtitle("Growth Rate over entire duration of exp (normalized by timepoint specific SA")

# checking to see if initial coral size was correlated with growth rate 
ggplot(growth.normal, aes (x=initial_weight, y = full.growth.rate))+ geom_point()
ggplot(growth.normal, aes (x=day1_weight, y = full.growth.rate))+ geom_point()

summary(growth.normal)
#make data long again
growth.normal %>% 
  pivot_longer(cols = c(pt1.growth.rate, pt2.growth.rate),
               names_to = "exp_period",
               values_to = "growthrate")-> growth.normal.long

ggplot(data = growth.normal.long, aes(x = exp_period, y = growthrate, shape = wound))+ 
  stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.1))+ 
  stat_summary( fun = "mean", geom = "line", size = 1, position = position_dodge(width = 0.1), linetype = "dashed", aes(group = wound))+ 
  facet_wrap(~temp)+ 
  theme_classic()+
  ggtitle("Growth Rate normalized to timepoint specific SA (mg/cm2/day)")


```

## Growth Rate standardized by post wound SA 
```{r}
growth%>%mutate(full.growth.rate = (full.growth / day1.SA),
                pt1.growth.rate = (pt1.growth / day1.SA),
                pt2.growth.rate = (pt2.growth / day1.SA))%>%left_join(sample.info, by = "coral_id") -> growth1.normal
growth1.normal$wound<- as.factor(growth1.normal$wound)

ggplot(growth1.normal, aes(x = wound, y = full.growth.rate, col = temp))+
  geom_boxplot()+ggtitle("Growth Rate over entire duration of exp (normalized by post wound SA")

# checking to see if initial coral size was correlated with growth rate 
ggplot(growth1.normal, aes (x=initial_weight, y = full.growth.rate))+ geom_point()
ggplot(growth1.normal, aes (x=day1_weight, y = full.growth.rate))+ geom_point()

summary(growth1.normal)
#make data long again
growth1.normal %>% 
  pivot_longer(cols = c(pt1.growth.rate, pt2.growth.rate),
               names_to = "exp_period",
               values_to = "growthrate")-> growth1.normal.long

ggplot(data = growth1.normal.long, aes(x = exp_period, y = growthrate, shape = wound))+
  stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.1))+ 
  stat_summary( fun = "mean", geom = "line", size = 1, position = position_dodge(width = 0.1), linetype = "dashed", aes(group = wound))+ 
  facet_wrap(~temp)+ 
  theme_classic()+
  ggtitle("Growth Rate normalized by post wound SA (mg/cm2/day)")

ggplot(data = growth1.normal.long, aes(x = wound, y = full.growth.rate, fill = temp))+geom_point(aes(col = temp))+
  stat_summary( fun = "mean", geom = "bar", position = position_dodge(width = 1))+
  theme_classic()+
  ggtitle("Mean Growth Rate")

ggplot(data = growth1.normal.long, aes(x = wound, y = growthrate, fill = temp))+geom_point(aes(col = temp))+
  stat_summary( fun = "mean", geom = "bar", position = position_dodge(width = 1))+
  facet_wrap(~exp_period)+ 
  theme_classic()+
  ggtitle("Mean Growth Rate split between exp time periods")


```
#models for effect of treatments and interaction of treatments: m1) for growth rate across entire exp, m2) growth rate during pt1, and m3) growth rate during pt 2
```{r}
#temperature, wounding, and the interactions as fixed effects and genotype (parent colony) as random effects.
growth1.normal$genotype<- as.factor(growth1.normal$genotype)

m1.rand <- lme(full.growth.rate ~ wound + temp + wound:temp, random = ~1|genotype, data = growth1.normal)
  summary(m1.rand)

m1.null<-gls(full.growth.rate ~ wound + temp + wound:temp , data = growth1.normal)
  summary(m1.null)

m2 <- lme(pt1.growth.rate ~ wound + temp + wound:temp, random = ~1|genotype, data = growth1.normal)
  summary(m2)
  
m3 <- lme(pt2.growth.rate ~ wound + temp + wound:temp, random = ~1|genotype, data = growth1.normal)
  summary(m3)

m4 <- lmer(full.growth.rate ~ wound + temp + wound:temp + (1|genotype), data = growth1.normal)
  summary(m4)
m4.null<- lm(full.growth.rate ~ wound + temp + wound:temp, data = growth1.normal)
  summary(m4.null)
  
a<- glm(full.growth.rate ~ wound*temp, data = growth1.normal)
  summary(a)
b<- glm(full.growth.rate ~ wound + temp, data = growth1.normal)
  summary(b)
  
# Obtain  estimated marginal means (EMMs) for all levels of the predictors
emmeans_results <- emmeans(m4, ~ wound + temp + wound:temp)
# Display the EMMs
print(emmeans_results)
# Pairwise comparisons for the levels within each predictor
pairwise_results <- pairs(emmeans_results)
# Display the pairwise comparisons
print(pairwise_results)

```
# temperature only 
```{r}
temp.null<- lm(full.growth.rate ~ temp, data = growth1.normal)
  summary(temp.null)
temp.model <- lmer(full.growth.rate ~ temp + (1|genotype), data = growth1.normal)
  summary(temp.model)
emmeans_temp <- emmeans(temp.model, ~ temp)
print(emmeans_temp)
pairwise_temp <- pairs(emmeans_temp)
print(pairwise_temp)
  
```





```{r}
growth1.normal.long$genotype<- as.factor(growth1.normal.long$genotype)

t1 <- lm(growthrate ~ exp_period + wound + temp + wound:temp, data = growth1.normal.long)
  summary(t1)
  
t2 <- lmer(growthrate ~ exp_period + wound + temp + wound:temp + (1|genotype), data = growth1.normal.long)
  summary(t2)
  
# Obtain  estimated marginal means (EMMs) for all levels of the predictors
emmeans_results1 <- emmeans(t2, ~ exp_period + wound + temp + wound:temp)
# Display the EMMs
print(emmeans_results1)
# Pairwise comparisons for the levels within each predictor
pairwise_results1 <- pairs(emmeans_results1)
# Display the pairwise comparisons
print(pairwise_results1)
  
```




## standardized by final SA only 
```{r}
growth%>%mutate(full.growth.rate = (full.growth / final.SA),
                pt1.growth.rate = (pt1.growth / final.SA),
                pt2.growth.rate = (pt2.growth / final.SA))%>%left_join(sample.info, by = "coral_id") -> growth2.normal
growth2.normal$wound<- as.factor(growth2.normal$wound)

ggplot(growth2.normal, aes(x = wound, y = full.growth.rate, col = temp))+
  geom_boxplot()+ggtitle("Growth Rate over entire duration of exp (normalized by final SA")

# checking to see if initial coral size was correlated with growth rate 
ggplot(growth2.normal, aes (x=initial_weight, y = full.growth.rate))+ geom_point()
ggplot(growth2.normal, aes (x=day1_weight, y = full.growth.rate))+ geom_point()

summary(growth2.normal)
#make data long again
growth2.normal %>% 
  pivot_longer(cols = c(pt1.growth.rate, pt2.growth.rate),
               names_to = "exp_period",
               values_to = "growthrate")-> growth2.normal.long

ggplot(data = growth2.normal.long, aes(x = exp_period, y = growthrate, shape = wound))+ 
  stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.1))+ 
  stat_summary( fun = "mean", geom = "line", size = 1, position = position_dodge(width = 0.1), linetype = "dashed", aes(group = wound))+ 
  facet_wrap(~temp)+ 
  theme_classic()+
  ggtitle("Growth Rate normalized by final SA (mg/cm2/day)")
```

# Growth Rate of control corals only, starting from initial timepoint 
```{r}
mass.SA%>%mutate(initial.weight.norm = initial_weight / final.SA,
                 day1.weight.norm = day1_weight / final.SA,
                 day10.weight.norm = day10_weight / final.SA, #this is just final SA, lol
                 final.weight.norm = final_weight / final.SA) -> mass.SA.normal.controlsonly
growth.rates3<- mass.SA.normal.controlsonly%>%mutate(full.growth.rate = ((final.weight.norm - initial.weight.norm)*1000)/19,
                                      pt1.growth.rate = ((day1.weight.norm - initial.weight.norm)*1000)/7,
                                      pt2.growth.rate = ((day10.weight.norm - day1.weight.norm)*1000)/9,
                                      pt3.growth.rate = ((final.weight.norm - day10.weight.norm)*1000)/9)%>%left_join(sample.info, by = "coral_id")%>%filter(wound == 0)
growth.rates3$wound<- as.factor(growth.rates3$wound)

ggplot(growth.rates3, aes(x = wound, y = full.growth.rate, col = temp))+
  geom_boxplot()+ggtitle("Growth Rate of non-wounded corals standardized by final SA over entire exp (mg/cm2/day)")

# checking to see if initial coral size was correlated with growth rate 
ggplot(growth.rates3, aes (x=initial.weight.norm, y = full.growth.rate))+ geom_point()
ggplot(growth.rates3, aes (x=day1.weight.norm, y = full.growth.rate))+ geom_point()

summary(growth.rates3)
#make data long again
growth.rates3 %>% 
  pivot_longer(cols = c(pt1.growth.rate, pt2.growth.rate, pt3.growth.rate),
               names_to = "exp_period",
               values_to = "growthrate")%>%filter(!coral_id == 32)-> growth.rates.long3 #filtered out coral 32 here bc it has irrationally large growth rate from initial to day 1 (maybe i switched it out for another coral?)

ggplot(data = growth.rates.long3, aes(x = exp_period, y = growthrate))+ geom_boxplot()+
  stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.1))+ 
  stat_summary( fun = "mean", geom = "line", size = 1, position = position_dodge(width = 0.1), linetype = "dashed", aes(group = wound))+ 
  theme_classic()+
  ggtitle("Growth Rate of non-wounded corals standardized by final SA (mg/cm2/day)")

ggplot(data = growth.rates.long3, aes(x = exp_period, y = growthrate))+ geom_point()+
  stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.1))+ 
  stat_summary( fun = "mean", geom = "line", size = 1, position = position_dodge(width = 0.1), linetype = "dashed", aes(group = wound))+ 
  facet_wrap(~temp)+ 
  theme_classic()+
  ggtitle("Growth Rate of non-wounded corals standardized by final SA (mg/cm2/day)")

```

