---
title: "ACR_Regen_Growth"
author: "Ninah Munk"
date: "2024-02-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("dplyr")
library("ggplot2")
library("readxl")
library("janitor")
library('readxl')
library("emmeans")
library("lme4")
library("Matrix")
library("nlme")
library("Rmisc")
```

# Calculating Skeletal Mass from Buoyant Weight 

Read in Raw Data. Combine into one long data frame.
```{r}
weight_initial<- read_xlsx("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Growth/Data/bouyantweight_initial.xlsx", sheet= "raw_data")%>%clean_names()%>%select(-calibration)%>%mutate(timepoint = "initial")
weight_24hr<- read_xlsx("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Growth/Data/bouyantweight_24hr.xlsx", sheet= "raw_data")%>%clean_names()%>%select(-bouyantweight_redo_if_applicable)%>%mutate(timepoint = "postwound")
weight_day10<- read_xlsx("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Growth/Data/bouyantweight_day10.xlsx", sheet= "raw_data")%>%clean_names()%>%select(-bouyantweight_redo_if_applicable)%>%mutate(timepoint = "day10")
weight_final<- read_xlsx("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Growth/Data/bouyantweight_final.xlsx", sheet= "raw_data")%>%clean_names()%>%mutate(timepoint = "final")

rbind(weight_initial, weight_24hr, weight_day10, weight_final)%>%select(-date, -group, -temp_c, -salinity_ppt, -salt_temp_c ) -> weight.long
```

# Percent change in Bouyant Weight
```{r}
weight.long%>%select(coral_id, bouyantweight_g, timepoint)%>%pivot_wider(names_from = timepoint, values_from = bouyantweight_g) -> weight.wide

weight.wide%>%mutate(percent.change = ((final - postwound)/postwound)*100) -> bouyant.change
#clean data
corals_to_remove <- c(11, 5, 37, 104, 71, 17, 29, 100)
bouyant.change <- bouyant.change[!(bouyant.change$coral_id %in% corals_to_remove), ]

#bring in sample info\
sample.info<- read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Respiration/Data/samp_info.csv")

bouyant.change%>%left_join(sample.info, by = "coral_id")-> bouyant.change

bouyant.change%>%unite("treatment", wound:temp, remove = F) -> bouyant.change

ggplot(data = bouyant.change, aes(x = treatment, y = percent.change))+geom_boxplot()

# Perform ANOVA
bw.model <- aov(percent.change ~ treatment, data = bouyant.change)
summary(bw.model)
TukeyHSD(bw.model)

bw.model2 <- aov(percent.change ~ as.factor(wound)*as.factor(temp), data = bouyant.change)
TukeyHSD(bw.model2)

AIC(bw.model, bw.model2)

mod<- lmer(percent.change ~ as.factor(wound)*as.factor(temp) + (1|genotype), data = bouyant.change)
anova(mod)

bouyant.change%>%aggregate(percent.change ~ treatment, FUN=mean)-> meanchange
# Calculating standard error for error bars in the plot.
bouyant.change%>%aggregate(percent.change ~ treatment, FUN = function(x) sd(x)/sqrt(length(x))) -> std_err 

# Assuming all contrasts were significant for simplicity; this may change based on actual results.
meanchange$letter <- c("a", "a", "a","a", "a", "a")

# Plotting
# Creating a bar plot with error bars and significant letters to clearly show the differences between groups.
# Using specified fill colors for clarity.
ggplot(meanchange, aes(x=treatment, y=percent.change, label=letter)) +
  geom_bar(stat="identity", fill=c("gray", "skyblue", "darkblue","purple", "pink", "orange")) +
  geom_errorbar(aes(ymin=percent.change-std_err$percent.change, ymax=percent.change+std_err$percent.change), width=.2) +
  geom_text(aes(label=letter), nudge_y=0.0005, color="red") +
  labs(title="Percent Change Bouyant Weight", x="Treatment", y="Growth (g/cm^2/day)")+ 
  scale_x_discrete(labels = c('No Wound, Ambient','No Wound, Hot','Fragmentation, Ambient', 'Fragmentation, Hot', 'Abrasion, Ambient', 'Abrasion, Hot'))+
  theme_classic()


```
```{r}
#FILTERING BY WORM DAMAGE

read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Data/worm_damage.csv")-> worm

bouyant.change%>%left_join(worm, by = "coral_id") -> bouyant.change.worm

ggplot(data = bouyant.change.worm, aes(x = percent_affected, y = percent.change))+geom_point()

bouyant.change.lowworm = subset(bouyant.change.worm, percent_affected == 0)

mod2<- lmer(percent.change ~ wound.x*temp + (1|genotype.x), data = bouyant.change.lowworm)
anova(mod2)

ggplot(data = bouyant.change.lowworm, aes(x = percent_affected, y = percent.change))+geom_point()
```



```{r}
weight.long%>% 
  mutate(density_stopper = (air_weight_g * 0.9965)/(air_weight_g - fresh_weight_g))%>%
  mutate(density_sw = (air_weight_g - salt_weight_g)/ (air_weight_g / density_stopper))%>%
  mutate(dry_mass = bouyantweight_g / (1 - (density_sw/density_aragonite)))%>%
  mutate(coral_vol = (dry_mass / (density_aragonite - density_sw))) -> skeletal.mass

#clean data
corals_to_remove <- c(11, 5, 37, 104, 71, 17, 29, 100)
skeletal.mass.clean <- skeletal.mass[!(skeletal.mass$coral_id %in% corals_to_remove), ]
  
# coral dry mass and volume should be positively correlated with buoyant weight
ggplot( data = skeletal.mass.clean, aes(x= bouyantweight_g, y = dry_mass))+ geom_point() + ggtitle("Bouyant Weight X Dry Mass")
ggplot( data = skeletal.mass.clean, aes(x= bouyantweight_g, y = coral_vol))+ geom_point()+ ggtitle("Bouyant Weight X Coral Volume")
```
 
# Calculating Coral Surface Area 
 
 ## Geometric coral surface areas
```{r}
# read in SA and calculate SA of a curved cylinder of each branch for every coral frag
data<- read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Surface_Area/Data/geometric_SA_initial.csv")%>%clean_names()%>%
  mutate(branch_height_cm = branch_height_mm / 10) %>%
  mutate(avg_diameter_mm= (diameter_base_mm + diameter_tip_mm) / 2)%>%
  mutate(radius_cm = (avg_diameter_mm/2)/10)%>%
  mutate(radius_tip_cm = (diameter_tip_mm/2)/10)%>%
  mutate(CSA_cm2= (2*3.14*(radius_cm*branch_height_cm) + 3.14*(radius_tip_cm)^2))

# add surface areas of branches together 
  summarized_initial_data <- data %>%
  group_by(coral_id) %>%
  reframe(SA = sum(CSA_cm2))
  
# calculate the mean apical tip diameter to use for removing SA of abrasion wound (type 2) from initial surface areas
wound_2_avg_tip_diameter<- mean(data$diameter_tip_mm)

# read in data sheet with height, tip, & base measurements for fragment wound (type 1)
wound_amount<-read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Surface_Area/Data/wound_type_1_measurements.csv")%>%clean_names()
# filling in height for wound type 2 (10 mm ~ 1cm)
wound_amount$branch_h_mm <- ifelse(wound_amount$wound == 2 & wound_amount$coral_id %in% unique(wound_amount$coral_id[wound_amount$wound == 2]), 10, wound_amount$branch_h_mm)
#filling in tip diameter for wound type 2 based on average diameter of apical branch tips from initial geometric measurements (see above)
wound_amount$tip_d_mm <- ifelse(wound_amount$wound == 2 & wound_amount$coral_id %in% unique(wound_amount$coral_id[wound_amount$wound == 2]), wound_2_avg_tip_diameter, wound_amount$tip_d_mm)
# make all other NAs in the data frame zero 
wound_amount[is.na(wound_amount)] <- 0

# calculate SA of removed coral tissue from wounding treatments
data2<- wound_amount%>%
  mutate(branch_height_cm = branch_h_mm / 10) %>%
  mutate(avg_diameter_mm= (base_d_mm + tip_d_mm) / 2)%>%
  mutate(radius_cm = (avg_diameter_mm/2)/10)%>%
  mutate(radius_tip_cm = (tip_d_mm/2)/10)%>%
  mutate(CSA_cm2= (2*3.14*(radius_cm*branch_height_cm) + 3.14*(radius_tip_cm)^2))

# new data frame with SA to subtract from initial SA
SA_to_subtract<-data2%>%select(coral_id, wound, CSA_cm2)
# join data frame with data frame of initial SA's
initialSA_woundSA<- left_join(summarized_initial_data, SA_to_subtract, by= 'coral_id')
# subtract removed SA from initial SA to get post wound SA
SA_postwound<-initialSA_woundSA%>%mutate(SA_postwound = SA - CSA_cm2)
```

try making SA removed equal for all fragmented corals
```{r}
#mean(SA_to_subtract$CSA_cm2[SA_to_subtract$wound ==1])
#SA_to_subtract$CSA_cm2[SA_to_subtract$wound == 1] <- 1.07
```


 ## Wax dipping method calibration curve
```{r}
#wax dipping calibration: calculate surface area of dowels 
calibration<- read.csv('/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Surface_Area/Data/20230712_wax_calibration.csv')%>%clean_names()%>%
  mutate(wax_weight_g = postwax_weight_g - prewax_weight_g)%>%
  mutate(cal_radius_mm = diameter_mm / 2)%>%
  mutate(cal_radius_cm = cal_radius_mm /10)%>%
  mutate(height_cm = height_mm / 10)%>%
  mutate(CSA_cm2= ((2*3.14*cal_radius_cm*height_cm) + 3.14*(cal_radius_cm)^2)) #curved surface area (CSA) = 2piRH + piR^2 (one area of circle for top of coral)

#calculate the curve coefficients for slope and intercept to apply as the standard
stnd.curve <- lm(CSA_cm2~wax_weight_g, data=calibration)
plot(CSA_cm2~wax_weight_g, data=calibration)
stnd.curve$coefficients
summary(stnd.curve)$r.squared

#bring in the datasheet with coral samples 
smpls<- read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Surface_Area/Data/20230712_wax_weights.csv")%>%clean_names()%>%
#subtract postwax weight from prewax weight
mutate(wax_weight_g = postwax_weight_g - prewax_weight_g)
#Calculate surface area using the standard curve
smpls$SA <- stnd.curve$coefficients[2] * smpls$wax_weight_g + stnd.curve$coefficients[1]

# Can we create another standard curve using final wax weights and final coral skeletal mass
skeletal.mass%>%filter(timepoint == "final")%>%select(coral_id, dry_mass)-> final.mass

# combine wax weight and coral weight 
smpls%>%select(coral_id, wax_weight_g)%>%
  left_join(final.mass, by = "coral_id")-> waxXmass

stnd.curve1 <- lm(dry_mass~wax_weight_g, data=waxXmass)
plot(dry_mass~wax_weight_g, data=waxXmass)
stnd.curve1$coefficients
summary(stnd.curve1)$r.squared #decent rsquared

#bring in sample info to filter out controls
sample.info<- read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Respiration/Data/samp_info.csv")

# calculate mass of removed coral skeleton
skeletal.mass%>%select(coral_id, timepoint, dry_mass)%>%pivot_wider(names_from = timepoint,
              values_from = dry_mass)%>%mutate(removed.mass = initial - postwound)%>%left_join(sample.info, by = "coral_id")%>%filter(!wound == 0) -> removed.mass

# lets check that final SA is positively correlated to final weight 
smpls%>%left_join(final.mass, by = "coral_id")-> final.SAXmass
ggplot(data = final.SAXmass, aes(x = dry_mass, y = SA))+ geom_point()

stnd.curve2 <- lm(dry_mass~SA, data=final.SAXmass)
plot(dry_mass~SA, data=final.SAXmass)
stnd.curve2$coefficients
summary(stnd.curve2)$r.squared #decent rsquared

```
 
 SA's derived geometrically are positively correlated to wax dip derived SA
```{r}
SA_postwound%>%select(coral_id, SA, SA_postwound)%>%dplyr::rename(SA_initial = SA)%>%left_join(smpls%>%select(coral_id, SA)%>%dplyr::rename(SA_final = SA))%>%left_join(sample.info)-> all.SA

stnd.curve3 <- lm(SA_initial~SA_final, data=all.SA)
plot(SA_initial~SA_final, data=all.SA)
stnd.curve3$coefficients
summary(stnd.curve3)$r.squared

ggplot(data = all.SA, aes(x= SA_postwound, y = SA_final, col = temp))+geom_point()+
  stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x)


```
 
Corals weight more at day 1 post wounding then they do at initial weighing (i.e., Corals grew after initial weighing so subtracting post wound weight from initial weight to get weight of removed mass wont work). This makes sense because corals had high growth rates before exp started (see below growth rates of unwounded corals pre start of experiment). Below I try standardizing corals a few different ways. I subtract geometrically calculated SA of wounds from final SA for standardizing the postwound(day1) measurements. 

```{r}
#final wax SA - wound geometric SA = post wound SA
smpls%>%left_join(SA_to_subtract, by = "coral_id")%>%mutate(postwound.SA = SA - CSA_cm2)%>%select(coral_id, postwound.SA) -> postwound.SA
```

Combine Initial, Postwound, and Final Surface Areas into one dataframe
```{r}
summarized_initial_data%>%mutate(timepoint = "initial") -> SA.1
postwound.SA%>%dplyr::rename(SA = postwound.SA)%>%mutate(timepoint = "postwound") -> SA.2
smpls%>%select(coral_id, SA)%>%mutate(timepoint = "day10") -> SA.3
smpls%>%select(coral_id, SA)%>%mutate(timepoint = "final") -> SA.4


rbind(SA.1, SA.2, SA.3, SA.4) -> SA.long
#clean data
corals_to_remove <- c(11, 5, 37, 104, 71, 17, 29, 100) #100 is an outlier it has double the growth rate of the next fastest growing coral 
SA.long.clean <- SA.long[!(SA.long$coral_id %in% corals_to_remove), ]

```


# Growth Rates analyze with varying SA standardizations

## Calculate Growth Rates NOT standardized by SA
```{r}
skeletal.mass.clean%>%select(coral_id, timepoint, dry_mass)%>%pivot_wider(names_from = timepoint,
              values_from = dry_mass)%>%
  dplyr::rename(initial_weight = initial,
         day1_weight = postwound,
         day10_weight = day10, 
         final_weight = final) -> skeletal.mass.clean.wide

SA.long.clean%>%pivot_wider(names_from = timepoint,
              values_from = SA)%>%
  dplyr::rename(initial.SA = initial,
         day1.SA = postwound,
         day10.SA = day10, 
         final.SA = final) -> SA.clean.wide

SA.clean.wide%>%left_join(skeletal.mass.clean.wide, by = "coral_id") -> mass.SA

mass.SA%>%mutate(full.growth = (((final_weight - day1_weight)/19)*1000),
                 pt1.growth = (((day10_weight - day1_weight)/9)*1000),
                 pt2.growth = (((final_weight - day10_weight)/9)*1000))-> growth
growth.not.normal<- growth%>%left_join(sample.info, by = "coral_id")

growth.not.normal$wound<- as.factor(growth.not.normal$wound)

ggplot(growth.not.normal, aes(x = wound, y = full.growth, col = temp))+
  geom_boxplot()+ggtitle("Growth Rate over entire duration of exp (not normalized by SA)")

# checking to see if initial coral size was correlated with growth rate 
ggplot(growth.not.normal, aes (x=initial_weight, y = full.growth))+ geom_point()
ggplot(growth.not.normal, aes (x=day1_weight, y = full.growth))+ geom_point()

ggplot(data = growth.not.normal, aes(x = day1.SA, y = full.growth, col = temp))+
  geom_point()+
  geom_smooth(method='lm')+
  facet_wrap(~wound)+
  ggtitle("Growth (mg) ~ Initial size")

ggplot(data = growth.not.normal, aes(x = day1.SA, y = full.growth, col = wound))+
  geom_point()+
  geom_smooth(method='lm')+
  #facet_wrap(~temp)+
  ggtitle("Growth (mg) ~ Initial size")

summary(growth.not.normal)
#make data long again
growth.not.normal %>% 
  pivot_longer(cols = c(pt1.growth, pt2.growth),
               names_to = "exp_period",
               values_to = "growthrate")-> growth.not.normal.long

ggplot(data = growth.not.normal.long, aes(x = day1.SA, y = growthrate, shape = wound))+ 
  stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.1))+ 
 # stat_summary( fun = "mean", geom = "line", size = 1, position = position_dodge(width = 0.1), linetype = "dashed", aes(group = wound))+ 
  facet_wrap(~temp)+ 
  theme_classic()+
  ggtitle("not standardized growth = mg/day")

ggplot(data = growth.not.normal.long, aes(x = day1.SA, y = growthrate, col = wound))+
  geom_point()+
  geom_smooth(method='lm')+
  facet_wrap(~exp_period + temp)+
  ggtitle("Growth (mg) ~ Initial size")




```
## WOUNDING: plot growth as change in weight not as a rate 
```{r}
skeletal.mass.clean%>%select(coral_id, timepoint, dry_mass)%>%left_join(all.SA, by = "coral_id") -> weight

weight%>%filter(timepoint == "initial")%>%mutate(weight = dry_mass / SA_initial)-> initial.data
weight%>%filter(!timepoint == "initial")%>%mutate(weight = dry_mass / SA_postwound)-> postwound.data
rbind(initial.data, postwound.data) -> weight.norm

weight.norm%>%filter(!wound == 0)%>%mutate(wound_status = 'wounded')-> wounded
weight.norm%>%filter(wound == 0)%>%mutate(wound_status = 'not wounded')-> not.wounded
rbind(wounded, not.wounded)-> weight.norm

# Define a mapping from the original values to the desired numeric values
timepoints.numeric <- c("initial" = -7, "postwound" = 1, "day10" = 10, "final" = 19)

# Use mutate to create a new column with the numeric values based on the mapping
weight.norm <- weight.norm %>%
  mutate(timepoint_numeric = case_when(
    timepoint %in% names(timepoints.numeric) ~ timepoints.numeric[timepoint],
    TRUE ~ NA_real_
  ))


ggplot(data = weight.norm, aes(x = timepoint_numeric, y = weight, col = wound_status))+geom_point()+geom_smooth(method = 'lm')+theme_classic()
ggplot(data = weight.norm, aes(x = timepoint_numeric, y = weight, col = as.factor(wound)))+geom_point()+geom_smooth(method = 'lm')
# FILTERING OUT INITIAL TIMEPOINT 
ggplot(data = weight.norm%>%filter(!timepoint == "initial"), aes(x = timepoint_numeric, y = weight, col = wound_status))+geom_point()+geom_smooth(method = 'lm')+theme_classic()
ggplot(data = weight.norm%>%filter(!timepoint == "initial"), aes(x = timepoint_numeric, y = weight, col = as.factor(wound)))+geom_point()+geom_smooth(method = 'lm')+theme_classic()+facet_wrap(~temp)+  
  labs(title = "Change in Weight", y = "Weight (g/cm2)", x = "Day")



```
# glm(weight~time*wound)
WOUNDING: plot growth rate, weights were standardized to g/cm2 accounting for wounding treatments above so now can look at growth rate across entire experiment 
```{r}
weight.norm%>%filter(!timepoint == "postwound", !timepoint == "day10")%>%select(-dry_mass, -SA_initial, -SA_postwound, -SA_final, -timepoint_numeric)%>%pivot_wider(names_from = timepoint, values_from = weight)-> weight.norm.wide

weight.norm.wide%>%mutate(growthrate = (final - initial)/26)%>%pivot_longer(cols = c(initial, final), names_to = "timepoint", values_to = "weight") -> weight.norm.long

ggplot(data = weight.norm.long, aes(x = wound_status, y = growthrate, col = wound_status))+geom_boxplot()
ggplot(data = weight.norm.long, aes(x = as.factor(wound), y = growthrate, col = as.factor(wound)))+geom_boxplot()

# FILTERING OUT INITIAL TIMEPOINT 
weight.norm%>%filter(!timepoint == "initial", !timepoint == "day10")%>%select(-dry_mass, -SA_initial, -SA_postwound, -SA_final, -timepoint_numeric)%>%pivot_wider(names_from = timepoint, values_from = weight)-> weight.norm.wide2

weight.norm.wide2%>%mutate(growthrate = (final - postwound)/19)%>%pivot_longer(cols = c(postwound, final), names_to = "timepoint", values_to = "weight")%>%filter(timepoint == "final")%>%mutate(growthrate = growthrate*1000) -> weight.norm.long2

ggplot(data = weight.norm.long2, aes(x = wound_status, y = growthrate, col = wound_status))+geom_boxplot()+theme_classic()+labs(x = "Wound Status", y = "Growth (mg/cm2/day)")
ggplot(data = weight.norm.long2, aes(x = as.factor(wound), y = growthrate, col = as.factor(wound)))+geom_boxplot()
ggplot(data = weight.norm.long2, aes(x = as.factor(wound), y = growthrate, col = as.factor(wound)))+geom_boxplot()+facet_wrap(~temp)

```
```{r}
# FILTERING OUT INITIAL TIMEPOINT 
weight.norm%>%filter(!timepoint == "initial", !timepoint == "final")%>%select(-dry_mass, -SA_initial, -SA_postwound, -SA_final, -timepoint_numeric)%>%pivot_wider(names_from = timepoint, values_from = weight)-> weight.norm.wide3

weight.norm.wide3%>%mutate(growthrate = (day10 - postwound)/10)%>%pivot_longer(cols = c(postwound, day10), names_to = "timepoint", values_to = "weight")%>%filter(timepoint == "day10")%>%mutate(growthrate = growthrate*1000) -> weight.norm.long3

ggplot(data = weight.norm.long3, aes(x = wound_status, y = growthrate, col = wound_status))+geom_boxplot()+theme_classic()+labs(x = "Wound Status", y = "Growth (mg/cm2/day)")
ggplot(data = weight.norm.long3, aes(x = as.factor(wound), y = growthrate, col = as.factor(wound)))+geom_boxplot()
ggplot(data = weight.norm.long3, aes(x = as.factor(wound), y = growthrate, col = as.factor(wound)))+geom_boxplot()+facet_wrap(~temp)
```



#aov w tukey but check normality 
## TEMPERATURE: plot growth as change in weight not as a rate w respect to temp treatment
```{r}
ggplot(data = weight.norm, aes(x = timepoint_numeric, y = weight, col = temp))+geom_point()+geom_smooth(method = "lm")
ggplot(data = weight.norm.long, aes(x = temp, y = growthrate, col = temp))+geom_boxplot()


# FILTERING OUT INITIAL TIMEPOINT 
ggplot(data = weight.norm%>%filter(!timepoint == "initial"), aes(x = timepoint_numeric, y = weight, col = temp))+geom_point()+geom_smooth(method = "lm")
ggplot(data = weight.norm.long2, aes(x = temp, y = growthrate, col = temp))+geom_boxplot()
```

```{r}
weight.norm%>%unite("treatment", wound:temp, remove = F) -> weight.norm.interaction
weight.norm.long%>%unite("treatment", wound:temp, remove = F) -> weight.norm.interaction.long
weight.norm.long2%>%unite("treatment", wound:temp, remove = F) -> weight.norm.interaction.long2

ggplot(data = weight.norm.interaction, aes(x = timepoint_numeric, y = weight, col = treatment))+geom_point()+geom_smooth(method = "lm")
ggplot(data = weight.norm.interaction.long, aes(x = as.factor(wound), y = growthrate, col = temp))+geom_boxplot()
ggplot(data = weight.norm.interaction.long, aes(x = treatment, y = growthrate, col = treatment))+geom_boxplot()

# FILTERING OUT INITIAL TIMEPOINT 
ggplot(data = weight.norm.interaction%>%filter(!timepoint == "initial"), aes(x = timepoint_numeric, y = weight, col = treatment))+geom_point()+geom_smooth(method = "lm")
ggplot(data = weight.norm.interaction.long2, aes(x = as.factor(wound), y = growthrate, col = temp))+geom_boxplot()
ggplot(data = weight.norm.interaction.long2, aes(x = treatment, y = growthrate, col = treatment))+geom_boxplot()
ggplot(data = weight.norm.interaction.long2, aes(x = as.factor(wound), y = growthrate, col = temp))+geom_boxplot()

```

#aov + tukey 
```{r}
hist(weight.norm.interaction.long2$growthrate)
shane<- aov(growthrate ~ treatment, data = weight.norm.interaction.long2%>%filter(timepoint == "final"))
plot(shane)
summary(shane)
TukeyHSD(shane)
# only report p value from aov 
```


# code from adrian on orthogonal contrasts
```{r}
weight.norm.interaction.long2%>%filter(timepoint == "final")%>%select(-timepoint)%>%mutate(treatment = as.factor(treatment), wound = as.factor(wound)) -> growthrate
ggplot(data = growthrate, aes(x = treatment, y = growthrate, col = treatment))+geom_boxplot()
# Perform ANOVA
growth.model <- aov(data = growthrate, growthrate ~ treatment)

#these next 2 lines of code do the same thing 
TukeyHSD(growth.model)
#comp <- glht(growth.model, linfct = mcp(treatment = "Tukey"))
#summary(comp)

ancova_model <- aov(growthrate ~ as.factor(wound) + temp, data = growthrate)
TukeyHSD(ancova_model)

# Adjusted p-values for multiple comparisons
# This plot helps visualize if there are any significant differences between groups after adjustments.
#plot(comp)

# Step 3: Plot data with significant letters
# Calculating mean growth for each group to plot.
growthrate%>%aggregate(growthrate ~ treatment, FUN=mean)-> meangrowthrate
# Calculating standard error for error bars in the plot.
growthrate%>%aggregate(growthrate ~ treatment, FUN = function(x) sd(x)/sqrt(length(x))) -> std_err 

# Assuming all contrasts were significant for simplicity; this may change based on actual results.
meangrowthrate$letter <- c("a", "a", "a","a", "a", "a")

# Plotting
# Creating a bar plot with error bars and significant letters to clearly show the differences between groups.
# Using specified fill colors for clarity.
ggplot(meangrowthrate, aes(x=treatment, y=growthrate, label=letter)) +
  geom_bar(stat="identity", fill=c("gray", "skyblue", "darkblue","purple", "pink", "orange")) +
  geom_errorbar(aes(ymin=growthrate-std_err$growthrate, ymax=growthrate+std_err$growthrate), width=.2) +
  geom_text(aes(label=letter), nudge_y=0.0005, color="red") +
  labs(title="Mean Coral Growth", x="Treatment", y="Growth (g/cm^2/day)")+ 
  scale_x_discrete(labels = c('No Wound, Ambient','No Wound, Hot','Fragmentation, Ambient', 'Fragmentation, Hot', 'Abrasion, Ambient', 'Abrasion, Hot'))+
  theme_classic()

```




```{r}
mod2<- lmer(growthrate ~ wound*temp + (1|genotype), data = growthrate)
anova(mod2)

model<- aov(full.growth ~ wound + as.factor(temp), data = growth.not.normal)
summary(model)

model2<- glm(full.growth ~ wound + as.factor(temp), data = growth.not.normal)
summary(model2)

model3<- lm(full.growth ~ wound + as.factor(temp) , data = growth.not.normal)
summary(model3)

emmeans_results <- emmeans(model2, ~ wound + temp)
# Display the EMMs
print(emmeans_results)
# Pairwise comparisons for the levels within each predictor
pairwise_results <- pairs(emmeans_results)
# Display the pairwise comparisons
print(pairwise_results)

```
```{r}
# FILTERING BY WORM DAMAGE 
#read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Data/worm_damage.csv")-> worm
library(rstatix)
library(Rmisc)
library(lmerTest)
growthrate%>%left_join(worm, by = "coral_id") -> growthrate.worm
#full model
mod2<- lmer(growthrate ~ wound.x*temp*percent_affected + (1|genotype.x), data = growthrate.worm)
anova(mod2)

#reduced model
mod2<- lmer(growthrate ~ wound.x*temp+wound.x*percent_affected + (1|genotype.x), data = growthrate.worm)

anova(mod2)

ggplot(data = growthrate.worm, aes(x = wound.x, y = growthrate, color = temp))+geom_boxplot()+facet_wrap(~as.factor(percent_affected))

#growthrate.lowworm = subset(growthrate.worm, genotype.x == 65 & percent_affected < 20 | genotype.x == 66 | genotype.x == 67 & percent_affected < 20)

growthrate.lowworm = subset(growthrate.worm, percent_affected <= 10)

#full model, percent affected is no longer significant 
mod2<- lmer(growthrate ~ wound.x*temp*percent_affected + (1|genotype.x), data = growthrate.lowworm)
anova(mod2)

#reduced
mod2<- lmer(growthrate ~ wound.x*temp + temp*percent_affected + (1|genotype.x), data = growthrate.lowworm)
anova(mod2)

ggplot(data = growthrate.lowworm, aes(x = wound.x, growthrate))+ geom_boxplot()+facet_wrap(~temp)
ggplot(data = growthrate.lowworm, aes(x = temp, growthrate))+ geom_boxplot()+facet_wrap(~wound.x)

emmeans_noworms <- emmeans(mod2, ~ wound.x + temp)
# Display the EMMs
print(emmeans_results)
# Pairwise comparisons for the levels within each predictor
pairwise_results <- pairs(emmeans_noworms)
# Display the pairwise comparisons
print(pairwise_results)

ggplot(data = growthrate.lowworm, aes(x = wound.x, y = growthrate, color = temp))+geom_boxplot()+facet_wrap(~as.factor(percent_affected))
ggplot(data = growthrate.lowworm, aes(x = wound.x, y = growthrate, color = temp))+geom_boxplot()
ggplot(data = worm, aes(x = percent_affected))+geom_histogram()+facet_wrap(~genotype)

ggplot(data = worm, aes(x = genotype))

summarySE(data = growthrate.lowworm, measurevar = "growthrate", groupvars = c("wound.x", "temp", "genotype.x"),
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE)

test <- growthrate.lowworm %>% group_by(wound.x, temp, genotype.x) %>% reframe(samp.size=n()) 
```


# checking to see how many corals have regeneration data
```{r}
read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Data/all.wounds.csv") -> regen

growthrate.lowworm%>%left_join(regen, by = "coral_id")%>%drop_na() -> growthregen
```


## Growth rate standardized by post wound and final SA 
```{r}
growth%>%mutate(full.growth.rate = (full.growth / final.SA),
                pt1.growth.rate = (pt1.growth / day1.SA),
                pt2.growth.rate = (pt2.growth / final.SA))%>%left_join(sample.info, by = "coral_id") -> growth.normal
growth.normal$wound<- as.factor(growth.normal$wound)

ggplot(growth.normal, aes(x = wound, y = full.growth.rate, col = temp))+
  geom_boxplot()+ggtitle("Growth Rate over entire duration of exp (normalized by timepoint specific SA")

# checking to see if initial coral size was correlated with growth rate 
ggplot(growth.normal, aes (x=initial_weight, y = full.growth.rate))+ geom_point()
ggplot(growth.normal, aes (x=day1_weight, y = full.growth.rate))+ geom_point()

summary(growth.normal)
#make data long again
growth.normal %>% 
  pivot_longer(cols = c(pt1.growth.rate, pt2.growth.rate),
               names_to = "exp_period",
               values_to = "growthrate")-> growth.normal.long

ggplot(data = growth.normal.long, aes(x = exp_period, y = growthrate, shape = wound))+ 
  stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.1))+ 
  stat_summary( fun = "mean", geom = "line", size = 1, position = position_dodge(width = 0.1), linetype = "dashed", aes(group = wound))+ 
  facet_wrap(~temp)+ 
  theme_classic()+
  ggtitle("Growth Rate normalized to timepoint specific SA (mg/cm2/day)")


```

## Growth Rate standardized by post wound SA 
```{r}
growth%>%mutate(full.growth.rate = (full.growth / day1.SA),
                pt1.growth.rate = (pt1.growth / day1.SA),
                pt2.growth.rate = (pt2.growth / day1.SA))%>%left_join(sample.info, by = "coral_id") -> growth1.normal
growth1.normal$wound<- as.factor(growth1.normal$wound)

ggplot(growth1.normal, aes(x = wound, y = full.growth.rate, col = temp))+
  geom_boxplot()+ggtitle("Growth Rate over entire duration of exp (normalized by post wound SA")

# checking to see if initial coral size was correlated with growth rate 
ggplot(growth1.normal, aes (x=initial_weight, y = full.growth.rate))+ geom_point()
ggplot(growth1.normal, aes (x=day1_weight, y = full.growth.rate))+ geom_point()

summary(growth1.normal)
#make data long again
growth1.normal %>% 
  pivot_longer(cols = c(pt1.growth.rate, pt2.growth.rate),
               names_to = "exp_period",
               values_to = "growthrate")-> growth1.normal.long

ggplot(data = growth1.normal.long, aes(x = exp_period, y = growthrate, shape = wound))+
  stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.1))+ 
  stat_summary( fun = "mean", geom = "line", size = 1, position = position_dodge(width = 0.1), linetype = "dashed", aes(group = wound))+ 
  facet_wrap(~temp)+ 
  theme_classic()+
  ggtitle("Growth Rate normalized by post wound SA (mg/cm2/day)")

ggplot(data = growth1.normal.long, aes(x = wound, y = full.growth.rate, fill = temp))+geom_point(aes(col = temp))+
  stat_summary( fun = "mean", geom = "bar", position = position_dodge(width = 1))+
  theme_classic()+
  ggtitle("Mean Growth Rate")

ggplot(data = growth1.normal.long, aes(x = wound, y = growthrate, fill = temp))+geom_point(aes(col = temp))+
  stat_summary( fun = "mean", geom = "bar", position = position_dodge(width = 1))+
  facet_wrap(~exp_period)+ 
  theme_classic()+
  ggtitle("Mean Growth Rate split between exp time periods")


```
# models for effect of treatments and interaction of treatments: m1) for growth rate across entire exp, m2) growth rate during pt1, and m3) growth rate during pt 2
```{r}
#temperature, wounding, and the interactions as fixed effects and genotype (parent colony) as random effects.
growth1.normal$genotype<- as.factor(growth1.normal$genotype)

m1.rand <- lme(full.growth.rate ~ wound + temp + wound:temp, random = ~1|genotype, data = growth1.normal)
  summary(m1.rand)

m1.null<-gls(full.growth.rate ~ wound + temp + wound:temp , data = growth1.normal)
  summary(m1.null)

m2 <- lme(pt1.growth.rate ~ wound + temp + wound:temp, random = ~1|genotype, data = growth1.normal)
  summary(m2)
  
m3 <- lme(pt2.growth.rate ~ wound + temp + wound:temp, random = ~1|genotype, data = growth1.normal)
  summary(m3)

m4 <- lmer(full.growth.rate ~ wound + temp + wound:temp + (1|genotype), data = growth1.normal)
  summary(m4)
m4.null<- lm(full.growth.rate ~ wound + temp + wound:temp, data = growth1.normal)
  summary(m4.null)
  
a<- glm(full.growth.rate ~ wound*temp, data = growth1.normal)
  summary(a)
b<- glm(full.growth.rate ~ wound + temp, data = growth1.normal)
  summary(b)
  
# Obtain  estimated marginal means (EMMs) for all levels of the predictors
emmeans_results <- emmeans(m4, ~ wound + temp + wound:temp)
# Display the EMMs
print(emmeans_results)
# Pairwise comparisons for the levels within each predictor
pairwise_results <- pairs(emmeans_results)
# Display the pairwise comparisons
print(pairwise_results)

```
# temperature only 
```{r}
temp.null<- lm(full.growth.rate ~ temp, data = growth1.normal)
  summary(temp.null)
temp.model <- lmer(full.growth.rate ~ temp + (1|genotype), data = growth1.normal)
  summary(temp.model)
emmeans_temp <- emmeans(temp.model, ~ temp)
print(emmeans_temp)
pairwise_temp <- pairs(emmeans_temp)
print(pairwise_temp)
  
```





```{r}
growth1.normal.long$genotype<- as.factor(growth1.normal.long$genotype)

t1 <- lm(growthrate ~ exp_period + wound + temp + wound:temp, data = growth1.normal.long)
  summary(t1)
  
t2 <- lmer(growthrate ~ exp_period + wound + temp + wound:temp + (1|genotype), data = growth1.normal.long)
  summary(t2)
  
# Obtain  estimated marginal means (EMMs) for all levels of the predictors
emmeans_results1 <- emmeans(t2, ~ exp_period + wound + temp + wound:temp)
# Display the EMMs
print(emmeans_results1)
# Pairwise comparisons for the levels within each predictor
pairwise_results1 <- pairs(emmeans_results1)
# Display the pairwise comparisons
print(pairwise_results1)
  
```




# standardized by final SA only 
```{r}
growth%>%mutate(full.growth.rate = (full.growth / final.SA),
                pt1.growth.rate = (pt1.growth / final.SA),
                pt2.growth.rate = (pt2.growth / final.SA))%>%left_join(sample.info, by = "coral_id") -> growth2.normal
growth2.normal$wound<- as.factor(growth2.normal$wound)

ggplot(growth2.normal, aes(x = wound, y = full.growth.rate, col = temp))+
  geom_boxplot()+ggtitle("Growth Rate over entire duration of exp (normalized by final SA")

# checking to see if initial coral size was correlated with growth rate 
ggplot(growth2.normal, aes (x=initial_weight, y = full.growth.rate))+ geom_point()
ggplot(growth2.normal, aes (x=day1_weight, y = full.growth.rate))+ geom_point()

summary(growth2.normal)
#make data long again
growth2.normal %>% 
  pivot_longer(cols = c(pt1.growth.rate, pt2.growth.rate),
               names_to = "exp_period",
               values_to = "growthrate")-> growth2.normal.long

ggplot(data = growth2.normal.long, aes(x = exp_period, y = growthrate, shape = wound))+ 
  stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.1))+ 
  stat_summary( fun = "mean", geom = "line", size = 1, position = position_dodge(width = 0.1), linetype = "dashed", aes(group = wound))+ 
  facet_wrap(~temp)+ 
  theme_classic()+
  ggtitle("Growth Rate normalized by final SA (mg/cm2/day)")
```

# Growth Rate of control corals only, starting from initial timepoint 
```{r}
mass.SA%>%mutate(initial.weight.norm = initial_weight / final.SA,
                 day1.weight.norm = day1_weight / final.SA,
                 day10.weight.norm = day10_weight / final.SA, #this is just final SA, lol
                 final.weight.norm = final_weight / final.SA) -> mass.SA.normal.controlsonly
growth.rates3<- mass.SA.normal.controlsonly%>%mutate(full.growth.rate = ((final.weight.norm - initial.weight.norm)*1000)/19,
                                      pt1.growth.rate = ((day1.weight.norm - initial.weight.norm)*1000)/7,
                                      pt2.growth.rate = ((day10.weight.norm - day1.weight.norm)*1000)/9,
                                      pt3.growth.rate = ((final.weight.norm - day10.weight.norm)*1000)/9)%>%left_join(sample.info, by = "coral_id")%>%filter(wound == 0)
growth.rates3$wound<- as.factor(growth.rates3$wound)

ggplot(growth.rates3, aes(x = wound, y = full.growth.rate, col = temp))+
  geom_boxplot()+ggtitle("Growth Rate of non-wounded corals standardized by final SA over entire exp (mg/cm2/day)")

# checking to see if initial coral size was correlated with growth rate 
ggplot(growth.rates3, aes (x=initial.weight.norm, y = full.growth.rate))+ geom_point()
ggplot(growth.rates3, aes (x=day1.weight.norm, y = full.growth.rate))+ geom_point()

summary(growth.rates3)
#make data long again
growth.rates3 %>% 
  pivot_longer(cols = c(pt1.growth.rate, pt2.growth.rate, pt3.growth.rate),
               names_to = "exp_period",
               values_to = "growthrate")%>%filter(!coral_id == 32)-> growth.rates.long3 #filtered out coral 32 here bc it has irrationally large growth rate from initial to day 1 (maybe i switched it out for another coral?)

ggplot(data = growth.rates.long3, aes(x = exp_period, y = growthrate))+ geom_boxplot()+
  stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.1))+ 
  stat_summary( fun = "mean", geom = "line", size = 1, position = position_dodge(width = 0.1), linetype = "dashed", aes(group = wound))+ 
  theme_classic()+
  ggtitle("Growth Rate of non-wounded corals standardized by final SA (mg/cm2/day)")

ggplot(data = growth.rates.long3, aes(x = exp_period, y = growthrate))+ geom_point()+
  stat_summary( fun = "mean", geom = "point", size = 3, position = position_dodge(width = 0.1))+ 
  stat_summary( fun = "mean", geom = "line", size = 1, position = position_dodge(width = 0.1), linetype = "dashed", aes(group = wound))+ 
  facet_wrap(~temp)+ 
  theme_classic()+
  ggtitle("Growth Rate of non-wounded corals standardized by final SA (mg/cm2/day)")

```

