---
title: "ACR Wound Closure Rate"
author: "Ninah Munk"
date: "2024-03-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(janitor)
library(ggplot2)
library(stringr)
library(lme4)
library(emmeans)
library(car)
library(ggpubr)
library(rstatix)
library(lmerTest)


```

# DARKHOURS: Load in the data + make some basic plots
```{r}
data<- read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Data/darkhours/wound_closure_measurements.csv")%>%clean_names()

data$time_point_initial[data$time_point_initial == "say14"] <- "day14"


ggplot(data = data, aes(x = area_mm2, y = perimeter))+geom_point()+ggtitle("Fig. 1 Wound Perimeter x Area")

ggplot(data = data, aes(x = area_mm2, y = ferets_diameter_mm))+geom_point()+ggtitle("Fig 2. Wound Diameter x Area")

#divide perimeter by area to get ratio P:A
data<- data%>%select(-notes)%>%mutate(P.A = (perimeter / area_mm2))%>%mutate(time_point_initial = as.factor(time_point_initial))

#desired_order<- c("initial", "day7", "day14", "final")
data$time_point_initial <- factor(data$time_point_initial, levels = c("initial", "day7", "day14", "final"))

ggplot(data = data, aes(x = time_point_initial, y = P.A))+geom_boxplot()+ggtitle("Fig. 3 Perimeter:Area Ratio - P:R increasing")


```
Fig 1. Wound area and perimeter increase together. Fig 2. Wound diameter and area increase together. Fig 3. P:A ratio gets larger as wound heals.
Wounds that are larger have lower P:A ratios because there is more surface area than perimeter. In the literature, wounds with low P:A ratio take longer to heal, which supports the hypothesis that resources from tissue directly boarder the lesion are responsible for wound repair AND/OR there is a limited pot of energy available for repair which large wounds exhaust AND/OR repair becomes too expensive and the colony stops fueling repair   

## DARKHOURS:load in sample info and join with data
```{r}
samp.info<- read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Data/darkhours/darkhours_mastersheetcsv.csv")

data2<- left_join(data, samp.info, by = "coral_id")

```



# DARKHOURS:Area 
```{r}
ggplot(data = data2, aes(x = time_point_initial, y = area_mm2))+geom_boxplot()+ggtitle("Fig. 1 Wound Area through time")+facet_wrap(~temp)

```
Fig. 1 Wound area gets smaller through time (ambient n=5, hot n=5)

# DARKHOURS:Presence of Apical Polyps 
```{r}
data2 <- data2 %>% mutate(apical_polyp_tentacles = str_squish(apical_polyp_tentacles)) 

ggplot(data = data2, aes(x = apical_polyp_tentacles))+geom_bar(aes(fill = temp))+facet_wrap(~time_point_initial, ncol = 4)+ ggtitle("Fig 1. Presence of Apical Polyp")

data2%>%filter(time_point_initial == "day7")-> filtered

ggplot(data = filtered, aes(x =apical_polyp_tentacles))+geom_bar()+ggtitle("Fig 2. Day 7")+facet_wrap(~temp)
```
Fig. 1 Apical polyp is at the very tip of the coral. This figure shows at each time point if that polyp was present for each coral in both temp treatments. Fig. 2 Apical polyp is visible for 2 corals at day 7 in the ambient treatment only.

# DARKHOURS: Number of new polyps 
```{r}
ggplot(data = data2, aes(x = time_point_initial, y = num_new_polyps))+geom_point(aes(col = as.factor(coral_id)))+facet_wrap(~temp)+ggtitle("Fig 1. New polyps for each coral throughout experiment") #facet wrap as individual 

ggplot(data = data2, aes(x = time_point_initial, y = num_new_polyps))+geom_boxplot(aes(fill = temp))+facet_wrap(~temp)+ggtitle("Fig 2. Number of new polyps throughout experiment")

data2%>%filter(time_point_initial == "final") -> final.timepoint

ggplot(data = final.timepoint, aes(x = temp, y = num_new_polyps))+ geom_boxplot(aes(fill = temp))+ggtitle("Fig 3. Number of new polyps formed between day 14 and final")

# Get total num of new polyps
# Group by coral_id and summarize to get the sum of num_new_polyps for each coral_id
sum_polyps <- data2 %>%
  group_by(coral_id) %>%
  summarize(total_polyps = sum(num_new_polyps, na.rm = TRUE))

# Join the sum_polyps data frame with the original data frame based on coral_id
data2 <- left_join(data2, sum_polyps, by = "coral_id")

ggplot(data = data2, aes(x = temp, y = total_polyps))+geom_boxplot(aes(fill = temp))+ggtitle("Fig 4. total number of new polyps by end of experiment")
```
New polyps formed across 20 days post fragmentation wound. By the end of experiment corals in both treatments had similar number of new polyps formed (Fig. 3), but corals in the hot treatment had delayed polyp formation at beginning OR enhanced polyp formation towards end. 

# Wound healing analysis of main ACR regen project - FRAG ONLY

```{r}
data3<- read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Data/fragment.only.measurements.csv")%>%clean_names()

data3$time_point_initial <- factor(data3$time_point_initial, levels = c("initial", "24hr", "day10", "final"), labels = c("0", "1", "10", "19")) 

ggplot(data = data3, aes(x = area_mm2, y = perimeter))+geom_point()+ggtitle("Fig. 1 Wound Perimeter x Area")

ggplot(data = data3, aes(x = area_mm2, y = perimeter, col = time_point_initial))+geom_point()+ggtitle("Fig 2. perimeter x area")+geom_smooth(method = "lm")

ggplot(data = data3, aes(x = area_mm2, y = ferets_diameter_mm))+geom_point()+ggtitle("Fig 3. Wound Diameter x Area")

data3<- data3%>%select(-notes)%>%mutate(P.A = (perimeter / area_mm2))%>%mutate(time_point_initial = as.factor(time_point_initial))

ggplot(data = data3, aes(x = time_point_initial, y = P.A))+geom_point()+ggtitle("Fig 4. Perimeter:Area Ratio")


wound.measurements.model<- lm(area_mm2 ~ perimeter + time_point_initial, data = data3)
summary(wound.measurements.model)
pairwise_comparisons <- emmeans(wound.measurements.model, pairwise ~ time_point_initial, adjust = "tukey")
summary(pairwise_comparisons)


```
# Wound healing analysis of main ACR regen project - FRAG ONLY cont.
```{r}
samp.info2<- read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Respiration/Data/samp_info.csv")
data4<- left_join(data3, samp.info2, by = "coral_id")

ggplot(data = data4, aes(x = time_point_initial, y = area_mm2))+geom_boxplot()+facet_wrap(~temp)+ggtitle("Wound Closure (area), Frag only")

ggplot(data = data4, aes(x = time_point_initial, y = perimeter))+geom_boxplot()+facet_wrap(~temp)+ggtitle("Wound Closure (perimeter), Frag only")

ggplot(data = data4, aes(x = area_mm2, y = perimeter, col = time_point_initial))+geom_point()+ggtitle("perimeter x area")+geom_smooth(method = "lm")+ facet_wrap(~temp)

ggplot(data = data4, aes(x = area_mm2, y = ferets_diameter_mm, col = time_point_initial))+geom_point()+ggtitle("diameter x area")+geom_smooth(method = "lm")+ facet_wrap(~temp)

ggplot(data = data4, aes(x = area_mm2, y = perimeter, col = temp))+geom_point()+ggtitle("perimeter x area")+geom_smooth(method = "lm")+ facet_wrap(~time_point_initial)

ggplot(data = data4, aes(x = area_mm2, y = ferets_diameter_mm, col = temp))+geom_point()+ggtitle("diameter x area")+geom_smooth(method = "lm")+ facet_wrap(~time_point_initial)





```
```{r}
# Fragment Wound healing statstics of main ACR regen project - FRAG ONLY cont. 

# Convert 'time_point_initial' and 'coral_id' to factors (random factors)
data4$time_point_initial <- as.factor(data4$time_point_initial)
data4$coral_id <- as.factor(data4$coral_id)
data4$temp <- as.factor(data4$temp)

# Get levels of grouping variable 
levels(data4$temp)

# summary statistics
data4 %>%
  group_by(temp) %>%
  get_summary_stats(area_mm2, type = "mean_sd")

# visualize 
bxp <- ggboxplot(data4, x = "temp", y = "area_mm2")
bxp

#check outliers
data4 %>%
  group_by(temp) %>%
  identify_outliers(area_mm2) # no outliers

# check normality of all groups together
model<- lm(area_mm2 ~ temp, data = data4)

ggqqplot(residuals(model))

shapiro_test(residuals(model))#data is normal p> 0.05  p> 0.05

#check normality within groups 
data4 %>%
  group_by(temp) %>%
  shapiro_test(area_mm2) #data is normal p> 0.05 

# QQ plot within groups 
ggqqplot(data4, "area_mm2", facet.by = "temp")

#check homogeneity of variance 
plot(model, 1) #no relationship between residuals and fitted values

data4%>%levene_test(area_mm2 ~ temp) # p> 0.05, assumption met, no sig difference in variance across groups 

# computation of one way ANOVA
res.aov <- data4%>%anova_test(area_mm2 ~ temp)

get_anova_table(res.aov) 
# p = 0.278, not significant 
# ges (effect size) 2.6% of the area differences can be explained by temperature 



```

## repeated measures one way ANOVA 
```{r}
# Fragment Wound healing statstics of main ACR regen project - FRAG ONLY cont. 
# Get levels of grouping variable 
levels(data4$time_point_initial)

# summary statistics
data4 %>%
  group_by(time_point_initial) %>%
  get_summary_stats(area_mm2, type = "mean_sd")

# visualize 
bxp <- ggboxplot(data4, x = "time_point_initial", y = "area_mm2", add = "point")
bxp

#check outliers
data4 %>%
  group_by(time_point_initial) %>%
  identify_outliers(area_mm2) # three outliers all coral 80, only timepoint 1 is extreme

#check normality within groups 
data4 %>%
  group_by(time_point_initial) %>%
  shapiro_test(area_mm2) #data is not normal for timepoints 1 and 19, p < 0.05 

# QQ plot within groups 
ggqqplot(data4, "area_mm2", facet.by = "time_point_initial")

res.aov <- anova_test(data = data4, dv = area_mm2, wid = coral_id, within = time_point_initial)
get_anova_table(res.aov)

#post-hoc tests, pairwise comparisons (t.tests between lebels of the within subjects factor (i.e., time))

pwc <- data4 %>%
  pairwise_t_test(
    area_mm2 ~ time_point_initial, paired = TRUE,
    p.adjust.method = "bonferroni"
    )
pwc

# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "time")
bxp + 
  stat_pvalue_manual(pwc) +
  labs(
    subtitle = get_test_label(res.aov, detailed = TRUE),
    caption = get_pwc_label(pwc)
  )

```


```{r}
# since normality assumption is broken for one-way anova, here I am trying the Friedman test

# get summary statistics
data4 %>%
  group_by(time_point_initial) %>%
  get_summary_stats(area_mm2, type = "common")

# visualize 
ggboxplot(data4, x = "time_point_initial", y = "area_mm2", add = "jitter")

# computation, non-parametric friedman test 
res.fried<- data4%>%friedman_test(area_mm2 ~ time_point_initial | coral_id)
res.fried # area is significantly different at the different time points p <0.05 

#effect size 
data4%>%friedman_effsize(area_mm2 ~ time_point_initial | coral_id) # W = 0.65 large effect size according to Cohen's interpretation guidelines (>=0.5 is large) W = 0.65

# pairwise comparisons
pwc <- data4 %>%
  wilcox_test(area_mm2 ~ time_point_initial, paired = TRUE, p.adjust.method = "bonferroni")
pwc

# Visualization: box plots with p-values
pwc <- pwc %>% add_xy_position(x = "time")
ggboxplot(data4, x = "time_point_initial", y = "area_mm2", add = "point") +
  stat_pvalue_manual(pwc, hide.ns = TRUE) +
  labs(
    subtitle = get_test_label(res.fried,  detailed = TRUE),
    caption = get_pwc_label(pwc)
  )
```

REPEATED MEASURES TWO WAY ANOVA --> wont run because area has 0 values and the test expects them to be NA 
# res.aov <- anova_test(
#   data = data4, dv = area_mm2, wid = coral_id,
#   within = c(temp, time_point_initial)
#   ) 
# get_anova_table(res.aov)

## Two Way MIXED ANOVA
```{r}
# get summary statistics 
data4 %>%
  group_by(temp, time_point_initial) %>%
  get_summary_stats(area_mm2, type = "mean_sd")

# visualization
bxp <- ggboxplot(
  data4, x = "time_point_initial", y = "area_mm2",
  color = "temp", palette = "jco"
  )
bxp

# check outliers
data4 %>%
  group_by(temp, time_point_initial) %>%
  identify_outliers(area_mm2) # two extreme outliers, both coral 80

# normality
data4 %>%
  group_by(temp, time_point_initial) %>%
  shapiro_test(area_mm2) # some groups not normal

ggqqplot(data4, "area_mm2", ggtheme = theme_bw()) +
  facet_grid(time_point_initial ~ temp, labeller = "label_both")

# homogeneity of variance 
data4%>%
  group_by(time_point_initial)%>%
  levene_test(area_mm2 ~ temp) # p>0.5

# homogeneity of covariances
box_m(data4[, "area_mm2", drop = FALSE], data4$temp) #p> 0.001

res.aov <- anova_test(
  data = data4, dv = area_mm2, wid = coral_id,
  between = temp, within = time_point_initial
  )
get_anova_table(res.aov)



```
Okay.. the data is not normally distributed, theres outliers, and theres cases of 0 values. I am going to use a linear mixed effects model which i think is robust to these things 
```{r}
# Fit a linear mixed-effects model
model1 <- lmer(area_mm2 ~ temp + (1|time_point_initial), data = data4)
model2 <- lmer(area_mm2 ~ temp + (1|time_point_initial)  + (1|coral_id), data = data4)
model3 <- lmer(area_mm2 ~ temp*time_point_initial + (1|coral_id), data = data4) # lowest AIC

AIC(model1, model2, model3)
# Check the model summary
summary(model3)
```

## Calculating effect sizes (Cohen's D)
```{r}
# Extract fixed effects coefficients from the lmer model above
fixed_effects <- fixef(model3)

# Extract standard errors of fixed effects coefficients
se_fixed_effects <- sqrt(diag(vcov(model3)))

# Calculate Cohen's d for each fixed effect
cohen_d <- fixed_effects / se_fixed_effects

# Print Cohen's d for each fixed effect
print(cohen_d)
```

```{r}
install.packages("compute.es")
library(compute.es)

# Compute means, standard deviations, and sample sizes for each group
means <- tapply(data4$area_mm2, list(data4$temp, data4$time_point_initial), mean)
sds <- tapply(data4$area_mm2, list(data4$temp, df$time_point_initial), sd)
ns <- tapply(data4$area_mm2, list(df$temp, df$time_point_initial), length)

# Convert to vectors
means_vec <- as.vector(means)
sds_vec <- as.vector(sds)
ns_vec <- as.vector(ns)

# Calculate Cohen's d with confidence intervals
cohen_d_result <- esci(means1 = means_vec[1:16], means2 = means_vec[17:32], 
                       sd1 = sds_vec[1:16], sd2 = sds_vec[17:32], 
                       n1 = ns_vec[1:16], n2 = ns_vec[17:32])

# Print the result
print(cohen_d_result)
```


# calculating percent wound closure
```{r}
data4%>%filter(temp == "A") -> data4.ambient
data4%>%filter(temp == "H") -> data4.hot

data4.ambient%>%filter(time_point_initial == 0)%>%select(coral_id, area_mm2)%>%rename(initial.area = area_mm2) -> ambient.initial
ambient.initial%>%left_join(data4.ambient, by = "coral_id") -> data4.ambient

data4.ambient%>%mutate(percent.wound.closure = ((area_mm2 - initial.area) / initial.area)*100) -> data4.ambient

data4.hot%>%filter(time_point_initial == 0)%>%select(coral_id, area_mm2)%>%rename(initial.area = area_mm2) -> hot.initial
hot.initial%>%left_join(data4.hot, by = "coral_id") -> data4.hot
data4.hot%>%mutate(percent.wound.closure = ((area_mm2 - initial.area) / initial.area)*100) -> data4.hot
```


## making tables of the wound measurement metrics
```{r}

# Aggregate the data by time_point_initial and calculate mean and standard deviation for each variable
summary_table_ambient <- aggregate(. ~ time_point_initial, data = data4.ambient[, c("area_mm2", "perimeter", "ferets_diameter_mm", "time_point_initial", "percent.wound.closure")],
                           function(x) c(mean = mean(x, na.rm = TRUE), sd = sd(x, na.rm = TRUE)))

# Flatten the resulting data frame for better readability
summary_table_ambient <- do.call(data.frame, summary_table_ambient)

# Rename the columns for better clarity
colnames(summary_table_ambient) <- c("Day", 
                             "Mean Area", "Area SD", 
                             "Mean Perimeter", "Perimeter SD", 
                             "Mean Ferets Diameter", "Ferets Diameter SD", "Mean % Wound Closure", "Mean % Wound Closure SD")

# Display the summary table
print(summary_table_ambient)

# Export the summary table to a CSV file
#write.csv(summary_table_ambient, "/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Output/summary_table_ambient.csv", row.names = FALSE)

summary_table_hot <- aggregate(. ~ time_point_initial, data = data4.hot[, c("area_mm2", "perimeter", "ferets_diameter_mm", "time_point_initial", "percent.wound.closure")],
                           function(x) c(mean = mean(x, na.rm = TRUE), sd = sd(x, na.rm = TRUE)))

# Flatten the resulting data frame for better readability
summary_table_hot <- do.call(data.frame, summary_table_hot)

# Rename the columns for better clarity
colnames(summary_table_hot) <- c("Day", 
                             "Mean Area", "Area SD", 
                             "Mean Perimeter", "Perimeter SD", 
                             "Mean Ferets Diameter", "Ferets Diameter SD", "Mean % Wound Closure", "Mean % Wound Closure SD")

# Display the summary table
print(summary_table_hot)

# Export the summary table to a CSV file
#write.csv(summary_table_hot, "/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Output/summary_table_hot.csv", row.names = FALSE)


```

```{r}
rbind(data4.ambient, data4.hot)-> test.means

# Define a function to compute mean and standard error
mean_se <- function(x) {
  n <- length(x)
  mean_val <- mean(x, na.rm = TRUE)
  se_val <- sd(x, na.rm = TRUE) / sqrt(n)
  c(mean = mean_val, se = se_val)
}

# Aggregate the data by time_point_initial and temp and calculate mean and standard deviation for each variable
summary_table <- aggregate(. ~ time_point_initial + temp, 
                                  data = test.means[, c("area_mm2", "perimeter", "ferets_diameter_mm", "percent.wound.closure", "time_point_initial", "temp")],
                                  mean_se)

# Flatten the resulting data frame for better readability
summary_table <- do.call(data.frame, summary_table)

# Rename the columns for better clarity
colnames(summary_table) <- c("Day", "Temp",
                                     "Area", "Area SE", 
                                     "Perimeter", "Perimeter SE", 
                                     "Ferets Diameter", "Ferets Diameter SE", 
                                     "Wound Closure (%)", "Wound Closure (%) SE")
print(summary_table)
```

```{r}
summary_table%>%clean_names()%>%mutate(wound_closure_percent = abs(wound_closure_percent))-> summary_table
ggplot(data = summary_table, aes(x = day, y = wound_closure_percent, color = temp))+ geom_point()

# Create the ggplot object
ggplot(summary_table, aes(x = day, y = wound_closure_percent, fill = temp)) +
  # Add the bar plot
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), width = 0.6) +
  # Add the error bars
  geom_errorbar(aes(ymin = wound_closure_percent - wound_closure_percent_se,
                    ymax = wound_closure_percent + wound_closure_percent_se),
                position = position_dodge(width = 0.9), width = 0.2) +
  # Customize plot aesthetics
  labs(x = "Day", y = "Wound Closure (%)", fill = "Temperature") +
  theme_minimal() +
  theme(legend.position = "top") +
  scale_fill_manual(values = c("A" = "blue", "H" = "red"))  # Set custom fill colors for temperature treatments
```

# Darkhours Wound Closure
```{r}

darkdata<- read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Data/darkhours/wound_closure_measurements.csv")%>%clean_names()

darkdata$time_point_initial <- factor(darkdata$time_point_initial, levels = c("initial", "day7", "day14", "final"), labels = c("0", "7", "14", "20")) 
darkdata$time_point_initial[is.na(darkdata$time_point_initial)] <- "14"

darkdata%>%select(-num_new_polyps, -apical_polyp_tentacles, -notes)%>%left_join(samp.info, by = "coral_id")-> darkdata.closure

darkdata.closure%>%filter(temp == "A") -> darkdata.ambient
darkdata.closure%>%filter(temp == "H") -> darkdata.hot

darkdata.ambient%>%filter(time_point_initial == 0)%>%select(coral_id, area_mm2)%>%rename(initial.area = area_mm2) -> dark.ambient.initial
dark.ambient.initial%>%left_join(darkdata.ambient, by = "coral_id") -> darkdata.ambient
darkdata.ambient%>%mutate(percent.wound.closure = ((area_mm2 - initial.area) / initial.area)*100) -> darkdata.ambient

darkdata.hot%>%filter(time_point_initial == 0)%>%select(coral_id, area_mm2)%>%rename(initial.area = area_mm2) -> dark.hot.initial
dark.hot.initial%>%left_join(darkdata.hot, by = "coral_id") -> darkdata.hot
darkdata.hot%>%mutate(percent.wound.closure = ((area_mm2 - initial.area) / initial.area)*100) -> darkdata.hot

rbind(darkdata.ambient, darkdata.hot) -> darkdata.means

# Aggregate the data by time_point_initial and temp and calculate mean and standard deviation for each variable
darkdata_summary_table <- aggregate(. ~ time_point_initial + temp, 
                                  data = darkdata.means[, c("area_mm2", "perimeter", "ferets_diameter_mm", "percent.wound.closure", "time_point_initial", "temp")],
                                  mean_se)

# Flatten the resulting data frame for better readability
darkdata_summary_table <- do.call(data.frame, darkdata_summary_table)

# Rename the columns for better clarity
colnames(darkdata_summary_table) <- c("Day", "Temp",
                                     "Area", "Area SE", 
                                     "Perimeter", "Perimeter SE", 
                                     "Ferets Diameter", "Ferets Diameter SE", 
                                     "Wound Closure (%)", "Wound Closure (%) SE")
print(darkdata_summary_table)

darkdata_summary_table%>%clean_names()%>%mutate(wound_closure_percent = abs(wound_closure_percent))-> darkdata_summary_table
ggplot(data = darkdata_summary_table, aes(x = day, y = wound_closure_percent, color = temp))+ geom_point()

# Create the ggplot object
ggplot(darkdata_summary_table, aes(x = day, y = wound_closure_percent, fill = temp)) +
  # Add the bar plot
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), width = 0.6) +
  # Add the error bars
  geom_errorbar(aes(ymin = wound_closure_percent - wound_closure_percent_se,
                    ymax = wound_closure_percent + wound_closure_percent_se),
                position = position_dodge(width = 0.9), width = 0.2) +
  # Customize plot aesthetics
  labs(x = "Day", y = "Wound Closure (%)", fill = "Temperature") +
  theme_minimal() +
  theme(legend.position = "top") +
  scale_fill_manual(values = c("A" = "blue", "H" = "red"))

```
# Models for Dark Hours fragment wound closure
```{r}
darkdata.closure%>%mutate(temp = as.factor(temp))%>%mutate(coral_id = as.factor(coral_id))-> darkdata.closure
# Fit a linear mixed-effects model
model4 <- lmer(area_mm2 ~ temp + (1|time_point_initial), data = darkdata.closure)
model5 <- lmer(area_mm2 ~ temp + (1|time_point_initial)  + (1|coral_id), data = darkdata.closure)
model6 <- lmer(area_mm2 ~ temp*time_point_initial + (1|coral_id), data = darkdata.closure) # lowest AIC

AIC(model4, model5, model6)
# Check the model summary
summary(model6)
```


```{r}

# T-test to test compare mean % wound closure between ambient and hot at each timepoint  
test.means%>%filter(time_point_initial == 1)%>%select(coral_id,temp, percent.wound.closure)%>%pivot_wider(names_from = "temp", values_from = "percent.wound.closure")-> t1

test.t1 <- t.test(x = t1$A,
                       y = t1$H,
                       alternative = "two.sided",
                       mu = 0, paired = FALSE, var.equal = FALSE,
                      conf.level = 0.95)
# Print results
print(test.t1)


test.means%>%filter(time_point_initial == 10)%>%select(coral_id,temp, percent.wound.closure)%>%pivot_wider(names_from = "temp", values_from = "percent.wound.closure")-> t2

test.t2 <- t.test(x = t2$A,
                       y = t2$H,
                       alternative = "two.sided",
                       mu = 0, paired = FALSE, var.equal = FALSE,
                      conf.level = 0.95)
# Print results
print(test.t2)


test.means%>%filter(time_point_initial == 19)%>%select(coral_id,temp, percent.wound.closure)%>%pivot_wider(names_from = "temp", values_from = "percent.wound.closure")-> t3

test.t3 <- t.test(x = t3$A,
                       y = t3$H,
                       alternative = "two.sided",
                       mu = 0, paired = FALSE, var.equal = FALSE,
                      conf.level = 0.95)
# Print results
print(test.t3)
```


```{r}
data4 <- data4 %>% mutate(apical_polyp_tentacles = str_squish(apical_polyp_tentacles))

ggplot(data = data4, aes(x = apical_polyp_tentacles))+geom_bar(aes(fill = temp))+facet_wrap(~time_point_initial)+ggtitle("Presence of Apical Polyp Tentacles")

data4%>%filter(time_point_initial == "10")-> filtereddata

ggplot(data = filtereddata, aes(x =apical_polyp_tentacles))+geom_bar()+ggtitle("Presence of Apical Polyp Tentacles - Day 10")+facet_wrap(~temp)
```
# chisquared
# number of new polyps
```{r}
ggplot(data = data4, aes(x = num_new_polyps))+geom_bar(aes(fill = temp))+facet_wrap(~time_point_initial, ncol= 4)+ggtitle("Number of new polyps")

ggplot(data = data4, aes(x = time_point_initial, y = num_new_polyps))+geom_boxplot()+facet_wrap(~temp)+ggtitle("Number of new polyps")
```
# MAIN PROJ: Fragment and Abrasion analysis
```{r}
read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Data/all.wounds.csv")%>%left_join(samp.info2, "coral_id")-> data5
read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Data/all.wounds.csv") -> data5
read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Surface_Area/Output/post_wound_surface_areas.csv") -> SA

data5%>%left_join(SA, "coral_id")-> data5
data5[data5 == "no"] <- "0"
data5[data5 == "no "] <- "0"
data5[data5 == "yes"] <- "1"
data5[data5 == " yes"] <- "1"
data5[data5 == "nyes"] <- "1"
data5[data5 == "initial"] <- "0"
data5[data5 == "24hr"] <- "1"
data5[data5 == "day10"] <- "10"
data5[data5 == "final"] <- "19"

data5%>%mutate(radial_coralite_structure = as.factor(radial_coralite_structure), 
               radial_coralite_polyps = as.factor(radial_coralite_polyps),
               apical_coralite.structure = as.factor(apical_coralite.structure),
               healed = as.factor(healed),
               tissue_over_wound = as.factor(apical_coralite.structure),
               wound = as.factor(wound),
               temp = as.factor(temp),
               genotype = as.factor(genotype))%>%select(-wound_type)-> data5

ggplot(data = data5%>%filter(time_point == "10"), aes(x = SA_post_wound, y = healed, col = temp))+ geom_point()+facet_wrap(~wound)+ggtitle("Healed status Day 10")+theme_classic()
ggplot(data = data5%>%filter(time_point == "10"), aes(x = wound, fill = healed))+ geom_bar() +facet_wrap(~temp)



```
# binomial logistic regression

# MAIN PROJ: stats for day 10 healing results 
```{r}
m1<- glm(healed ~ temp*wound, data = data5%>%filter(time_point == "10"), family = binomial)
summary(m1)
m2<- glm(healed ~ temp + wound, data = data5%>%filter(time_point == "10"), family = binomial)
summary(m2)


AIC(m1, m2) # only thing that matters is which one is lower and to be actually lower, needs to be a difference of 2. can see model 2 is better 

emmeans_results <- emmeans(m2, ~ wound)
# Display the EMMs
print(emmeans_results)
# Pairwise comparisons for the levels within each predictor
pairwise_results <- pairs(emmeans_results)
# Display the pairwise comparisons
print(pairwise_results)

```
# how many coral were fully healed based on signs of healing 
```{r}
#view(data5%>%filter(time_point == 19, wound == 1)) # 12/12 = 100% of fragmented corals
#view(data5%>%filter(time_point == 19, wound == 2)) # 8/12 = 67% of abrasion corals
```

# Shane helps w survival analysis 
```{r}
library(ggfortify)
library(survival)
library(gtsummary)
#clean data frame for columns we want, change timepoint and healed to numeric for model
data5%>%select(coral_id, time_point, healed, wound, temp)%>%group_by(coral_id)%>%filter(lag(healed)!=1)%>%ungroup()%>%filter(healed == 1 | time_point == 19) %>% mutate(time_point = as.numeric(time_point), healed = as.character(healed), healed = as.numeric(healed))-> survival
# survfit is model function (like  a glm), surv() function to create dependent df (when the event happened (time) and what was event (event)) as a function of wound
survfit(Surv(time = time_point, event = healed)~ wound, data = survival) #can see median time to healing is lower for wound 1, but cant say that is sig bc no upper CI 
summary(survfit(Surv(time = time_point, event = healed)~ wound, data = survival))
# put model inside autoplot() to visualize, fun = event to switch direction of healing 
autoplot(survfit(Surv(time = time_point, event = healed)~ wound, data = survival), fun = 'event') -> p1
# cox.zph() assumption check to see if hazard ratio is changing over time (p<0.05 is bad)
cox.zph(coxph(Surv(time_point, healed)~wound + temp, data = survival))
# harard ratio, CI, and p value for wound 2 with respect to wound 1. HR<1, CI not crossing 1, p<0.05 all mean that at any point in time individuals with wound 2 have less probability to be healed 
coxph(Surv(time_point, healed)~wound + temp, data = survival) %>% tbl_regression(exp = T)

p1+ theme_classic()+ xlab("Time (days")+ ylab("Survival (%)")+
  guides(fill=FALSE) +
  labs(colour = "Wound Type") +
  scale_color_manual(labels = c("Fragmentation", "Abrasion"), values = c(1, 2)) 
```


```{r}
data5%>%mutate(radial_coralite_structure = as.character(radial_coralite_structure),
               radial_coralite_polyps = as.character(radial_coralite_polyps),
               apical_coralite.structure = as.character(apical_coralite.structure))%>%
  mutate(radial_coralite_structure = as.numeric(radial_coralite_structure),
               radial_coralite_polyps = as.numeric(radial_coralite_polyps),
               apical_coralite.structure = as.numeric(apical_coralite.structure))-> data5

data5%>%mutate(percent_healed = (radial_coralite_structure + radial_coralite_polyps + apical_coralite.structure) / 3) -> data5

data5%>%select(coral_id, time_point, healed, percent_healed)-> percent.healed
```


