---
title: "regeneration.20240524"
author: "Ninah Munk"
date: "2024-05-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggfortify)
library(survival)
library(gtsummary)
library(rstatix)
library(Rmisc)
library(lmerTest)
library(tidyverse)
library(janitor)
library(emmeans)
library(plyr)
library(ggthemes)
library(survminer)
#library(survMisc)
```


# Load data and combine 
```{r}
read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Data/all.wounds.csv")%>%select(-wound_type) -> regen
read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Respiration/Data/samp_info.csv")-> samp.info
read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Surface_Area/Output/post_wound_surface_areas.csv") -> SA
read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Data/worm_damage.csv")%>%select(coral_id, affected, percent_affected)-> worm

regen%>%left_join(samp.info, by = "coral_id")%>%left_join(SA, by = "coral_id")%>%left_join(worm, by = "coral_id")%>%select(-X) -> data

```


# modify data set 
```{r}

data[data == "no"] <- "0"
data[data == "no "] <- "0"
data[data == "yes"] <- "1"
data[data == " yes"] <- "1"
data[data == "nyes"] <- "1"
data[data == "initial"] <- "0"
data[data == "24hr"] <- "1"
data[data == "day10"] <- "10"
data[data == "final"] <- "19"

data%>%mutate(radial_coralite_structure = as.factor(radial_coralite_structure), 
               radial_coralite_polyps = as.factor(radial_coralite_polyps),
               apical_coralite.structure = as.factor(apical_coralite.structure),
               healed = as.factor(healed),
               tissue_over_wound = as.factor(apical_coralite.structure),
               wound = as.factor(wound),
               temp = as.factor(temp),
               genotype = as.factor(genotype))-> data
```


# FILTERING DATA TO REMOVE WORM EFFECT
Here I filter the data to use only the corals that had low worm damage. This threshold is <= 10% worm damage and was determined using the growth rate dataset. The remaining sample size is N = 16
```{r}
data = subset(data, percent_affected <= 10)
```

# Clean Data
```{r}
#clean data
corals_to_remove <- c(11, 5, 37, 104, 71, 17, 29, 100)
data <- data[!(data$coral_id %in% corals_to_remove), ]
```

# HOW MANY OF THE REMAINING CORALS DO I ALSO HAVE REGENERATION DATA FOR? 
Out of the 71 corals I have growth data on, I have regeneration data for 14 corals. (16 if i dont clean the dataset)
```{r}
print(data %>% group_by(coral_id) %>% reframe(samp.size=n()))
print(data %>% group_by(time_point, wound) %>% reframe(samp.size=n()))
```

# survival and cox models
```{r}

#clean data frame for columns we want, change timepoint and healed to numeric for model
data%>%select(coral_id, time_point, healed, wound, temp, genotype, percent_affected)%>%group_by(coral_id)%>%filter(lag(healed)!=1)%>%ungroup()%>%filter(healed == 1 | time_point == 19) %>% mutate(time_point = as.numeric(time_point), healed = as.character(healed), healed = as.numeric(healed))-> survival
# survfit is model function (like  a glm), surv() function to create dependent df (when the event happened (time) and what was event (event)) as a function of wound
surv.mod<- survfit(Surv(time = time_point, event = healed)~ temp, data = survival) #can see median time to healing is lower for wound 1, but cant say that is sig bc no upper CI 
summary(survfit(Surv(time = time_point, event = healed)~ temp, data = survival))
# put model inside autoplot() to visualize, fun = event to switch direction of healing 
autoplot(survfit(Surv(time = time_point, event = healed)~ temp, data = survival), fun = 'event') -> p1
# cox.zph() assumption check to see if hazard ratio is changing over time (p<0.05 is bad)
cox.zph(coxph(Surv(time_point, healed)~  temp, data = survival))
# harard ratio, CI, and p value for wound 2 with respect to wound 1. HR<1, CI not crossing 1, p<0.05 all mean that at any point in time individuals with wound 2 have less probability to be healed 
coxph(Surv(time_point, healed)~temp, data = survival) %>% tbl_regression(exp = T)

survdiff(Surv(time_point, healed) ~ temp, data = survival) # ask about this 

p1+xlab("Time (days")+ ylab("Survival (%)")+
  guides(fill="none") +
  labs(colour = "Injury Type") +
  scale_colour_brewer(type = "qual", palette = "Dark2",
                    guide = guide_legend(keywidth = 3, keyheight = 3), labels = c("Fragmentation", "Abrasion"))+theme_classic()

```

# glm

```{r}
m1<- glm(healed ~ wound*temp*time_point, data = survival, family = binomial)
summary(m1)
```

# how many corals healed by the end of experiment?
```{r}
data%>%filter(time_point == 19, wound == 1) # 7/7 = 100% of fragmented corals
data%>%filter(time_point == 19, wound == 2) # 6/7 
```

```{r}
survival%>%unite("treatment", wound:temp, remove = FALSE) -> survival

```

For censored data (where the event has not occurred by the end of the study or observation period), the event status is typically represented as Surv(time, event) where event is 0 (indicating censored) or 1 (indicating the event occurred).
For uncensored data (where the event has occurred), event is typically represented as 1.
```{r}

# survfit is model function (like  a glm), surv() function to create dependent df (when the event happened (time) and what was event (event)) as a function of wound
surv.mod<- survfit(Surv(time = time_point, event = healed) ~ wound + temp, data = survival) #can see median time to healing is lower for wound 1, but cant say that is sig bc no upper CI 
summary(survfit(Surv(time = time_point, event = healed)~ wound + temp, data = survival))
# put model inside autoplot() to visualize, fun = event to switch direction of healing 
autoplot(survfit(Surv(time = time_point, event = healed) ~ wound + temp, data = survival), fun = 'event') 
# cox.zph() assumption check to see if hazard ratio is changing over time (p<0.05 is bad)
cox.zph(coxph(Surv(time_point, healed)~ wound * temp, data = survival, id = coral_id))
# harard ratio, CI, and p value for wound 2 with respect to wound 1. HR<1, CI not crossing 1, p<0.05 all mean that at any point in time individuals with wound 2 have less probability to be healed 
coxph(Surv(time_point, healed)~wound + temp, data = survival, id = coral_id) %>% tbl_regression(exp = T)

p1+xlab("Time (days")+ ylab("Survival (%)")+
  guides(fill="none") +
  labs(colour = "Injury Type") +
  scale_colour_brewer(type = "qual", palette = "Dark2",
                    guide = guide_legend(keywidth = 3, keyheight = 3), labels = c("Fragmentation", "Abrasion"))+theme_classic()

```

```{r}
library(coxme)

coxme(Surv(time_point, healed)~wound*temp + (1|percent_affected), data = survival) %>% tbl_regression(exp = T)

coxme(Surv(time_point, healed)~treatment+ (1|genotype/cora) + (1|percent_affected), data = survival) %>% tbl_regression(exp = T)
cox.zph(coxme(Surv(time_point, healed)~treatment+ (1|percent_affected), data = survival))
```

```{r}

# Plot survival curves
ggsurvplot(surv_fit, data = survival, risk.table = TRUE, pval = TRUE, fun = "event")


# Customizing appearance
ggsurvplot(surv_fit, data = survival,
          palette = c("blue", "lightblue", "darksalmon", "red"),   # Custom color palette
           #linetype = c("solid", "dashed", "solid", "dashed"),                  # Different linetypes for groups
           xlab = "Time", ylab = "Probability of full regeneration",  # Axis labels
          # risk.table = "percentage", # Include risk table
          # tables.theme = theme_cleantable(),
           #pval = TRUE,
           fun = "event",
           #conf.int = TRUE,
           legend.title = "Treatments",
 legend.labs = c("FA", "FE", "AA", "AE"),
 #legend = "none",
          # surv.median.line = "v",
          # cumevents = TRUE
         # risk.table.title = " "
           ) 

```

```{r}
surv_fit <- survfit(Surv(time_point, healed) ~ treatment, data = survival)
# Convert survfit object to data frame
surv_data <- summary(surv_fit, times = c(10, 15, 19, 20))  # Specify times for risk table

# Create a data frame from summary
surv_df <- data.frame(
  time = surv_data$time,
  surv = surv_data$surv,
  strata = as.factor(surv_data$strata)
)

# Define color palette and linetypes
palette <- c("blue", "blue", "red", "red")
linetypes <- c("solid", "dashed", "solid", "dashed")

# Plot survival curves with ggplot2
ggplot(surv_df, aes(x = time, y = surv, color = strata, linetype = strata)) +
  geom_step(size = 1.5) +
  scale_color_manual(values = palette) +
  scale_linetype_manual(values = linetypes) +
  labs(title = "Survival Curves by Treatment", x = "Time", y = "Survival Probability") +
  theme_minimal()
```



# using treatment
```{r}
#make wound and temp factors
survival$treatment <- as.factor(survival$treatment)
# create survival object with time and healed data
surv_object <- Surv(time = survival$time_point, event = survival$healed)
# fit survival model
surv.mod <- survfit(surv_object~treatment, data = survival)
# confirm results of survival model with cox model
cox_model <- coxph(surv_object ~ treatment, data = survival)
# results of cox model
summary(cox_model)
# make sure hazard ratio is not changing over time (P>0.5)
cox.zph(cox_model)
# make table of cox results
coxph(surv_object ~ treatment, data = survival) %>% tbl_regression(exp = T)


```
Penalized regression techniques can handle perfect separation by adding a penalty to the model fitting process. Ridge regression and the Lasso (Least Absolute Shrinkage and Selection Operator) are common choices.
```{r}

library(glmnet)

# Prepare the data for glmnet
# glmnet expects the predictors as a matrix and the response as a survival object
x <- model.matrix(healed ~wound * temp, data = survival)[, -1]  # remove the intercept
y <- Surv(survival$time_point, survival$healed)

# Fit the penalized Cox model using glmnet
cox_penalized <- glmnet(x, y, family = "cox", alpha = 1)  # alpha = 0 for Ridge, alpha = 1 for Lasso
print(cox_penalized)

#To find the best lambda, use cross-validation
cv_fit <- cv.glmnet(x, y, family = "cox", alpha = 1)
plot(cv_fit)

# The lambda that gives the minimum mean cross-validated error
best_lambda <- cv_fit$lambda.min
print(best_lambda)

# Fit the final model with the best lambda
cox_lasso_best <- glmnet(x, y, family = "cox", alpha = 1, lambda = best_lambda)
print(cox_lasso_best)


# Extract the coefficients for the model with the best lambda
lasso_coefficients <- coef(cox_lasso_best)
print(lasso_coefficients)

# Convert the coefficients to a data frame and display non-zero coefficients
lasso_coefficients_df <- as.data.frame(as.matrix(lasso_coefficients))
lasso_coefficients_df <- lasso_coefficients_df[lasso_coefficients_df$s1 != 0, , drop = FALSE]
print(lasso_coefficients_df)

# Predict risk scores for each observation
risk_scores <- predict(cox_lasso_best, newx = x, type = "link")
print(risk_scores)

```


