---
title: "Regeneration 6"
author: "Ninah Munk"
date: "2024-07-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggfortify)
library(survival)
library(gtsummary)
library(rstatix)
library(Rmisc)
library(lmerTest)
library(tidyverse)
library(janitor)
library(emmeans)
library(plyr)
library(ggthemes)
library(survminer)
library(coxme)
```


# Load data and combine 
```{r}
read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Data/healing.signs.data.ninah.csv")%>%select(-wound) -> regen
read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Respiration/Data/samp_info.csv")-> samp.info
read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Surface_Area/Output/post_wound_surface_areas.csv") -> SA
read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Data/worm_damage.csv")%>%select(coral_id, affected, percent_affected)-> worm

regen%>%left_join(samp.info, by = "coral_id")%>%left_join(SA, by = "coral_id")%>%left_join(worm, by = "coral_id") -> data

```


# modify data set 
```{r}

data%>%mutate(polyp = as.factor(polyp), 
               corallite = as.factor(corallite),
               new_tissue = as.factor(new_tissue),
               healed.full = as.factor(healed.full),
               healed.2.3 = as.factor(healed.2.3),
               wound = as.factor(wound),
               temp = as.factor(temp),
               genotype = as.factor(genotype))-> data2
```

# Clean Data
```{r}
#clean data
corals_to_remove <- c(11, 5, 37, 104, 71, 17, 29, 100)
data2 <- data2[!(data$coral_id %in% corals_to_remove), ]
```


# FILTERING DATA TO REMOVE WORM EFFECT
Here I filter the data to use only the corals that had low worm damage. This threshold is <= 10% worm damage and was determined using the growth rate dataset. The remaining sample size is N = 16
```{r}
data3 = subset(data2, percent_affected <= 10)

print(data3 %>% group_by(coral_id) %>% reframe(samp.size=n()))
```

# POLYP
survival dataframe
```{r}
#clean data frame for columns we want, change timepoint and healed to numeric for model
data3 %>% select(coral_id, time, polyp, wound, temp, genotype, percent_affected)%>%group_by(coral_id) %>% filter(lag(polyp)!=1)%>%ungroup()%>%filter(polyp == 1 | time == 19) %>% mutate(time = as.numeric(time), polyp = as.character(polyp), polyp = as.numeric(polyp))%>%unite("treatment", wound:temp, remove = FALSE)%>% 
  mutate(wound = case_when(wound == 1 ~ "Fragmentation",
                           wound == 2 ~ "Abrasion"),
         wound = as.factor(wound)) -> polyp.survival
```
survival curve
```{r}
survfit(Surv(time, polyp) ~ wound + temp , data = polyp.survival)
surv.mod_polyp <- survfit(Surv(time, polyp) ~ wound + temp , data = polyp.survival)
summary(surv.mod)
```
cox model (mixed effects model with genotype as random effect)
```{r}
cox.zph(coxme(Surv(time, polyp)~wound*temp + (1|genotype) , data = polyp.survival))
coxme(Surv(time, polyp)~wound*temp + (1|genotype) , data = polyp.survival)
coxme(Surv(time, polyp)~wound*temp + (1|genotype), data = polyp.survival) %>% tbl_regression(exp = T)
```
using treatment
```{r}
cox.zph(coxme(Surv(time, polyp)~treatment + (1|genotype) , data = polyp.survival))
coxme(Surv(time, polyp)~treatment + (1|genotype) , data = polyp.survival)
coxme(Surv(time, polyp)~treatment + (1|genotype), data = polyp.survival) %>% tbl_regression(exp = T)
```
```{r}
# List of ggsurvplots
splots <- list()
```


survival figure
```{r}
ggsurvplot(
   fit = surv.mod_polyp,                     # survfit object with calculated statistics.
   data = polyp.survival,             # data used to fit survival curves.
   #risk.table = TRUE,       # show risk table.
   conf.int = FALSE,         # show confidence intervals for 
                           # point estimates of survival curves.
                            # survival estimates.
   xlab = " ",
   ylab = "Proportion With New Polyps",# customize X axis label
   ggtheme = theme_classic(), # customize plot and risk table with a theme.
 #risk.table.y.text.col = T, # colour risk table text annotations.
 # risk.table.y.text = FALSE, # show bars instead of names in text annotations
  legend = "none",
            # legend.title = "Treatments", legend.labs = c("FA", "FE", "AA", "AE"), legend = "top",
 fun = "event",
 facet.by = "wound",
 legend.labs = c("Ambient", "Warming"),
 palette = c("blue", "red"),
 short.panel.labs = T
 
                            # in legend of risk table
) -> splots_1
  


```

# CORALLITE
survival dataframe
```{r}
#clean data frame for columns we want, change timepoint and healed to numeric for model
data3 %>% select(coral_id, time, corallite, wound, temp, genotype, percent_affected)%>%group_by(coral_id) %>% filter(lag(corallite)!=1)%>%ungroup()%>%filter(corallite == 1 | time == 19) %>% mutate(time = as.numeric(time), corallite = as.character(corallite), corallite = as.numeric(corallite))%>%unite("treatment", wound:temp, remove = FALSE) %>% 
  mutate(wound = case_when(wound == 1 ~ "Fragmentation",
                           wound == 2 ~ "Abrasion"),
         wound = as.factor(wound)) -> corallite.survival
```
survival curve
```{r}
survfit(Surv(time, corallite) ~ wound + temp , data = corallite.survival)
surv.mod_corallite <- survfit(Surv(time, corallite) ~ wound + temp , data = corallite.survival)
summary(surv.mod)
```
cox model (mixed effects model with genotype as random effect), interaction wound*temp, breaks assumption so doing only single fixed effects, interaction was not significant anyway
```{r}
cox.zph(coxme(Surv(time, corallite)~wound + temp + (1|genotype) , data = corallite.survival))
coxme(Surv(time, corallite)~wound + temp + (1|genotype) , data = corallite.survival)
coxme(Surv(time, corallite)~wound + temp + (1|genotype), data = corallite.survival) %>% tbl_regression(exp = T)
```
breaks assumption. interaction also breaks assumption^ 
```{r}
cox.zph(coxme(Surv(time, corallite)~treatment + (1|genotype) , data = corallite.survival))
coxme(Surv(time, corallite)~treatment + (1|genotype) , data = corallite.survival)
coxme(Surv(time, corallite)~treatment + (1|genotype), data = corallite.survival) %>% tbl_regression(exp = T)
```


survival figure
```{r}
ggsurvplot(
   surv.mod_corallite,                     # survfit object with calculated statistics.
   data = corallite.survival,             # data used to fit survival curves.
  # risk.table = TRUE,       # show risk table.
   conf.int = FALSE,         # show confidence intervals for 
  ylab = "Proportion With New Corallites",
                            # point estimates of survival curves.
                          # survival estimates.
   xlab = " ",   # customize X axis label
   ggtheme = theme_classic(), # customize plot and risk table with a theme.
# risk.table.y.text.col = T, # colour risk table text annotations.
 # risk.table.y.text = FALSE, # show bars instead of names in text annotations
            # legend.title = "Treatments", legend.labs = c("FA", "FE", "AA", "AE"), legend = "top",
legend = "right",
  fun = "event",
 facet.by = "wound",
 legend.labs = c("Ambient", "Warming"),
 palette = c("blue", "red")
  

                            # in legend of risk table
) +
  theme(strip.text = element_blank())-> splots_2
```

# NEW TISSUE
survival dataframe
```{r}
#clean data frame for columns we want, change timepoint and healed to numeric for model
data3%>%select(coral_id, time, new_tissue, wound, temp, genotype, percent_affected)%>%group_by(coral_id)%>%filter(lag(new_tissue)!=1)%>%ungroup()%>%filter(new_tissue == 1 | time == 19) %>% mutate(time = as.numeric(time), new_tissue = as.character(new_tissue), new_tissue = as.numeric(new_tissue))%>%unite("treatment", wound:temp, remove = FALSE)%>% 
  mutate(wound = case_when(wound == 1 ~ "Fragmentation",
                           wound == 2 ~ "Abrasion"),
         wound = as.factor(wound)) -> newtissue.survival
```
survival curve
```{r}
survfit(Surv(time, new_tissue) ~ wound + temp , data = newtissue.survival)
surv.mod_tissue <- survfit(Surv(time, new_tissue) ~ wound + temp , data = newtissue.survival)
summary(surv.mod)
```
cox model (mixed effects model with genotype as random effect)
```{r}
cox.zph(coxme(Surv(time, new_tissue)~wound*temp + (1|genotype) , data = newtissue.survival))
coxme(Surv(time, new_tissue)~wound*temp + (1|genotype) , data = newtissue.survival)
coxme(Surv(time, new_tissue)~wound*temp + (1|genotype), data = newtissue.survival) %>% tbl_regression(exp = T)
```

using treatment instead of wound and temp gives the same results
```{r}
cox.zph(coxme(Surv(time, new_tissue)~treatment + (1|genotype) , data = newtissue.survival))
coxme(Surv(time, new_tissue)~treatment + (1|genotype) , data = newtissue.survival)
coxme(Surv(time, new_tissue)~treatment + (1|genotype), data = newtissue.survival) %>% tbl_regression(exp = T)
```

survival figure
```{r}
ggsurvplot(
   surv.mod_tissue,                     # survfit object with calculated statistics.
   data = newtissue.survival,             # data used to fit survival curves.
  # risk.table = TRUE,       # show risk table.
   conf.int = FALSE,         # show confidence intervals for 
                            # point estimates of survival curves.
                            # survival estimates.
   xlab = "Time (days)",   # customize X axis label
  ylab = "Proportion With New Tissue",
   ggtheme = theme_classic(), # customize plot and risk table with a theme.
 #risk.table.y.text.col = T, # colour risk table text annotations.
 # risk.table.y.text = FALSE, # show bars instead of names in text annotations
             legend.title = "Treatments", legend = "none",
 fun = "event",
 facet.by = "wound",
 legend.labs = c("Ambient", "Warming"),
 palette = c("blue", "red"),
 short.panel.labs = T
                            # in legend of risk table
)+
  theme(strip.text = element_blank()) -> splots_3
```


```{r}

library(patchwork)

surv.plot <- ((splots_1& theme(plot.tag.position = c(0,1)))/(splots_2& theme(plot.tag.position = c(0,1)))/(splots_3&  
  theme(plot.tag.position  = c(0,1)))) + plot_annotation(tag_levels = "A")



ggsave("survival.jpg", path = here(), plot = surv.plot,
       width = 15,
       height = 20,
       units = "in", 
       scale = 0.6)

```

