---
title: "Quantifying Wound Area and total loss of tissue surface area due to Injuries in ARC regen proj"
author: "Ninah Munk"
date: "2024-06-07"
output: html_document
---

load packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(janitor)

```

# load data: fragment.area is the wound closure data which lexi collected in imagej. SA.lost are the dimensions of removed coral branches as a result of the fragmentation injury.  
```{r}
fragment.area<- read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Data/fragment.only.measurements.csv")%>%clean_names()
abrasion.area<- read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Data/abrasion_msmts.csv")%>%clean_names()

frag.SA.lost<-read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Surface_Area/Data/wound_type_1_measurements.csv")%>%clean_names()

```

# clean data
```{r}
fragment.area$time_point_initial <- factor(fragment.area$time_point_initial, levels = c("initial", "24hr", "day10", "final"), labels = c("0", "1", "10", "19")) 
```



# clean frag.SA.lost data and calculate SA of the coral branch that was removed; MEAN SURFACE AREA LOST BY CREATING THE FRAGMENTATION WOUND WAS 1.0717 cm2
```{r}
frag.SA.lost%>%drop_na()-> frag.SA.lost # only fragmented corals now

frag.SA.lost%>%
  mutate(branch_height_cm = branch_h_mm / 10) %>%
  mutate(avg_diameter_mm= (base_d_mm + tip_d_mm) / 2)%>%
  mutate(radius_cm = (avg_diameter_mm/2)/10)%>%
  mutate(radius_tip_cm = (tip_d_mm/2)/10)%>%
  mutate(SA.lost= (2*3.14*(radius_cm*branch_height_cm) + 3.14*(radius_tip_cm)^2)) -> frag.SA.lost

summary(frag.SA.lost)
summarySE(data = frag.SA.lost, measurevar = "SA.lost",
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE)

frag.SA.lost%>%select(coral_id, genotype, wound, temperature, SA.lost)-> frag.SA
```

# caluclate SA of abrasion injury using lengths of the abrased area collected in imagej by lexi, these measurements are day 1 timepoint. MEAN SURFACE AREA LOST AND LEFT BEHIND BY CREATING THE ABRASION WOUND WAS 0.9493 cm2
```{r}
# first getting average branch radius and average tip radius to use for calculating SA of abrasion injury
abrasion.radius<- mean(frag.SA.lost$radius_cm) # 0.1495 cm
abrasion.radius.tip<- mean(frag.SA.lost$radius_tip_cm) #0.1008056 cm

abrasion.area%>%
  mutate(branch_height_cm = length_mm /10)%>%
  mutate(radius_cm = 0.1495)%>%
  mutate(radius_tip_cm = 0.1008056)%>%
  mutate(SA.lost= (2*3.14*(radius_cm*branch_height_cm) + 3.14*(radius_tip_cm)^2)) -> abrasion.SA

summarySE(data = abrasion.SA, measurevar = "SA.lost",
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE)

summary(abrasion.SA)
```

# WHAT WAS THE AVERAGE SURFACE AREA LEFT BEHIND BY THE FRAGMENTATION WOUND? derived from imagej: 0.2228  cm2
```{r}
fragment.area%>%
  filter(time_point_initial == 0)%>%
  mutate(area.cm2 = area_mm2 / 100)-> frag.initial.area

summarySE(data = frag.initial.area, measurevar = "area.cm2",
          na.rm = FALSE, conf.interval = 0.95, .drop = TRUE)
summary(frag.initial.area)
```

# WHAT WAS THE AVERAGE SURFACE AREA LEFT BEHIND BY THE FRAGMENTATION WOUND? derived from fragment base diameter: 0.12875 cm2
```{r}
frag.SA.lost%>%
  mutate(base_d_cm = base_d_mm / 10)%>%
  mutate(base.r.cm = base_d_cm/2)%>%
  mutate(area.base = 3.14*((base.r.cm)^2))-> area.base
summary(area.base)
```



