---
title: "Regeneration.5"
author: "Ninah Munk"
date: "2024-06-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggfortify)
library(survival)
library(gtsummary)
library(rstatix)
library(Rmisc)
library(lmerTest)
library(tidyverse)
library(janitor)
library(emmeans)
library(plyr)
library(ggthemes)
library(survminer)
library(coxme)
```


# Load data and combine 
```{r}
read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Data/healing.signs.data.ninah.csv")%>%select(-wound) -> regen
read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Respiration/Data/samp_info.csv")-> samp.info
read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Surface_Area/Output/post_wound_surface_areas.csv") -> SA
read.csv("/Users/ninahmunk/Desktop/Projects/Acropora_Regeneration-main/Wound_Closure/Data/worm_damage.csv")%>%select(coral_id, affected, percent_affected)-> worm

regen%>%left_join(samp.info, by = "coral_id")%>%left_join(SA, by = "coral_id")%>%left_join(worm, by = "coral_id") -> data

```


# modify data set 
```{r}

data%>%mutate(polyp = as.factor(polyp), 
               corallite = as.factor(corallite),
               new_tissue = as.factor(new_tissue),
               healed.full = as.factor(healed.full),
               healed.2.3 = as.factor(healed.2.3),
               wound = as.factor(wound),
               temp = as.factor(temp),
               genotype = as.factor(genotype))-> data2
```

# Clean Data
```{r}
#clean data
corals_to_remove <- c(11, 5, 37, 104, 71, 17, 29, 100)
data2 <- data2[!(data$coral_id %in% corals_to_remove), ]
```


# using polyp as event. Here I am running mixed effects cox models with percent affected as a covariate
```{r}
#clean data frame for columns we want, change timepoint and healed to numeric for model
data2%>%select(coral_id, time, polyp, wound, temp, genotype, percent_affected)%>%group_by(coral_id)%>%filter(lag(polyp)!=1)%>%ungroup()%>%filter(polyp == 1 | time == 19) %>% mutate(time = as.numeric(time), polyp = as.character(polyp), polyp = as.numeric(polyp))%>%unite("treatment", wound:temp, remove = FALSE)-> survival

cox.zph(coxme(Surv(time, polyp)~wound*temp*percent_affected + (1|genotype), data = survival))
coxme(Surv(time, polyp)~wound*temp*percent_affected + (1|genotype), data = survival)
coxme(Surv(time, polyp)~wound*temp*percent_affected + (1|genotype), data = survival) %>% tbl_regression(exp = T) 

cox.zph(coxme(Surv(time, polyp)~treatment*percent_affected + (1|genotype), data = survival))
coxme(Surv(time, polyp)~treatment*percent_affected+ (1|genotype), data = survival)
coxme(Surv(time, polyp)~treatment*percent_affected + (1|genotype), data = survival) %>% tbl_regression(exp =T)

```

# using corallite as event with percent affected as random effect
```{r}
#clean data frame for columns we want, change timepoint and healed to numeric for model
data2 %>% select(coral_id, time, corallite, wound, temp, genotype, percent_affected)%>%group_by(coral_id) %>% filter(lag(corallite)!=1)%>%ungroup()%>%filter(corallite == 1 | time == 19) %>% mutate(time = as.numeric(time), corallite = as.character(corallite), corallite = as.numeric(corallite))%>%unite("treatment", wound:temp, remove = FALSE) -> survival

cox.zph(coxme(Surv(time, corallite)~wound*temp + (1|genotype) + (1|percent_affected), data = survival))
coxme(Surv(time, corallite)~wound*temp + (1|genotype) + (1|percent_affected), data = survival)
coxme(Surv(time, corallite)~wound*temp + (1|genotype), data = survival) %>% tbl_regression(exp = T)

cox.zph(coxme(Surv(time, corallite)~treatment + (1|genotype) + (1|percent_affected), data = survival))
coxme(Surv(time, corallite)~treatment+ (1|genotype) + (1|percent_affected), data = survival)
coxme(Surv(time, corallite)~treatment + (1|genotype) + (1|percent_affected), data = survival) %>% tbl_regression(exp =T)
```

# using new tissue as event with percent affected as covariate
```{r}
#clean data frame for columns we want, change timepoint and healed to numeric for model
data2%>%select(coral_id, time, new_tissue, wound, temp, genotype, percent_affected)%>%group_by(coral_id)%>%filter(lag(new_tissue)!=1)%>%ungroup()%>%filter(new_tissue == 1 | time == 19) %>% mutate(time = as.numeric(time), new_tissue = as.character(new_tissue), new_tissue = as.numeric(new_tissue))%>%unite("treatment", wound:temp, remove = FALSE)-> survival

cox.zph(coxme(Surv(time, new_tissue)~wound*temp*percent_affected + (1|genotype) , data = survival))
coxme(Surv(time, new_tissue)~wound*temp*percent_affected + (1|genotype), data = survival)
coxme(Surv(time, new_tissue)~wound*temp + (1|genotype), data = survival) #reduced
coxme(Surv(time, new_tissue)~wound*temp + (1|genotype) + (1|percent_affected), data = survival) %>% tbl_regression(exp = T)

cox.zph(coxme(Surv(time, new_tissue)~treatment + (1|genotype) + (1|percent_affected), data = survival))
coxme(Surv(time, new_tissue)~treatment+ (1|genotype) + (1|percent_affected), data = survival)
coxme(Surv(time, new_tissue)~treatment + (1|genotype) + (1|percent_affected), data = survival) %>% tbl_regression(exp =T)
```

#  having 2 out of 3 healing metrics as an event
```{r}
#clean data frame for columns we want, change timepoint and healed to numeric for model
data2%>%select(coral_id, time, healed.2.3, wound, temp, genotype, percent_affected)%>%group_by(coral_id)%>%filter(lag(healed.2.3)!=1)%>%ungroup()%>%filter(healed.2.3 == 1 | time == 19) %>% mutate(time = as.numeric(time), healed.2.3 = as.character(healed.2.3), healed.2.3 = as.numeric(healed.2.3))%>%unite("treatment", wound:temp, remove = FALSE)-> survival

cox.zph(coxme(Surv(time, healed.2.3)~wound*temp + (1|genotype) + (1|percent_affected), data = survival))
coxme(Surv(time, healed.2.3)~wound*temp + (1|genotype) + (1|percent_affected), data = survival)
coxme(Surv(time, healed.2.3)~wound*temp + (1|genotype) + (1|percent_affected), data = survival) %>% tbl_regression(exp = T)

cox.zph(coxme(Surv(time, healed.2.3)~treatment + (1|genotype) + (1|percent_affected), data = survival))
coxme(Surv(time, healed.2.3)~treatment+ (1|genotype) + (1|percent_affected), data = survival)
coxme(Surv(time, healed.2.3)~treatment + (1|genotype) + (1|percent_affected), data = survival) %>% tbl_regression(exp =T)
```



# FILTERING DATA TO REMOVE WORM EFFECT
Here I filter the data to use only the corals that had low worm damage. This threshold is <= 10% worm damage and was determined using the growth rate dataset. The remaining sample size is N = 16
```{r}
data3 = subset(data2, percent_affected <= 10)
```

# 2 out of 3 healing metrics as event with filtered data
```{r}
#clean data frame for columns we want, change timepoint and healed to numeric for model
data3 %>% select(coral_id, time, healed.2.3, wound, temp, genotype, percent_affected)%>%group_by(coral_id) %>% filter(lag(healed.2.3)!=1)%>%ungroup()%>%filter(healed.2.3 == 1 | time == 19) %>% mutate(time = as.numeric(time), healed.2.3 = as.character(healed.2.3), healed.2.3 = as.numeric(healed.2.3))%>%unite("treatment", wound:temp, remove = FALSE) -> survival

cox.zph(coxme(Surv(time, healed.2.3)~wound*temp + (1|genotype) , data = survival))
coxme(Surv(time, healed.2.3)~wound*temp + (1|genotype) , data = survival)
coxme(Surv(time, healed.2.3)~wound*temp + (1|genotype), data = survival) %>% tbl_regression(exp = T)

cox.zph(coxme(Surv(time, healed.2.3)~treatment + (1|genotype), data = survival))
coxme(Surv(time, healed.2.3)~treatment+ (1|genotype), data = survival)
coxme(Surv(time, healed.2.3)~treatment + (1|genotype), data = survival) %>% tbl_regression(exp =T)


```

# polyp as event with filtered data
Polyp development is the same between treatments
```{r}
#clean data frame for columns we want, change timepoint and healed to numeric for model
data3 %>% select(coral_id, time, polyp, wound, temp, genotype, percent_affected)%>%group_by(coral_id) %>% filter(lag(polyp)!=1)%>%ungroup()%>%filter(polyp == 1 | time == 19) %>% mutate(time = as.numeric(time), polyp = as.character(polyp), polyp = as.numeric(polyp))%>%unite("treatment", wound:temp, remove = FALSE) -> survival

cox.zph(coxme(Surv(time, polyp)~wound*temp + (1|genotype) , data = survival))
coxme(Surv(time, polyp)~wound*temp + (1|genotype) , data = survival)
coxme(Surv(time, polyp)~wound*temp + (1|genotype), data = survival) %>% tbl_regression(exp = T)

cox.zph(coxme(Surv(time, polyp)~treatment + (1|genotype), data = survival))
coxme(Surv(time, polyp)~treatment+ (1|genotype), data = survival)
coxme(Surv(time, polyp)~treatment + (1|genotype), data = survival) %>% tbl_regression(exp =T)
```
# corallite as event with filtered data
```{r}
#clean data frame for columns we want, change timepoint and healed to numeric for model
data3 %>% select(coral_id, time, corallite, wound, temp, genotype, percent_affected)%>%group_by(coral_id) %>% filter(lag(corallite)!=1)%>%ungroup()%>%filter(corallite == 1 | time == 19) %>% mutate(time = as.numeric(time), corallite = as.character(corallite), corallite = as.numeric(corallite))%>%unite("treatment", wound:temp, remove = FALSE) -> survival

cox.zph(coxme(Surv(time, corallite)~wound*temp + (1|genotype) , data = survival))
coxme(Surv(time, corallite)~wound*temp + (1|genotype) , data = survival)
coxme(Surv(time, corallite)~wound*temp + (1|genotype), data = survival) %>% tbl_regression(exp = T)

cox.zph(coxme(Surv(time, corallite)~treatment + (1|genotype), data = survival))
coxme(Surv(time, corallite)~treatment+ (1|genotype), data = survival)
coxme(Surv(time, corallite)~treatment + (1|genotype), data = survival) %>% tbl_regression(exp =T)
```
# using new tissue as event with filtered data
```{r}
#clean data frame for columns we want, change timepoint and healed to numeric for model
data3%>%select(coral_id, time, new_tissue, wound, temp, genotype, percent_affected)%>%group_by(coral_id)%>%filter(lag(new_tissue)!=1)%>%ungroup()%>%filter(new_tissue == 1 | time == 19) %>% mutate(time = as.numeric(time), new_tissue = as.character(new_tissue), new_tissue = as.numeric(new_tissue))%>%unite("treatment", wound:temp, remove = FALSE)-> survival

cox.zph(coxme(Surv(time, new_tissue)~wound*temp + (1|genotype) , data = survival))
coxme(Surv(time, new_tissue)~wound*temp + (1|genotype), data = survival)
coxme(Surv(time, new_tissue)~wound*temp + (1|genotype) , data = survival) %>% tbl_regression(exp = T)

cox.zph(coxme(Surv(time, new_tissue)~treatment + (1|genotype) , data = survival))
coxme(Surv(time, new_tissue)~treatment+ (1|genotype) , data = survival)
coxme(Surv(time, new_tissue)~treatment + (1|genotype) , data = survival) %>% tbl_regression(exp =T)
```

# new column representing polyp and new tissue 
```{r}
data4 <- data3 %>%
  mutate(polyp.tissue=case_when(
    polyp=="1" & new_tissue=="1" ~ "1",
    polyp=="0" & new_tissue=="0" ~ "0",
    polyp=="1" & new_tissue=="0" ~ "0",
    polyp=="0" & new_tissue=="1" ~ "0"
  ))
```

```{r}
#clean data frame for columns we want, change timepoint and healed to numeric for model
data4%>%select(coral_id, time, polyp.tissue, wound, temp, genotype, percent_affected)%>%group_by(coral_id)%>%filter(lag(polyp.tissue)!=1)%>%ungroup()%>%filter(polyp.tissue == 1 | time == 19) %>% mutate(time = as.numeric(time), polyp.tissue = as.character(polyp.tissue), polyp.tissue = as.numeric(polyp.tissue))%>%unite("treatment", wound:temp, remove = FALSE)-> survival

cox.zph(coxme(Surv(time, polyp.tissue)~wound*temp + (1|genotype) , data = survival))
coxme(Surv(time, polyp.tissue)~wound*temp + (1|genotype), data = survival)
coxme(Surv(time, polyp.tissue)~wound*temp + (1|genotype) , data = survival) %>% tbl_regression(exp = T)

cox.zph(coxme(Surv(time, polyp.tissue)~treatment + (1|genotype) , data = survival))
coxme(Surv(time, polyp.tissue)~treatment+ (1|genotype) , data = survival)
coxme(Surv(time, polyp.tissue)~treatment + (1|genotype) , data = survival) %>% tbl_regression(exp =T)
```


```{r}
data5 <- data3 %>%
  mutate(corallite.tissue=case_when(
    corallite=="1" & new_tissue=="1" ~ "1",
    corallite=="0" & new_tissue=="0" ~ "0",
    corallite=="1" & new_tissue=="0" ~ "0",
    corallite=="0" & new_tissue=="1" ~ "0"
  ))

```

```{r}
#clean data frame for columns we want, change timepoint and healed to numeric for model
data5%>%select(coral_id, time, corallite.tissue, wound, temp, genotype, percent_affected)%>%group_by(coral_id)%>%filter(lag(corallite.tissue)!=1)%>%ungroup()%>%filter(corallite.tissue == 1 | time == 19) %>% mutate(time = as.numeric(time), corallite.tissue = as.character(corallite.tissue), corallite.tissue = as.numeric(corallite.tissue))%>%unite("treatment", wound:temp, remove = FALSE)-> survival

cox.zph(coxme(Surv(time, corallite.tissue)~wound*temp + (1|genotype) , data = survival))
coxme(Surv(time, corallite.tissue)~wound*temp + (1|genotype), data = survival)
coxme(Surv(time, corallite.tissue)~wound*temp + (1|genotype) , data = survival) %>% tbl_regression(exp = T)

cox.zph(coxme(Surv(time, polyp.tissue)~treatment + (1|genotype) , data = survival))
coxme(Surv(time, polyp.tissue)~treatment+ (1|genotype) , data = survival)
coxme(Surv(time, polyp.tissue)~treatment + (1|genotype) , data = survival) %>% tbl_regression(exp =T)
```
```{r}
data3 = subset(data2, percent_affected <= 10)
```

```{r}
data3%>%select(coral_id,time, polyp, genotype, wound, temp)%>%group_by(coral_id)%>%filter(lag(polyp)!=1)%>%ungroup()%>%filter(polyp == 1 | time == 19)%>%mutate(event = "polyp")%>%rename(status = polyp) -> polyp
data3%>%select(coral_id,time, corallite, genotype, wound, temp)%>%group_by(coral_id)%>%filter(lag(corallite)!=1)%>%ungroup()%>%filter(corallite == 1 | time == 19)%>%mutate(event = "corallite")%>%rename(status = corallite)  -> corallite
data3%>%select(coral_id,time, new_tissue, genotype, wound, temp)%>%group_by(coral_id)%>%filter(lag(new_tissue)!=1)%>%ungroup()%>%filter(new_tissue == 1 | time == 19)%>%mutate(event = "new_tissue")%>%rename(status = new_tissue)  -> new_tissue

rbind(polyp, corallite, new_tissue)-> full

full%>%mutate(time = as.numeric(time), status = as.character(status), status = as.numeric(status))%>%unite("treatment", wound:temp, remove = FALSE)-> data6

# Convert the event variable to a factor
data6$event <- factor(data6$event)

cox.zph(coxph(Surv(time, status) ~ wound * temp , data = data6%>%filter(event == "polyp"), id = coral_id))
fit <- coxph(Surv(time, status) ~ wound * temp , data = data6%>%filter(event == "polyp"), id = coral_id)
summary(fit)
coxph(Surv(time, status) ~ wound * temp , data = data6%>%filter(event == "polyp"), id = coral_id)%>% tbl_regression(exp =T)

cox.zph(coxph(Surv(time, status) ~ wound * temp , data = data6%>%filter(event == "corallite"), id = coral_id))
fit <- coxph(Surv(time, status) ~ wound * temp , data = data6%>%filter(event == "corallite"), id = coral_id)
summary(fit)
coxph(Surv(time, status) ~ wound * temp , data = data6%>%filter(event == "corallite"), id = coral_id)%>% tbl_regression(exp =T)

cox.zph(coxph(Surv(time, status) ~ wound * temp , data = data6%>%filter(event == "new_tissue"), id = coral_id))
fit <- coxph(Surv(time, status) ~ wound * temp, data = data6%>%filter(event == "new_tissue"), id = coral_id)
summary(fit)
coxph(Surv(time, status) ~ wound * temp , data = data6%>%filter(event == "new_tissue"), id = coral_id)%>% tbl_regression(exp =T)


```

```{r}
# survfit is model function (like  a glm), surv() function to create dependent df (when the event happened (time) and what was event (event)) as a function of wound

survfit(Surv(time, status) ~ wound + temp , data = data6%>%filter(event == "polyp"))
surv.mod <- survfit(Surv(time, status) ~ wound + temp , data = data6%>%filter(event == "polyp"))
summary(surv.mod)

survfit(Surv(time, status) ~ wound + temp , data = data6%>%filter(event == "corallite"))
surv.mod <- survfit(Surv(time, status) ~ wound + temp , data = data6%>%filter(event == "corallite"))
summary(surv.mod)

survfit(Surv(time, status) ~ wound + temp , data = data6%>%filter(event == "new_tissue"))
surv.mod <- survfit(Surv(time, status) ~ wound + temp , data = data6%>%filter(event == "new_tissue"))
summary(surv.mod)


```


```{r}
library(survival)
library(survminer)

# Fit survival models for each event type
polyp_mod <- survfit(Surv(time, status) ~ wound + temp, data = data6 %>% filter(event == "polyp"))
corallite_mod <- survfit(Surv(time, status) ~ wound + temp, data = data6 %>% filter(event == "corallite"))
new_tissue_mod <- survfit(Surv(time, status) ~ wound + temp, data = data6 %>% filter(event == "new_tissue"))

# Plot survival curves
ggsurvplot(polyp_mod, data = data6 %>% filter(event == "polyp"), 
                         pval = TRUE, title = "Survival Curves for Polyp",
                         legend.title = "Treatments", legend.labs = c("FA", "FE", "AA", "AE"), legend = "top")


 ggsurvplot(corallite_mod, data = data6 %>% filter(event == "corallite"), 
                             pval = TRUE, title = "Survival Curves for Corallite",
             legend.title = "Treatments", legend.labs = c("FA", "FE", "AA", "AE"), legend = "top")


 ggsurvplot(new_tissue_mod, data = data6 %>% filter(event == "new_tissue"), 
                              pval = TRUE, title = "Survival Curves for New Tissue",
             legend.title = "Treatments", legend.labs = c("FA", "FE", "AA", "AE"), legend = "top")


```

```{r}
polyp_mod <- survfit(Surv(time, status) ~ wound + temp, data = data6 %>% filter(event == "polyp"))

ggsurvplot(
   polyp_mod,                     # survfit object with calculated statistics.
   data = data6%>% filter(event == "polyp"),             # data used to fit survival curves.
   risk.table = TRUE,       # show risk table.
   conf.int = TRUE,         # show confidence intervals for 
                            # point estimates of survival curves.
                            # survival estimates.
   xlab = "Time in days",   # customize X axis label
   ggtheme = theme_light(), # customize plot and risk table with a theme.
 risk.table.y.text.col = T, # colour risk table text annotations.
  risk.table.y.text = FALSE, # show bars instead of names in text annotations
             legend.title = "Treatments", legend.labs = c("FA", "FE", "AA", "AE"), legend = "top"
                            # in legend of risk table
)

corallite_mod <- survfit(Surv(time, status) ~ wound + temp, data = data6 %>% filter(event == "corallite"))

ggsurvplot(
   corallite_mod,                     # survfit object with calculated statistics.
   data = data6%>% filter(event == "corallite"),             # data used to fit survival curves.
   risk.table = TRUE,       # show risk table.
   conf.int = TRUE,         # show confidence intervals for 
                            # point estimates of survival curves.
                            # survival estimates.
   xlab = "Time in days",   # customize X axis label
   ggtheme = theme_light(), # customize plot and risk table with a theme.
 risk.table.y.text.col = T, # colour risk table text annotations.
  risk.table.y.text = FALSE, # show bars instead of names in text annotations
             legend.title = "Treatments", legend.labs = c("FA", "FE", "AA", "AE"), legend = "top"
                            # in legend of risk table
)

new_tissue_mod <- survfit(Surv(time, status) ~ wound + temp, data = data6 %>% filter(event == "new_tissue"))

ggsurvplot(
   new_tissue_mod,                     # survfit object with calculated statistics.
   data = data6%>% filter(event == "new_tissue"),             # data used to fit survival curves.
   risk.table = TRUE,       # show risk table.
   conf.int = TRUE,         # show confidence intervals for 
                            # point estimates of survival curves.
                            # survival estimates.
   xlab = "Time in days",   # customize X axis label
   ggtheme = theme_light(), # customize plot and risk table with a theme.
 risk.table.y.text.col = T, # colour risk table text annotations.
  risk.table.y.text = FALSE, # show bars instead of names in text annotations
             legend.title = "Treatments", legend.labs = c("FA", "FE", "AA", "AE"), legend = "top"
                            # in legend of risk table
)

```

