---
title: "Temperature treatments in ACR regeneration experiment"
author: "Ninah Munk"
date: "2024-05-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(janitor)
library(emmeans)
library(lme4)
library(Matrix)
library(readxl)
library(tidyr)
```

# Temperature Data 
## Load raw hobo data files, rename columns, separate date/time column
```{r}

hotsump<-read.csv('21512790_hot_sump_20230703.csv')
colnames(hotsump)[2] = "date_time"
colnames(hotsump)[3] = "temp_c"

hotsump%>%separate(col = 'date_time', into = c('date', 'time'), sep= ' ')%>%
  select(date, time, temp_c)%>%
  drop_na()-> hotsump_date

ambientsump<-read.csv('21512796_ambient_sump_20230703.csv')
colnames(ambientsump)[2] = "date_time"
colnames(ambientsump)[3] = "temp_c"

ambientsump%>%separate(col = 'date_time', into = c('date', "time"), sep = ' ')%>%
  select(date, time, temp_c)%>%
  drop_na()-> ambientsump_date


```

## Perform Mann-Whitney U test
```{r}
beep <- t.test(hotsump_date$temp_c, ambientsump_date$temp_c)
print(beep)

result <- wilcox.test(hotsump_date$temp_c, ambientsump_date$temp_c)
print(result)

hotsump_date%>%mutate(treatment = "elevated")-> elevated.data
ambientsump_date%>%mutate(treatment = 'ambient')-> ambient.data

rbind(elevated.data, ambient.data)-> full.temp.data


```


## Visualize variation in daily temperautre across each day of the experiment in the hot sump 
```{r}

ggplot(data = hotsump_date, aes(x = as.Date(date, format = "%m / %d / %Y"), y = temp_c))+
  geom_point(size = 1, color = "red")+  
  labs(title = "Raw temperature data - Elevated", y = "Temperature (째C)", x = "Date")+ 
  theme_bw()


```

## create data frame with separate columns for month, day, year to calculate daily mean temperatures
```{r}

hotsump_date%>%separate(col = 'date',
                  into = c('month', 'day', 'year'),
                  sep= '/',
                  remove = FALSE) -> hotsumpfull

hotsumpfull%>%group_by(year, month, day, date)%>%
  summarise(meantemp = mean(temp_c)) -> hotsump_daily_mean

# Calculating mean temp for each date to plot.
hotsumpfull%>%aggregate(temp_c ~ date, FUN = mean)-> hotsump_daily_mean
# Calculating standard error for error bars in the plot.
hotsumpfull%>%aggregate(temp_c ~ date, FUN = function(x) sd(x)/sqrt(length(x))) -> std_err 


ggplot(hotsump_daily_mean, aes(x=date, y=temp_c)) +
  geom_point(size = 1, color = "red") +
  geom_errorbar(aes(ymin=temp_c-std_err$temp_c, ymax=temp_c+std_err$temp_c), width=.2) +
  labs(title="Mean Daily Temperatue - Elevated", x="Date", y="Temperature (째C)")+ 
  theme_classic()
```
## Visualize variation in daily temperautre across each day of the experiment in the ambient sump 
```{r}
ggplot(data = ambientsump_date, aes(x = as.Date(date, format = "%m / %d / %Y"), y = temp_c))+
  geom_point(size = 1, color = "red")+  
  labs(title = "Raw temperature data - Ambient", y = "Temperature (째C)", x = "Date")+ 
  theme_bw()

```

# Calculating mean temp for each date to plot.
```{r}
ambientsump_date%>%separate(col = 'date',
                  into = c('month', 'day', 'year'),
                  sep= '/',
                  remove = FALSE) -> ambientsumpfull

ambientsumpfull%>%group_by(year, month, day, date)%>%
  summarise(meantemp = mean(temp_c)) -> ambientsump_daily_mean

# Calculating mean temp for each date to plot.
ambientsumpfull%>%aggregate(temp_c ~ date, FUN = mean)-> ambientsump_daily_mean
# Calculating standard error for error bars in the plot.
ambientsumpfull%>%aggregate(temp_c ~ date, FUN = function(x) sd(x)/sqrt(length(x))) -> std_err2


ggplot(ambientsump_daily_mean, aes(x=date, y=temp_c)) +
  geom_point(size = 1, color = "red") +
  geom_errorbar(aes(ymin=temp_c-std_err2$temp_c, ymax=temp_c+std_err2$temp_c), width=.2) +
  labs(title="Mean Daily Temperatue - Ambient", x="Date", y="Temperature (째C)")+ 
  theme_classic()
```

## join both sumps temperature data together
```{r}
hotsump_daily_mean%>%rename(elevated_temp_c = "temp_c")-> hotsump_daily_mean
ambientsump_daily_mean%>%rename(ambient_temp_c = "temp_c")-> ambientsump_daily_mean

hotsump_daily_mean%>%left_join(ambientsump_daily_mean, by = "date")-> mean_full

 mean_full%>%pivot_longer(cols = c(elevated_temp_c, ambient_temp_c),
               names_to = "treatment",
               values_to = "temp_c") -> mean_full

std_err%>%rename(SE = "temp_c")-> std_err
std_err2%>%rename(SE = "temp_c")-> std_err2


mean_full%>%filter(treatment == "elevated_temp_c")%>%left_join(std_err, by = "date")-> elevated
mean_full%>%filter(treatment == "ambient_temp_c")%>%left_join(std_err2, by = "date")-> ambient

rbind(elevated, ambient)-> sump.temp.final

ggplot( data = sump.temp.final, aes(x= date, y = temp_c, col = treatment))+ geom_point(size = 1)


```

